### 2009-11-10 report/paper.tex
added preliminary version of the report.

\documentclass[a4paper]{llncs}
\usepackage{graphicx}
\usepackage{url}
\usepackage{longtable}
\usepackage[portuguese]{babel}
\usepackage[latin1]{inputenc}
\urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
\bibliographystyle{splncs}
\begin{document}
\title{Descrição do dataset ``Adult Database''}
\author{Flávio Cruz\inst{1} \and João Azevedo\inst{1}}
\institute{Faculdade de Engenharia da Universidade do Porto\\
    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL
    \mails}
\maketitle

\begin{abstract}
Usado de forma extremamente frequente em testes experimentais relacionados com
data mining, a ``Adult Database'' reveste-se de algumas propriedades
interessantes que a tornam bastante útil para o teste da eficácia de algoritmos
de classificação. Este relatório pretende descrever a ``Adult Database'' de um
ponto de vista estatístico e tendo em conta o seu suporte para tarefas de data
mining.
\end{abstract}

\section{Introdução}
\label{sec:intro}

\section{Formato dos Dados}

O ``Adult Database'' consiste de 48842 entradas de informação de pessoas
recolhida durante um censos de 1994 realizado no estado da Califórnia, nos
Estados Unidos da América. Dos dados globais, foi extraída uma amostra
representativa de pessoas com idades entre os 16 e os 100 anos e um peso
superior a 1.

\section{Análise dos Dados}

\section{Trabalho Futuro}

\section{Conclusões}

\begin{thebibliography}{1}
\end{thebibliography}

\end{document}

### 2009-11-10 stats.r
Disconnect as soon as possible.

@@ -5,6 +5,7 @@ library(car)
 ### connects and loads the table data
 con <- dbConnect(dbDriver("MySQL"), dbname="adult", user="root", password="root")
 data <- dbReadTable(con, "adult")
+dbDisconnect(con)
 
 ### replaces unknown with empty values for missing values processing
 data[data == "unknown"] <- NA
@@ -145,5 +146,3 @@ generate_plot(data$education_num, data$capital_gain, "plot_education_num_capital
 generate_plot(data$education_num, data$capital_loss, "plot_education_num_capital_loss.png", "Education and capital loss", xlab = "Education", ylab = "Capital loss", labeldist = 0)
 generate_plot(data$education_num, data$capital_gain - data$capital_loss, "plot_education_num_capital_result.png", "Education and capital result", xlab = "Education", ylab = "Capital result", labeldist = 0)
 generate_plot(data$plus_50, data$education_num, "plot_plus_50_education_num.png", "Plus 50K and education", xlab = ">50K", ylab = "Education", labeldist = 0)
-
-dbDisconnect(con)

### 2009-11-10 stats.r
Put common stuff in common.r

@@ -1,11 +1,9 @@
 ### loads the required libraries
 library(RMySQL)
 library(car)
+source('common.r')
 
-### connects and loads the table data
-con <- dbConnect(dbDriver("MySQL"), dbname="adult", user="root", password="root")
-data <- dbReadTable(con, "adult")
-dbDisconnect(con)
+data <- get_dataset("root")
 
 ### replaces unknown with empty values for missing values processing
 data[data == "unknown"] <- NA
@@ -24,19 +22,9 @@ data$plus_50        <- factor(data$plus_50)
 ### data without values from the us
 data_no_us          <- data[data$native_country != "united_states",]
 
-get_filename <- function (name)
-{
-  return(paste("R", "/", name, sep = ""))
-}
-
 proportion_table <- function (data, exclude = NULL)
 {
   return(prop.table(table(data, exclude = exclude)) * 100.0)
-}
-
-set_png_output <- function (filename)
-{
-  png(file=get_filename(filename), bg = "white", width = 1000, height = 800)
 }
 
 ### generates graphics

### 2009-11-10 report/paper.tex
fixing some stuff on the report.
added data description to report.

@@
 \section{Introdução}
-\label{sec:intro}
 
@@
 O ``Adult Database'' consiste de 48842 entradas de informação de pessoas
-recolhida durante um censos de 1994 realizado no estado da Califórnia, nos
-Estados Unidos da América. Dos dados globais, foi extraída uma amostra
-representativa de pessoas com idades entre os 16 e os 100 anos e um peso
-superior a 1.
+recolhida durante um censos de 1994 realizado nos Estados Unidos da América. Dos
+dados globais, foi extraída uma amostra representativa de pessoas com idades
+entre os 16 e os 100 anos e um peso superior a 1.
+
+Cada entrada é composta por um conjunto de 14 atributos e por uma classe,
+identificando-a como uma pessoa que ganha mais ou menos de 50K. Os atributos e
+respectivos tipos são:
+
+\begin{itemize}
+  \item{\textbf{Age}: idade do indíviduo, sendo um tipo de dados contínuo.}
+  \item{\textbf{Workclass}: Classe operária do indivíduo, sendo um tipo de dados
+        categórico: [[Private], [Self-emp-not-inc], [Self-emp-inc], 
+        [Federal-gov], [Local-gov], [State-gov], [Without-pay], [Never-worked]]}
+  \item{\textbf{FNLWGT}: Medida demográfico do indíviduo. Sendo um tipo de dados
+        contínuo, deve ser próximo entre indivíduos com as mesmas
+        características sócio-económicas.}
+  \item{\textbf{Education}: Nível educacional do indivíduo, sendo um tipo de
+        dados ordinal: [1. [Preschool], 2. [1st-4th], 3. [5th-6th], 
+        4. [7th-8th], 5. [9th < 10th], 6. [11th < 12th], 7. [HS-grad], 
+        8. [Prof-school], 9. [Assoc-acdm], 10. [Assoc-voc], 11. [Some-college],
+        12. [Bachelors], 13. [Masters], 14. [Doctorate]]}
+  \item{\textbf{Education-num}: Representação numérica do atributo
+        \textbf{Education}, sendo um tipo de dados contínuo.}
+  \item{\textbf{Marital-status}: Estado civil do indivíduo, sendo um tipo de
+        dados categórico: [[Married-civ-spouse], [Divorced], [Never-married],
+        [Separated], [Widowed], [Married-spouse-absent], [Married-AF-spouse]]}
+  \item{\textbf{Occupation}: Área de ocupação profissional do indivíduo, sendo
+        um tipo de dados categórico: [[Tech-support], [Craft-repair],
+        [Other-service], [Sales], [Exec-managerial], [Prof-specialty], 
+        [Handlers-cleaners], [Machine-op-inspct], [Adm-clerical], 
+        [Farming-fishing], [Transport-moving], [Priv-house-serv], 
+        [Protective-serv], [Armed-Forces]]}
+  \item{\textbf{Relationship}: Tipo de relação familiar do indivíduo, sendo um
+        tipo de dados categórico: [[Wife], [Own-child], [Husband], 
+        [Not-in-family], [Other-relative], [Unmarried]]}
+  \item{\textbf{Race}: Raça do indivíduo, sendo um tipo de dados categórico: 
+      [[White], [Asian-Pac-Islander], [Amer-Indian-Eskimo], [Other], [Black]]}
+  \item{\textbf{Sex}: Sexo do indivíduo, sendo um tipo de dados categórico:
+      [[Male], [Female]]}
+  \item{\textbf{Capital-gain}: Ganho de capital do indivíduo, sendo um tipo de
+        dados contínuo}
+  \item{\textbf{Capital-loss}: Perda de capital do indivíduo, sendo um tipo de
+        dados contínuo}
+  \item{\textbf{Hours-per-week}: Horas de trabalho por semana do indivíduo,
+        sendo um tipo de dados contínuo}
+  \item{\textbf{Native-country}: Naturalidade do indivíduo, sendo um tipo de
+        dados categórico.}
+\end{itemize}
 
 \section{Análise dos Dados}
-
 \section{Trabalho Futuro}

### 2009-11-11 report/paper.tex
More sections.
added preliminary data analysis:

@@
 \usepackage[latin1]{inputenc}
+\usepackage{subfigure}
 \urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
@@
 \title{Descrição do dataset ``Adult Database''}
-\author{Flávio Cruz\inst{1} \and João Azevedo\inst{1}}
+\author{Flávio Cruz \and João Azevedo}
 \institute{Faculdade de Engenharia da Universidade do Porto\\
@@
 
+\section{Identificação do problema}
+
 \section{Análise dos Dados}
+
+A tabela \ref{tab:summary} apresenta um sumário da distribuição de valores no
+dataset estudado. Numa primeira observação da distribuição de valores, pode-se 
+constatar que são poucas as classes com \textit{missing vales} nos dados. Por
+outro lado, nalguns atributos, alguns valores têm uma frequência bastante
+superior, como são exemplo a naturalidade 'Estados Unidos', a classe operária
+'Private' e a raça 'White'. Nas horas de trabalho semanais, verifica-se a
+existência de bastantes \textit{outliers}, com valores pouco realistas,
+nomeadamente próximos das 100 horas ou menores que 10 em indivíduos com
+profissões definidas, como pode ser visto na figura \ref{fig:box_hours}.
+Variáveis como o ganho e perda capital apresentam poucos valores superiores a 0,
+sendo atributos bastante específicos de determinadas classes de indivíduos. O
+sexo também apresenta alguma discrepância na amostra, sendo o número de
+indivíduos homens mais do dobro das mulheres.
+
+\begin{table}
+\centering
+\subtable[Age]{
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 17 \\
+       \hline
+       1st Qu. & 28 \\
+       \hline
+       Median & 37 \\
+       \hline
+       Mean & 38.58 \\
+       \hline
+       3rd Qu. & 48 \\
+       \hline
+       Max & 90 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Workclass]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       private & 22696 \\
+       \hline
+       self\_emp\_not\_inc & 2541 \\
+       \hline
+       local\_gov & 2093 \\
+       \hline
+       state\_gov & 1298 \\
+       \hline
+       self\_emp\_inc & 1116 \\
+       \hline
+       (Other) & 981 \\
+       \hline
+       NA's & 1836 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Education]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       hs\_grad & 10501 \\
+       \hline
+       some\_college & 7291 \\
+       \hline
+       bachelors & 5535 \\
+       \hline
+       masters & 1723 \\
+       \hline
+       assoc\_voc & 1382 \\
+       \hline
+       11th & 1175 \\
+       \hline
+       (Other) & 5134 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Marital Status]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       divorced & 4443 \\
+       \hline
+       married\_af\_spouse & 23 \\
+       \hline
+       married\_civ\_spouse & 14976 \\
+       \hline
+       married\_spouse\_absent & 418 \\
+       \hline
+       never\_married & 10683 \\
+       \hline
+       separated & 1025 \\
+       \hline
+       widowed & 993 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Occupation]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       prof\_specialty & 4140 \\
+       \hline
+       craft\_repair & 4099 \\
+       \hline
+       exec\_managerial & 4066 \\
+       \hline
+       adm\_clerical & 3770 \\
+       \hline
+       sales & 3650 \\
+       \hline
+       (Other) & 10993 \\
+       \hline
+       NA's & 1843 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Relatonship]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       husband & 13193 \\
+       \hline
+       not\_in\_family & 8305 \\
+       \hline
+       other\_relative & 981 \\
+       \hline
+       own\_child & 5068 \\
+       \hline
+       unmarried & 3446 \\
+       \hline
+       wife & 1568 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Race]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       amer\_indian\_eskimo & 311 \\
+       \hline
+       asian\_pac\_islander & 1039 \\
+       \hline
+       black & 3124 \\
+       \hline
+       other & 271 \\
+       \hline
+       white & 27816 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Sex]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       female & 10771 \\
+       \hline
+       male & 21790 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Capital Gain]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 0 \\
+       \hline
+       1st Qu. & 0 \\
+       \hline
+       Median & 0 \\
+       \hline
+       Mean & 1078 \\
+       \hline
+       3rd Qu. & 0 \\
+       \hline
+       Max. & 99999 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Capital Loss]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 0 \\
+       \hline
+       1st Qu. & 0 \\
+       \hline
+       Median & 0 \\
+       \hline
+       Mean & 87.3 \\
+       \hline
+       3rd Qu. & 0 \\
+       \hline
+       Max. & 4356 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Hours Per Week]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 1 \\
+       \hline
+       1st Qu. & 40 \\
+       \hline
+       Median & 40 \\
+       \hline
+       Mean & 40.44 \\
+       \hline
+       3rd Qu. & 45 \\
+       \hline
+       Max. & 99 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Native Country]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       United States & 29170 \\
+       \hline
+       Mexico & 643 \\
+       \hline
+       Philippines & 198 \\
+       \hline
+       Germany & 137 \\
+       \hline
+       Canada & 121 \\
+       \hline
+       (Other) & 1709 \\
+       \hline
+       NA's & 583 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Plus 50]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       No & 24720 \\
+       \hline
+       Yes & 7841 \\
+       \hline
+       \end{tabular}
+}
+\caption{Sumário da distribuição de valores no dataset}
+\label{tab:summary}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=11cm]{boxplot_hours_per_week.png}
+\caption{Boxplot da variável correspondente às horas de trabalho semanais}
+\label{fig:box_hours}
+\end{figure} %}
+
+\subsection{Regras de associacao}
+
+\subsection{Clusters}
+
 \section{Trabalho Futuro}

### 2009-11-11 report/paper.tex
added data description.

@@
 \label{fig:box_hours}
-\end{figure} %}
+\end{figure}
+
+Importa também verificar algumas correlações entre variáveis do dataset. Para
+isso, testamos algumas situações mais óbvias, como a influência que a idade,
+horas de trabalho semanais e nível educacional influenciam o ganho do indivíduo.
+As figuras \ref{fig:age_plus50}, \ref{fig:education_plus50} e 
+\ref{fig:hpw_plus50} mostram que existe uma tendência para que pessoas mais
+velhas, pessoas com um nível educacional mais elevado e pessoas que trabalham 
+mais horas por semana ganhem mais.
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{plot_plus_50_age.png}
+\caption{Boxplot da idade dos indivíduos, filtrado pelo ganho salarial}
+\label{fig:age_plus50}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{plot_plus_50_education_num.png}
+\caption{Boxplot do nível educacional dos indivíduos, filtrado pelo ganho salarial}
+\label{fig:education_plus50}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{plot_plus_50_hpw.png}
+\caption{Boxplot das horas de trabalho semanais, fitrado pelo ganho salarial}
+\label{fig:hpw_plus50}
+\end{figure}
 

### 2009-11-11 report/paper.tex
report translated to english.

--- 2009-11-11 report/paper.tex
@@
 \usepackage{longtable}
-\usepackage[portuguese]{babel}
 \usepackage[latin1]{inputenc}
 \usepackage{subfigure}
+\usepackage{a4wide}
 \urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
@@
 \begin{document}
-\title{Descrição do dataset ``Adult Database''}
+\title{Progress Report on the ``Adult Database'' analysis}
 \author{Flávio Cruz \and João Azevedo}
@@
 \begin{abstract}
-Usado de forma extremamente frequente em testes experimentais relacionados com
-data mining, a ``Adult Database'' reveste-se de algumas propriedades
-interessantes que a tornam bastante útil para o teste da eficácia de algoritmos
-de classificação. Este relatório pretende descrever a ``Adult Database'' de um
-ponto de vista estatístico e tendo em conta o seu suporte para tarefas de data
-mining.
+Highly frequent in experimental tests related to data mining, the Adult Database
+has some interesting properties which make it quite useful to test the 
+efficiency of classification algorithms. This report aims to describe the Adult
+Database from a statistical perspective and taking in account its support for
+data mining tasks.
 \end{abstract}
 
-\section{Introdução}
+\section{Introduction}
 
-\section{Formato dos Dados}
+\section{Data Format}
 
-O ``Adult Database'' consiste de 48842 entradas de informação de pessoas
-recolhida durante um censos de 1994 realizado nos Estados Unidos da América. Dos
-dados globais, foi extraída uma amostra representativa de pessoas com idades
-entre os 16 e os 100 anos e um peso superior a 1.
+The ``Adult Database'' consists of 48842 entries of individual information of 
+people gathered in a 1994 census on the United States of America. From the
+global collection of data, a representative sample was collected from people
+with ages between 16 and 100 and a final weight above 1.
 
-Cada entrada é composta por um conjunto de 14 atributos e por uma classe,
-identificando-a como uma pessoa que ganha mais ou menos de 50K. Os atributos e
-respectivos tipos são:
+Each entry is composed by a set of 14 attributes and a class, identifying
+persons who gain more or less than 50K. The attributes and types are:
 
 \begin{itemize}
-  \item{\textbf{Age}: idade do indíviduo, sendo um tipo de dados contínuo.}
-  \item{\textbf{Workclass}: Classe operária do indivíduo, sendo um tipo de dados
-        categórico: [[Private], [Self-emp-not-inc], [Self-emp-inc], 
-        [Federal-gov], [Local-gov], [State-gov], [Without-pay], [Never-worked]]}
-  \item{\textbf{FNLWGT}: Medida demográfico do indíviduo. Sendo um tipo de dados
-        contínuo, deve ser próximo entre indivíduos com as mesmas
-        características sócio-económicas.}
-  \item{\textbf{Education}: Nível educacional do indivíduo, sendo um tipo de
-        dados ordinal: [1. [Preschool], 2. [1st-4th], 3. [5th-6th], 
-        4. [7th-8th], 5. [9th < 10th], 6. [11th < 12th], 7. [HS-grad], 
-        8. [Prof-school], 9. [Assoc-acdm], 10. [Assoc-voc], 11. [Some-college],
-        12. [Bachelors], 13. [Masters], 14. [Doctorate]]}
-  \item{\textbf{Education-num}: Representação numérica do atributo
-        \textbf{Education}, sendo um tipo de dados contínuo.}
-  \item{\textbf{Marital-status}: Estado civil do indivíduo, sendo um tipo de
-        dados categórico: [[Married-civ-spouse], [Divorced], [Never-married],
-        [Separated], [Widowed], [Married-spouse-absent], [Married-AF-spouse]]}
-  \item{\textbf{Occupation}: Área de ocupação profissional do indivíduo, sendo
-        um tipo de dados categórico: [[Tech-support], [Craft-repair],
+  \item{\textbf{Age}: continuous.}
+  \item{\textbf{Workclass}: categorical: [[Private], [Self-emp-not-inc], 
+        [Self-emp-inc], [Federal-gov], [Local-gov], [State-gov], [Without-pay], 
+        [Never-worked]]}
+  \item{\textbf{FNLWGT}: continuous. Represents the final weight. People with 
+        the same demographical characteristics should have the same weight.}
+  \item{\textbf{Education}: ordinal: [1. [Preschool], 2. [1st-4th], 3. 
+        [5th-6th], 4. [7th-8th], 5. [9th < 10th], 6. [11th < 12th], 7. 
+        [HS-grad], 8. [Prof-school], 9. [Assoc-acdm], 10. [Assoc-voc], 11.
+        [Some-college], 12. [Bachelors], 13. [Masters], 14. [Doctorate]]}
+  \item{\textbf{Education-num}: continuous. Is a continuous representation of 
+        the \textbf{Education} attribute.}
+  \item{\textbf{Marital-status}: categorical: [[Married-civ-spouse], [Divorced],
+        [Never-married], [Separated], [Widowed], [Married-spouse-absent],
+        [Married-AF-spouse]]}
+  \item{\textbf{Occupation}: categorical: [[Tech-support], [Craft-repair],
         [Other-service], [Sales], [Exec-managerial], [Prof-specialty], 
@@
         [Protective-serv], [Armed-Forces]]}
-  \item{\textbf{Relationship}: Tipo de relação familiar do indivíduo, sendo um
-        tipo de dados categórico: [[Wife], [Own-child], [Husband], 
+  \item{\textbf{Relationship}: categorical: [[Wife], [Own-child], [Husband], 
         [Not-in-family], [Other-relative], [Unmarried]]}
-  \item{\textbf{Race}: Raça do indivíduo, sendo um tipo de dados categórico: 
-        [[White], [Asian-Pac-Islander], [Amer-Indian-Eskimo], [Other], [Black]]}
-  \item{\textbf{Sex}: Sexo do indivíduo, sendo um tipo de dados categórico:
-        [[Male], [Female]]}
-  \item{\textbf{Capital-gain}: Ganho de capital do indivíduo, sendo um tipo de
-        dados contínuo}
-  \item{\textbf{Capital-loss}: Perda de capital do indivíduo, sendo um tipo de
-        dados contínuo}
-  \item{\textbf{Hours-per-week}: Horas de trabalho por semana do indivíduo,
-        sendo um tipo de dados contínuo}
-  \item{\textbf{Native-country}: Naturalidade do indivíduo, sendo um tipo de
-        dados categórico.}
+  \item{\textbf{Race}: categorial: [[White], [Asian-Pac-Islander], 
+        [Amer-Indian-Eskimo], [Other], [Black]]}
+  \item{\textbf{Sex}: categorical: [[Male], [Female]]}
+  \item{\textbf{Capital-gain}: continuous.}
+  \item{\textbf{Capital-loss}: continuous.}
+  \item{\textbf{Hours-per-week}: continuous.}
+  \item{\textbf{Native-country}: categorical.}
 \end{itemize}
 
-\section{Identificação do problema}
+\section{Problem Identification}
 
-\section{Análise dos Dados}
+\section{Data Analysis}
 
-A tabela \ref{tab:summary} apresenta um sumário da distribuição de valores no
-dataset estudado. Numa primeira observação da distribuição de valores, pode-se 
-constatar que são poucas as classes com \textit{missing vales} nos dados. Por
-outro lado, nalguns atributos, alguns valores têm uma frequência bastante
-superior, como são exemplo a naturalidade 'Estados Unidos', a classe operária
-'Private' e a raça 'White'. Nas horas de trabalho semanais, verifica-se a
-existência de bastantes \textit{outliers}, com valores pouco realistas,
-nomeadamente próximos das 100 horas ou menores que 10 em indivíduos com
-profissões definidas, como pode ser visto na figura \ref{fig:box_hours}.
-Variáveis como o ganho e perda capital apresentam poucos valores superiores a 0,
-sendo atributos bastante específicos de determinadas classes de indivíduos. O
-sexo também apresenta alguma discrepância na amostra, sendo o número de
-indivíduos homens mais do dobro das mulheres.
+The table \ref{tab:summary} summarizes the distribution of values on the studied
+dataset. From a first observation, one can see that the amount of missing values
+is low. On the other side, in some of the attributes, some classes have a
+much higher frequency than others, as the 'united-states' native country, the
+'private' workclass and the 'white' race. On the hours of work per week
+attribute, there is a great amount of outliers, with some unrealistic
+values, namely with values close to the 100 hours or less than 10 in individuals
+with a defined occupation, as can be seen on figure \ref{fig:box_hours}.
+Variables such has the capital gain and capital loss have few values above 0,
+being quite specific to certain classes of individuals. The sex is also
+unbalanced in the sample data, as the amount of men is more than twice the
+amount of women.
+
+It is also important to verify some correlations between variables. To that
+matter, we've tested some apparent situations, such as the influence of the age,
+hours of work per week and educational level on the amount gained by an
+individual. The figures \ref{fig:age_plus50}, \ref{fig:education_plus50} and
+\ref{fig:hpw_plus50} show that there is a tendency for older people, people with
+a higher educational level and people who work a great amount of hours per week
+to gain a larger amount of money.
+
+\subsection{Association Rules}
+
+\subsection{Clusters}
+
+\section{Future Work}
+
+\section{Conclusions}
+
+\begin{thebibliography}{1}
+\end{thebibliography}
+
+\clearpage
+
+\appendix
+
+\section{Tables}
 
@@
 }
-\caption{Sumário da distribuição de valores no dataset}
+\caption{Summary of the value distribution in the database}
 \label{tab:summary}
@@
 
+\clearpage
+
+\section{Figures}
+
 \begin{figure}
 \centering
-\includegraphics[width=11cm]{boxplot_hours_per_week.png}
-\caption{Boxplot da variável correspondente às horas de trabalho semanais}
+\includegraphics[width=10cm]{boxplot_hours_per_week.png}
+\caption{Boxplot of the hours per week variable}
 \label{fig:box_hours}
-\end{figure} %}
+\end{figure}
 
-\subsection{Regras de associacao}
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{plot_plus_50_age.png}
+\caption{Boxplot of the individuals' age, for each class of gain}
+\label{fig:age_plus50}
+\end{figure}
 
-\subsection{Clusters}
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{plot_plus_50_education_num.png}
+\caption{Boxplot of the individuals' education level, for each class of gain}
+\label{fig:education_plus50}
+\end{figure}
 
-\section{Trabalho Futuro}
-
-\section{Conclusões}
-
-\begin{thebibliography}{1}
-\end{thebibliography}
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{plot_plus_50_hpw.png}
+\caption{Boxplot of the individuals' hours of work per week, for each class of 
+gain}
+\label{fig:hpw_plus50}
+\end{figure}
 

### 2009-11-11 report/paper.tex
fixing typos.

@@
 efficiency of classification algorithms. This report aims to describe the Adult
-Database from a statistical perspective and taking in account its support for
+Database from a statistical perspective and taking into account its support for
 data mining tasks.
@@
 is low. On the other side, in some of the attributes, some classes have a
-much higher frequency than others, as the 'united-states' native country, the
-'private' workclass and the 'white' race. On the hours of work per week
+much higher frequency than others, such as the 'united-states' native country, 
+the 'private' workclass and the 'white' race. On the hours of work per week
 attribute, there is a great amount of outliers, with some unrealistic
-values, namely with values close to the 100 hours or less than 10 in individuals
-with a defined occupation, as can be seen on figure \ref{fig:box_hours}.
-Variables such has the capital gain and capital loss have few values above 0,
-being quite specific to certain classes of individuals. The sex is also
-unbalanced in the sample data, as the amount of men is more than twice the
-amount of women.
+values, namely with amounts close to the 100 hours or less than 10 in 
+individuals with a defined occupation, as can be seen on figure 
+\ref{fig:box_hours}. Variables such has the capital gain and capital loss have
+few values above 0, being quite specific to certain classes of individuals. The
+sex is also unbalanced in the sample data, as the amount of men is more than
+twice the amount of women.
 

### 2009-11-11 report/paper.tex
Add problem identification.

--- 2009-11-11 report/paper.tex
@@
 
-Each entry is composed by a set of 14 attributes and a class, identifying
-persons who gain more or less than 50K. The attributes and types are:
+Each entry is composed by a set of 14 attributes and two classes, identifying
+persons who gain more or less than 50K dollars per year. The attributes and types are:
 
@@
 
+Deducing from the previous section, this is a classification problem. Given an adult
+with the mentioned attributes we want to classify him into one of the following groups:
+one gains more than 50000 dollars per year, the other group does not.
+
+One important objective in classifying things is to understand why the given object
+is put under one class and not in another, that is, we want to understand what social characteristics
+allows one person to gain more than someone else, if the sex is important, the education level,
+the race and what factors are less important.
+
+Given a specific set of persons from a certain society and from a certain time we could then
+learn what contributes to a person's social standing and, as we all know, money is an important
+metric in that department. We could, for example, learn if the race affects the persons success and if
+that society discriminates certain races.
+
+Understanding what affects the money gained through an year can gives a lot of information
+concerning society's behaviors.
+
 \section{Data Analysis}
@@
 
-\section{Future Work}
+\section{Work plan}
 

### 2009-11-11 report/paper.tex
More dataset information.
Dataset URL.

@@
 with ages between 16 and 100 and a final weight above 1.
+
+The data set is divided in two groups: the train set with 32561 records and the test set with 16281 entries.
 
@@
 \end{itemize}
+
+There are 3620 records with missing values. The attributes that contain missing values are: \textbf{Workclass}, \textbf{Occupation} and \textbf{Native-country}.
+
+This data set is available in the following URL: \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}.
 

### 2009-11-11 report/paper.tex
added preliminary cluster analysis.

@@
 \subsection{Clusters}
+
+Another method to analyze the dataset is based on the generation of clusters. 
+Clusters are sets of data in which objects in the same cluster are similar, and
+objects in different clusters are dissimilar. The first attempt at clustering
+used only three variables: age, education\_num and hours\_per\_week, the same
+variables we saw that had a greater impact on the amount gained by the
+individual. The data was normalized using the Z-Transformation method and the 
+K-Means method was used to obtain the final clusters, with a k of 6. Figures 
+\ref{fig:cluster_1_age_education} and \ref{fig:cluster_1_education_hpw} show
+scatter plots of the variables, colored by clusters. One thing one is able to
+note is that people with younger age belong to cluster \#2. In terms of
+education, people with higher education belong to cluster \#0, while people with
+lower education belong to cluster \#1. In terms of hours of work per week, 
+distribution is rather similar, but cluster \#2 and \#3 have a greater amount of
+lower and higher values, respectively. Figure \ref{fig:cluster_1_cluster_plus50}
+shows how clusters relate to the class values. Although the distribution is 
+balanced, clusters \#1 and \#2 have a greater amount of people gaining below 
+50K. Notice, as stated before, that this were the cluster having people with
+lower education and younger age, respective.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{cluster_1_age_education.png}
+\caption{Scatter plot of age and education}
+\label{fig:cluster_1_age_education}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{cluster_1_education_hpw.png}
+\caption{Scatter plot of education and hours of work per week}
+\label{fig:cluster_1_education_hpw}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{cluster_1_cluster_plus50.png}
+\caption{Scatter plot of cluster and class values of amount gained}
+\label{fig:cluster_1_cluster_plus50}
+\end{figure}
+
 \end{document}

### 2009-11-12 report/paper.tex
added cluster analysis.

@@
 \institute{Faculdade de Engenharia da Universidade do Porto\\
-    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL
+    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL\\
     \mails}
@@
 K-Means method was used to obtain the final clusters, with a k of 6. Figures 
-\ref{fig:cluster_1_age_education} and \ref{fig:cluster_1_education_hpw} show
+\ref{fig:cluster_1_age_education}, \ref{fig:cluster_1_education_hpw} and \ref{fig:cluster_1_age_hpw} show
 scatter plots of the variables, colored by clusters. One thing one is able to
@@
 balanced, clusters \#1 and \#2 have a greater amount of people gaining below 
-50K. Notice, as stated before, that this were the cluster having people with
-lower education and younger age, respective.
+50000 dollars. Notice, as stated before, that this were the clusters having 
+people with lower education and younger age, respectively.
+
+Clustering with the remaining attributes leads to results more difficult to
+analyze. The same process was applied, using the Z-Transformation method for
+normalization and the K-Means algorithm with a k of 6, but this time with all
+the attributes. The distribution of values is more balanced, although a very
+high value of capital gain, very unusual on the data set, leads to the creation
+of a very well identified cluster, as can be seen on table \ref{tab:centroids}.
+Particular values of race, workclass and native country also identify clusters
+\#2, \#3 and \#2, respectively. Notice that this attributes were the ones
+refered to before as having a given value more frequent than others, thus
+leading to this particular composition of clusters.
 
@@
 
+\begin{table}
+\centering
+\begin{tabular}{ | l | l | l | l | l | l | l | }
+\hline
+\textbf{Attribute} & \textbf{Centroid \#0} & \textbf{Centroid \#1} & \textbf{Centroid \#2} & \textbf{Centroid \#3} & \textbf{Centroid \#4} & \textbf{Centroid \#5} \\
+\hline
+age	& -0.874 & 0.692 & -0.166 & 0.246 & 0.570 & 0.190 \\
+\hline
+workclass & -0.272 & -0.063 & -0.036 & 2.208 & 0.455 & -0.364 \\
+\hline
+education\_num & -0.101 & -0.172 & -0.177 & 0.0309 & 1.103 & 0.102 \\
+\hline
+marital\_status & -0.679 & 1.693 & -0.083 & -0.150 & -0.007 & -0.108 \\
+\hline
+occupation & -0.305 & -0.303 & 0.091 & 0.952 & -0.434 & 0.010 \\
+\hline
+relationship & 0.703 & 0.526 & 0.189 & -0.053 & -0.351 & -0.507 \\
+\hline
+race & -0.086 & -0.020 & 2.630 & -0.147 & -0.002 & -0.230 \\
+\hline
+sex & 0.662 & 1.199 & 0.019 & -0.161 & -0.409 & -0.635 \\
+\hline
+capital\_gain & -0.116 & -0.091 & -0.086 & -0.029 & 13.394 & -0.040 \\
+\hline
+capital\_loss & -0.115 & -0.079 & -0.021 & 0.025 & -0.217 & 0.079 \\
+\hline
+hours\_per\_week & -0.523 & -0.231 & -0.031 &-0.071 & 0.758 & 0.339 \\
+\hline
+native\_country & -0.186 & -0.154 & 2.951 & -0.181 & -0.075 & -0.171 \\
+\hline
+\end{tabular}
+\caption{Centroid table for the results of the application of the K-Means algorithm with a k of 6}
+\label{tab:centroids}
+\end{table}
+
 \clearpage
@@
 \centering
+\includegraphics[width=10cm]{cluster_1_age_hpw.png}
+\caption{Scatter plot of age and hours of work per week}
+\label{fig:cluster_1_age_hpw}
+\end{figure}
+
+\begin{figure}
+\centering
 \includegraphics[width=10cm]{cluster_1_cluster_plus50.png}

### 2009-11-12 report/paper.tex
added introduction.

@@
 \begin{document}
-\title{Progress Report on the ``Adult Database'' analysis}
+\title{Progress Report on the ``Adult Database'' Analysis}
 \author{Flávio Cruz \and João Azevedo}
@@
 \section{Introduction}
+
+Classification in data mining is the activity of identifying a given object as 
+belonging to a pre-defined class. The ultimate goal is to learn a concept. This
+learning process is based on past observations, where the dataset performs a
+main role. This report gives a first overview on the ``Adult Database'' dataset,
+describing its format and identifying the problem it poses. Some data analysis
+is then done, in order to obtain some characteristics on the global set of
+values. Finally, the future work plan is presented and some conclusions on the
+subject are taken.
 

### 2009-11-12 report/paper.tex
added references and citations.

@@
 Classification in data mining is the activity of identifying a given object as 
-belonging to a pre-defined class. The ultimate goal is to learn a concept. This
+belonging to a pre-defined class \cite{1}. The ultimate goal is to learn a concept. This
 learning process is based on past observations, where the dataset performs a
@@
 One important objective in classifying things is to understand why the given object
-is put under one class and not in another, that is, we want to understand what social characteristics
+is put under one class and not in another \cite{2}\cite{3}, that is, we want to understand what social characteristics
 allows one person to gain more than someone else, if the sex is important, the education level,
@@
 individual. The data was normalized using the Z-Transformation method and the 
-K-Means method was used to obtain the final clusters, with a k of 6. Figures 
+K-Means method was used to obtain the final clusters, with a k of 6, using RapidMiner \cite{4}. Figures 
 \ref{fig:cluster_1_age_education}, \ref{fig:cluster_1_education_hpw} and \ref{fig:cluster_1_age_hpw} show
@@
 \begin{thebibliography}{1}
+\bibitem{1}
+Mitchel, T.; Machine Learning; McGraw-Hill, 1997
+\bibitem{2}
+Han, J., Kamber, M., Pei, J.; Data Mining: Concepts and Techniques, Second Edition; Morgan Kaufmann, 2005
+\bibitem{3}
+H. Witten, I., Frank, E.; Data Mining: Practical Machine Learning Tools and Techniques, Second Edition; Morgan Kaufmann, 2005
+\bibitem{4}
+RapidMiner; \url{http://www.rapidminer.com/}, November 2009
+\bibitem{5}
+Ross Quinlan, J.; C4.5: Programs for Machine Learning; Morgan Kaufmann, 1992
 \end{thebibliography}

### 2009-11-12 report/paper.tex
Association rules and work plan.

--- 2009-11-12 report/paper.tex
@@
 \usepackage{graphicx}
-\usepackage{url}
 \usepackage{longtable}
@@
 \usepackage{a4wide}
+\usepackage{listings}
+\usepackage[hyphens]{url}
 \urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
@@
 Given a specific set of persons from a certain society and from a certain time we could then
-learn what contributes to a person's social standing and, as we all know, money is an important
+learn what contributes to a person's social standing and, as we all know, income is an important
 metric in that department. We could, for example, learn if the race affects the persons success and if
-that society discriminates certain races.
+that society discriminates against certain races.
 
-Understanding what affects the money gained through an year can gives a lot of information
-concerning society's behaviors.
+Understanding what affects the income can give a lot of information concerning society's behaviors.
 
@@
 \subsection{Association Rules}
+
+Using association rules one can find frequent patterns and associations from a
+set of transactions. These rules can easily be used to make data analysis and
+to understand casual structures.
+
+To build a set of association rules we used the apriori algorithm that is featured
+in the R arules package. Before running the algorithm we made some data preprocessing
+for continuous attributes that are needed to build better and more general rules.
+
+The following attributes were transformed:
+
+\begin{itemize}
+  \item {\textbf{Age}: turned into categories. 15-25 (young), 25-45 (middle\_aged), 45-65 (senior), $>$ 65 (old).}
+  \item {\textbf{Hours-per-week}: turned into categories. 0-25 (part\_time), 25-40 (full\_time), 40-60 (overtime), $>$60 (workaholic).}
+  \item {\textbf{Capital-gain}: turned into categories. $<=$ 0 (none), $<$ median value (low), $>$ median value (high).}
+  \item {\textbf{Capital-loss}: turned into categories. $<=$ 0 (none), $<$ median value (low), $>$ median value (high).}
+\end{itemize}
+
+Next, we plotted the relative frequency of items in the transaction database
+with support greater than 10\%. The results are shown in figure \ref{fig:arules_item_frequency}.
+We can learn various things from that plot: the majority of people are from the united states,
+the white race is predominant, males outnumber females and that more people work in full time.
+
+Finally we run the apriori algorithm using the test data set to generate rules
+with support greater than 2\% and at least 60\% of confidence. The algorithm built
+50487 rules.
+
+We then built two sets of rules from the original set: one with the consequent component
+with plus\_50=0 (the person gains less than 50K a year) and the other with plus\_50=1
+(the person gains more than 50K a year).
+
+Each rule set was sorted by the confidence level, from more confident to less confident.
+
+The first 5 rules from the plus\_50=0 set are shown in listing \ref{arules_small_income} and
+are related to small incomes. We can easily make some observations
+about those rules: young people with a part-time are probably low income and having a child
+without being married is also not a very good signal.
+
+The second set, contrary to the ones above, shows us the top 5 rules for having large incomes.
+These rules are shown in listing \ref{arules_large_income}. A conclusion can readily be made:
+being a white husband with a high capital gain probably means we have a large income.
 
@@
 \section{Work plan}
+
+Given the problem at hand, we want to generate a classification mechanism
+that can tell if a specific person has a small or large income.
+
+For this we will use various kinds of classification algorithms, from classical
+decision trees to rule induction.
+
+We will try to compare the error rate among the various algorithms. We'll be using
+the train set to "train" the algorithm and then the test set to assess which algorithm
+is best suited to make income predictions.
+
+We'll also use some data-processing techniques to eliminate unknown values. It will
+also be useful to categorize continuous values for more generalized rules as we have
+done when generating association rules.
+
+The algorithms we will be using are: ID3, C4.5, CN2 and AQ. Depending on our progress
+we will experiment with neural networks and genetic algorithms, which are other
+mechanisms for building classifiers.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=20cm,angle=90]{item_frequency.png}
+\caption{Item frequencies of items with at least 10\% support.}
+\label{fig:arules_item_frequency}
+\end{figure}
+
+\clearpage
+\section{Listings}
+
+\begin{lstlisting}[language=r,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Association rules for small income.},label=arules_small_income]
+  lhs                               rhs            support confidence     lift
+1 {age=young,                                                                 
+   hours_per_week=part_time}     => {plus_50=0} 0.05730782          1 1.317193
+2 {age=young,                                                                 
+   occupation=unknown,                                                        
+   capital_gain=none}            => {plus_50=0} 0.02002396          1 1.317193
+3 {age=young,                                                                 
+   relationship=own_child,                                                    
+   hours_per_week=part_time}     => {plus_50=0} 0.04404042          1 1.317193
+4 {education=some_college,                                                    
+   relationship=own_child,                                                    
+   hours_per_week=part_time}     => {plus_50=0} 0.02069961          1 1.317193
+5 {marital_status=never_married,                                              
+   relationship=own_child,                                                    
+   hours_per_week=part_time}     => {plus_50=0} 0.04689659          1 1.317193
+\end{lstlisting}
+
+\begin{lstlisting}[language=r,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Association rules for large income.},label=arules_large_income]
+  lhs                                    rhs            support confidence     lift
+1 {relationship=husband,                                                           
+   race=white,                                                                     
+   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
+2 {marital_status=married_civ_spouse,                                              
+   relationship=husband,                                                           
+   race=white,                                                                     
+   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
+3 {relationship=husband,                                                           
+   race=white,                                                                     
+   sex=male,                                                                       
+   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
+4 {relationship=husband,                                                           
+   race=white,                                                                     
+   capital_gain=high,                                                              
+   capital_loss=none}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
+5 {relationship=husband,                                                           
+   capital_gain=high}                 => {plus_50=1} 0.02251159  0.9918809 4.118943
+\end{lstlisting}
+
 \end{document}

### 2009-11-12 report/paper.tex
added more references.

@@
 
-The algorithms we will be using are: ID3, C4.5, CN2 and AQ. Depending on our progress
+The algorithms we will be using are: ID3 \cite{1}, C4.5 \cite{5}, CN2 \cite{1} and AQ \cite{1}. Depending on our progress
 we will experiment with neural networks and genetic algorithms, which are other
@@
 \section{Conclusions}
+
+Taking into account the previous analysis of the data, we're now capable of
+applying a structured data mining process in order to address the classification
+problem posed by the dataset. The final report hopes to compare the performance
+and efficiency of different classification algorithms applied to this dataset.
+A tentative title for the paper to be delivered as the final report is 
+``A comparison of classification algorithms applied to the Adult Database''.
 

### 2009-11-12 report/paper.tex
adding more references.

@@
 To build a set of association rules we used the apriori algorithm that is featured
-in the R arules package. Before running the algorithm we made some data preprocessing
+in the R \cite{6} arules \cite{7} package. Before running the algorithm we made some data preprocessing
 for continuous attributes that are needed to build better and more general rules.
@@
 Ross Quinlan, J.; C4.5: Programs for Machine Learning; Morgan Kaufmann, 1992
+\bibitem{6}
+The R Project for Statistical Computing; http://www.r-project.org/, November 2009
+\bibitem{7}
+Hahsler, M., Gruen, B.; Hornik, K.; arules -- A Computational Environment for Mining Association Rules and Frequent Item Sets; Journal of Statistical Software 14/15, 2005
 \end{thebibliography}
@@
 \centering
-\includegraphics[width=10cm]{boxplot_hours_per_week.png}
+\includegraphics[width=9cm]{boxplot_hours_per_week.png}
 \caption{Boxplot of the hours per week variable}
@@
 \centering
-\includegraphics[width=10cm]{plot_plus_50_age.png}
+\includegraphics[width=8cm]{plot_plus_50_age.png}
 \caption{Boxplot of the individuals' age, for each class of gain}
@@
 \centering
-\includegraphics[width=10cm]{plot_plus_50_education_num.png}
+\includegraphics[width=9cm]{plot_plus_50_education_num.png}
 \caption{Boxplot of the individuals' education level, for each class of gain}
@@
 \centering
-\includegraphics[width=10cm]{plot_plus_50_hpw.png}
+\includegraphics[width=9cm]{plot_plus_50_hpw.png}
 \caption{Boxplot of the individuals' hours of work per week, for each class of 

### 2009-12-06 weka/C45.java
C4.5

@@ -1,0 +1,205 @@
+import weka.classifiers.trees.J48;
+import weka.classifiers.Evaluation;
+import weka.core.Attribute;
+import weka.core.Instance;
+import weka.core.Instances;
+import weka.experiment.InstanceQuery;
+import weka.filters.Filter;
+import weka.filters.unsupervised.attribute.ReplaceMissingValues;
+import weka.filters.unsupervised.attribute.MathExpression;
+import weka.filters.unsupervised.attribute.NumericToNominal;
+
+public class C45 {
+	
+	private static final int AGE_ATTRIBUTE = 0;
+	private static final int WORKCLASS_ATTRIBUTE = 1;
+	private static final int EDUCATION_ATTRIBUTE = 2;
+	private static final int MARITAL_STATUS_ATTRIBUTE = 3;
+	private static final int OCCUPATION_ATTRIBUTE = 4;
+	private static final int RELATIONSHIP_ATTRIBUTE = 5;
+	private static final int RACE_ATTRIBUTE = 6;
+	private static final int SEX_ATTRIBUTE = 7;
+	private static final int CAPITAL_GAIN_ATTRIBUTE = 8;
+	private static final int CAPITAL_LOSS_ATTRIBUTE = 9;
+	private static final int HOURS_PER_WEEK_ATTRIBUTE = 10;
+	private static final int NATIVE_COUNTRY_ATTRIBUTE = 11;
+	
+	private static String MAIN_SQL = "select age, workclass, education, " +
+	"marital_status, occupation, relationship, " +
+	"race, sex, capital_gain, capital_loss, " +
+	"hours_per_week, native_country, plus_50 from ";
+	private static String TRAIN_QUERY =  MAIN_SQL + "adult";
+	private static String TEST_QUERY = MAIN_SQL + "adult_test";
+	
+	private static Instances numericToNominal(Instances data, int []attributes) throws Exception {
+		NumericToNominal filter = new NumericToNominal();
+		filter.setAttributeIndicesArray(attributes);
+		filter.setInputFormat(data);
+		return Filter.useFilter(data, filter);
+	}
+	
+	private static void changeUnknown(Instances data, int attr)
+	{
+		for(int i = 0; i < data.numInstances(); ++i) {
+			Instance inst = data.instance(i);
+			if(inst.stringValue(attr).equals("unknown")) {
+				inst.setMissing(attr);
+			}
[…]

### 2009-12-06 generate_clauses.py
added clause generation script for ILP.

@@ -1,0 +1,375 @@
+#!/usr/bin/python
+
+# Script to generate .b, .f and .n files suitable for using with the ILP system
+# Aleph (http://www.comlab.ox.ac.uk/activities/machinelearning/Aleph/aleph.html)
+
+import sys
+
+if len(sys.argv) != 3:
+    print "generate_clauses.py <input> <output>"
+    sys.exit(1)
+
+def normalize_field(field):
+    return field.lower().replace('-', '_')
+
+def write_details(backf):
+    string_details = """
+:- modeh(1, more_than_50K(+person)).
+:- modeb(1, young(+person)).
+:- modeb(1, middle_aged(+person)).
+:- modeb(1, senior(+person)).
+:- modeb(1, old(+person)).
+:- modeb(1, work_private(+person)).
+:- modeb(1, work_self_emp_not_inc(+person)).
+:- modeb(1, work_self_emp_inc(+person)).
+:- modeb(1, work_federal_gov(+person)).
+:- modeb(1, work_local_gov(+person)).
+:- modeb(1, work_state_gov(+person)).
+:- modeb(1, work_without_pay(+person)).
+:- modeb(1, work_never_worked(+person)).
+:- modeb(1, education_bachelors(+person)).
+:- modeb(1, education_some_college(+person)).
+:- modeb(1, education_11th(+person)).
+:- modeb(1, education_hs_grad(+person)).
+:- modeb(1, education_prof_school(+person)).
+:- modeb(1, education_assoc_acdm(+person)).
+:- modeb(1, education_assoc_voc(+person)).
+:- modeb(1, education_9th(+person)).
+:- modeb(1, education_7th_8th(+person)).
+:- modeb(1, education_12th(+person)).
+:- modeb(1, education_masters(+person)).
+:- modeb(1, education_1st_4th(+person)).
+:- modeb(1, education_10th(+person)).
+:- modeb(1, education_doctorate(+person)).
+:- modeb(1, education_5th_6th(+person)).
+:- modeb(1, education_preschool(+person)).
+:- modeb(1, married_civ_spouse(+person)).
+:- modeb(1, divorced(+person)).
+:- modeb(1, never_married(+person)).
+:- modeb(1, separated(+person)).
+:- modeb(1, widowed(+person)).
+:- modeb(1, married_spouse_absent(+person)).
+:- modeb(1, married_af_spouse(+person)).
+:- modeb(1, occupation_tech_support(+person)).
+:- modeb(1, occupation_craft_repair(+person)).
[…]

### 2009-12-06 weka/C45.java
bayes networks using genetic search

@@ -1,170 +1,8 @@
-import weka.classifiers.trees.J48;
 import weka.classifiers.Evaluation;
-import weka.core.Attribute;
-import weka.core.Instance;
+import weka.classifiers.trees.J48;
 import weka.core.Instances;
-import weka.experiment.InstanceQuery;
-import weka.filters.Filter;
-import weka.filters.unsupervised.attribute.ReplaceMissingValues;
-import weka.filters.unsupervised.attribute.MathExpression;
-import weka.filters.unsupervised.attribute.NumericToNominal;
 
 public class C45 {
-	
-	private static final int AGE_ATTRIBUTE = 0;
-	private static final int WORKCLASS_ATTRIBUTE = 1;
-	private static final int EDUCATION_ATTRIBUTE = 2;
-	private static final int MARITAL_STATUS_ATTRIBUTE = 3;
-	private static final int OCCUPATION_ATTRIBUTE = 4;
-	private static final int RELATIONSHIP_ATTRIBUTE = 5;
-	private static final int RACE_ATTRIBUTE = 6;
-	private static final int SEX_ATTRIBUTE = 7;
-	private static final int CAPITAL_GAIN_ATTRIBUTE = 8;
-	private static final int CAPITAL_LOSS_ATTRIBUTE = 9;
-	private static final int HOURS_PER_WEEK_ATTRIBUTE = 10;
-	private static final int NATIVE_COUNTRY_ATTRIBUTE = 11;
-	
-	private static String MAIN_SQL = "select age, workclass, education, " +
-	"marital_status, occupation, relationship, " +
-	"race, sex, capital_gain, capital_loss, " +
-	"hours_per_week, native_country, plus_50 from ";
-	private static String TRAIN_QUERY =  MAIN_SQL + "adult";
-	private static String TEST_QUERY = MAIN_SQL + "adult_test";
-	
-	private static Instances numericToNominal(Instances data, int []attributes) throws Exception {
-		NumericToNominal filter = new NumericToNominal();
-		filter.setAttributeIndicesArray(attributes);
-		filter.setInputFormat(data);
-		return Filter.useFilter(data, filter);
-	}
-	
-	private static void changeUnknown(Instances data, int attr)
-	{
-		for(int i = 0; i < data.numInstances(); ++i) {
-			Instance inst = data.instance(i);
[…]

### 2009-12-06 weka/Genetic.java

@@ -1,0 +1,59 @@
+import weka.core.Instances;
+import weka.classifiers.Evaluation;
+import weka.classifiers.bayes.BayesNet;
+import weka.filters.Filter;
+import weka.filters.unsupervised.instance.RemovePercentage;
+
+public class Genetic {
+	public static void main(String[] args) throws Exception {
+		if(args.length != 2) {
+			System.err.println("usage: C45 <db user> <db password>");
+			System.exit(1);
+		}
+		
+		String user = args[0];
+		String password = args[1];
+	
+		
+		Instances train_data = Utils.getTrainSet(user, password);
+		/*System.out.println("====================================");
+		System.out.println("TRAIN DATA:");
+		System.out.println("====================================");
+		System.out.println(train_data.toSummaryString());
+		*/
+		Instances test_data = Utils.getTestSet(user, password);
+		/*System.out.println("====================================");
+		System.out.println("TEST DATA:");
+		System.out.println("====================================");
+		System.out.println(test_data.toSummaryString());
+*/
+		RemovePercentage rp = new RemovePercentage();
+		rp.setPercentage(99); // usar apenas 1% das instancias
+		rp.setInputFormat(train_data);
+		Instances new_data = Filter.useFilter(train_data, rp);
+		
+		BayesNet net = new BayesNet();
+		
+		String options = "" +
+			"-Q weka.classifiers.bayes.net.search.local.GeneticSearch -- -L 3 -A 3 -U 2 -C -M -R 1 " +
+			"-E weka.classifiers.bayes.net.estimate.SimpleEstimator -- -A 1.0";
+		System.out.println(options);
+		
+		net.setOptions(weka.core.Utils.splitOptions(options));
+		net.buildClassifier(new_data);
+		
+		System.out.println(net.toString());
+	
+		Evaluation train_eval = new Evaluation(train_data);
+		train_eval.evaluateModel(net, train_data);
+		System.out.println(train_eval.toSummaryString("\nTrain Results\n===============\n", false));
+	
+		Evaluation new_eval = new Evaluation(new_data);
+		new_eval.evaluateModel(net, new_data);
+		System.out.println(new_eval.toSummaryString("\nNew Results\n===============\n", false));
+	
+		Evaluation test_eval = new Evaluation(train_data);
+		test_eval.evaluateModel(net, test_data);
+		System.out.println(test_eval.toSummaryString("\nTest Results\n===============\n", false));
+	}
+}

### 2009-12-06 weka/Utils.java

@@ -1,0 +1,176 @@
+import weka.core.Attribute;
+import weka.core.Instance;
+import weka.core.Instances;
+import weka.experiment.InstanceQuery;
+import weka.filters.Filter;
+import weka.filters.unsupervised.attribute.MathExpression;
+import weka.filters.unsupervised.attribute.NumericToNominal;
+import weka.filters.unsupervised.attribute.ReplaceMissingValues;
+
+
+public class Utils {
+	private static final int AGE_ATTRIBUTE = 0;
+	private static final int WORKCLASS_ATTRIBUTE = 1;
+	private static final int EDUCATION_ATTRIBUTE = 2;
+	private static final int MARITAL_STATUS_ATTRIBUTE = 3;
+	private static final int OCCUPATION_ATTRIBUTE = 4;
+	private static final int RELATIONSHIP_ATTRIBUTE = 5;
+	private static final int RACE_ATTRIBUTE = 6;
+	private static final int SEX_ATTRIBUTE = 7;
+	private static final int CAPITAL_GAIN_ATTRIBUTE = 8;
+	private static final int CAPITAL_LOSS_ATTRIBUTE = 9;
+	private static final int HOURS_PER_WEEK_ATTRIBUTE = 10;
+	private static final int NATIVE_COUNTRY_ATTRIBUTE = 11;
+	
+	private static String MAIN_SQL = "select age, workclass, education, " +
+	"marital_status, occupation, relationship, " +
+	"race, sex, capital_gain, capital_loss, " +
+	"hours_per_week, native_country, plus_50 from ";
+	private static String TRAIN_QUERY =  MAIN_SQL + "adult";
+	private static String TEST_QUERY = MAIN_SQL + "adult_test";
+	
+	private static Instances numericToNominal(Instances data, int []attributes) throws Exception {
+		NumericToNominal filter = new NumericToNominal();
+		filter.setAttributeIndicesArray(attributes);
+		filter.setInputFormat(data);
+		return Filter.useFilter(data, filter);
+	}
+	
+	private static void changeUnknown(Instances data, int attr)
+	{
+		for(int i = 0; i < data.numInstances(); ++i) {
+			Instance inst = data.instance(i);
+			if(inst.stringValue(attr).equals("unknown")) {
+				inst.setMissing(attr);
+			}
+		}
+	}
[…]

### 2009-12-06 weka/Bayes.java
Simulated Annealing + K2

@@ -1,20 +1,13 @@
-import weka.core.Instances;
 import weka.classifiers.Evaluation;
 import weka.classifiers.bayes.BayesNet;
+import weka.core.Instances;
 import weka.filters.Filter;
 import weka.filters.unsupervised.instance.RemovePercentage;
 
-public class Genetic {
-	public static void main(String[] args) throws Exception {
-		if(args.length != 2) {
-			System.err.println("usage: C45 <db user> <db password>");
-			System.exit(1);
-		}
-		
-		String user = args[0];
-		String password = args[1];
-	
-		
+
+public class Bayes {
+	public static void runBayes(String user, String password, String search_alg, int perc) throws Exception
+	{
 		Instances train_data = Utils.getTrainSet(user, password);
 		/*System.out.println("====================================");
 		System.out.println("TRAIN DATA:");
@@ -28,14 +21,14 @@ public class Genetic {
 		System.out.println(test_data.toSummaryString());
 */
 		RemovePercentage rp = new RemovePercentage();
-		rp.setPercentage(99); // usar apenas 1% das instancias
+		rp.setPercentage(perc);
 		rp.setInputFormat(train_data);
 		Instances new_data = Filter.useFilter(train_data, rp);
 		
 		BayesNet net = new BayesNet();
 		
 		String options = "" +
-			"-Q weka.classifiers.bayes.net.search.local.GeneticSearch -- -L 3 -A 3 -U 2 -C -M -R 1 " +
+			"-Q " + search_alg + " " +
 			"-E weka.classifiers.bayes.net.estimate.SimpleEstimator -- -A 1.0";
 		System.out.println(options);
 		
@@ -55,5 +48,19 @@ public class Genetic {
 		Evaluation test_eval = new Evaluation(train_data);
 		test_eval.evaluateModel(net, test_data);
 		System.out.println(test_eval.toSummaryString("\nTest Results\n===============\n", false));
+	}
+	
+	public static void main(String[] args) throws Exception {
+		if(args.length != 4) {
+			System.err.println("usage: bayes <db user> <db password> <search alg>");
+			System.exit(1);
+		}
+
+		String user = args[0];
+		String password = args[1];
+		String alg = args[2];
+		int perc = Integer.parseInt(args[3]);
+
+		Bayes.runBayes(user, password, alg, 100 - perc);
 	}
 }

### 2009-12-06 weka/genetic.sh

@@ -1,0 +1,3 @@
+#!/bin/sh
+
+java Bayes $1 $2 "weka.classifiers.bayes.net.search.local.GeneticSearch -- -L 3 -A 3 -U 2 -C -M -R 1" 1

### 2009-12-06 weka/k2.sh

@@ -1,0 +1,3 @@
+#!/bin/sh
+
+java Bayes $1 $2 "weka.classifiers.bayes.net.search.local.K2 -- -P 100 -R" 10

### 2009-12-06 weka/simulated_annealing.sh

@@ -1,0 +1,3 @@
+#/bin/sh
+
+java Bayes $1 $2 "weka.classifiers.bayes.net.search.local.SimulatedAnnealing -- -U 50 -A 20 -D 0.4 -R 1" 10

### 2009-12-06 weka/hill_climber.sh
hill climbing + bayes

@@ -1,0 +1,3 @@
+#!/bin/sh
+
+java Bayes $1 $2 "weka.classifiers.bayes.net.search.local.HillClimber -- -P 100" 5

### 2009-12-07 weka/RandomForest.java
random forest algorithm

@@ -1,0 +1,44 @@
+import weka.classifiers.Evaluation;
+import weka.core.Instances;
+
+
+public class RandomForest {
+	public static void main(String []args) throws Exception {
+		if(args.length != 2) {
+			System.err.println("usage: C45 <db user> <db password>");
+			System.exit(1);
+		}
+
+		String user = args[0];
+		String password = args[1];
+
+		Instances train_data = Utils.getTrainSet(user, password);
+		System.out.println("====================================");
+		System.out.println("TRAIN DATA:");
+		System.out.println("====================================");
+		System.out.println(train_data.toSummaryString());
+
+		Instances test_data = Utils.getTestSet(user, password);
+		System.out.println("====================================");
+		System.out.println("TEST DATA:");
+		System.out.println("====================================");
+		System.out.println(test_data.toSummaryString());
+		
+		weka.classifiers.trees.RandomForest rf = new weka.classifiers.trees.RandomForest();
+		rf.setNumTrees(100); // change here
+		rf.setSeed((int)(System.currentTimeMillis() / 1000000));
+		
+		rf.buildClassifier(train_data);
+		
+		System.out.println(rf.toString());
+		
+		Evaluation train_eval = new Evaluation(train_data);
+		train_eval.evaluateModel(rf, train_data);
+		System.out.println(train_eval.toSummaryString("\nTrain Results\n===============\n", false));
+	
+		Evaluation test_eval = new Evaluation(train_data);
+		test_eval.evaluateModel(rf, test_data);
+		System.out.println(test_eval.toSummaryString("\nTest Results\n================\n", false));
+
+	}
+}

### 2009-12-07 weka/SimpleCart.java
simple cart

@@ -1,0 +1,43 @@
+import weka.classifiers.Evaluation;
+import weka.core.Instances;
+
+public class SimpleCart {
+	public static void main(String []args) throws Exception {
+		if(args.length != 2) {
+			System.err.println("usage: C45 <db user> <db password>");
+			System.exit(1);
+		}
+
+		String user = args[0];
+		String password = args[1];
+
+		Instances train_data = Utils.getTrainSet(user, password);
+		System.out.println("====================================");
+		System.out.println("TRAIN DATA:");
+		System.out.println("====================================");
+		System.out.println(train_data.toSummaryString());
+
+		Instances test_data = Utils.getTestSet(user, password);
+		System.out.println("====================================");
+		System.out.println("TEST DATA:");
+		System.out.println("====================================");
+		System.out.println(test_data.toSummaryString());
+		
+		weka.classifiers.trees.SimpleCart sc = new weka.classifiers.trees.SimpleCart();
+		sc.setUsePrune(true);
+		sc.setHeuristic(true);
+		sc.setSeed((int)(System.currentTimeMillis() / 1000000));
+		sc.buildClassifier(train_data);
+		
+		System.out.println(sc.toString());
+		
+		Evaluation train_eval = new Evaluation(train_data);
+		train_eval.evaluateModel(sc, train_data);
+		System.out.println(train_eval.toSummaryString("\nTrain Results\n===============\n", false));
+	
+		Evaluation test_eval = new Evaluation(train_data);
+		test_eval.evaluateModel(sc, test_data);
+		System.out.println(test_eval.toSummaryString("\nTest Results\n================\n", false));
+
+	}
+}

### 2009-12-07 weka/ADTree.java
ADTree

@@ -1,0 +1,44 @@
+import weka.classifiers.Evaluation;
+import weka.core.Instances;
+
+
+public class ADTree {
+	public static void main(String []args) throws Exception {
+		if(args.length != 2) {
+			System.err.println("usage: C45 <db user> <db password>");
+			System.exit(1);
+		}
+
+		String user = args[0];
+		String password = args[1];
+
+		Instances train_data = Utils.getTrainSet(user, password);
+		System.out.println("====================================");
+		System.out.println("TRAIN DATA:");
+		System.out.println("====================================");
+		System.out.println(train_data.toSummaryString());
+
+		Instances test_data = Utils.getTestSet(user, password);
+		System.out.println("====================================");
+		System.out.println("TEST DATA:");
+		System.out.println("====================================");
+		System.out.println(test_data.toSummaryString());
+		
+		weka.classifiers.trees.ADTree adt = new weka.classifiers.trees.ADTree();
+		
+		adt.setRandomSeed((int)(System.currentTimeMillis() / 1000000));
+		adt.setNumOfBoostingIterations(40);
+		adt.buildClassifier(train_data);
+		
+		System.out.println(adt.toString());
+		
+		Evaluation train_eval = new Evaluation(train_data);
+		train_eval.evaluateModel(adt, train_data);
+		System.out.println(train_eval.toSummaryString("\nTrain Results\n===============\n", false));
+	
+		Evaluation test_eval = new Evaluation(train_data);
+		test_eval.evaluateModel(adt, test_data);
+		System.out.println(test_eval.toSummaryString("\nTest Results\n================\n", false));
+
+	}
+}

### 2009-12-08 report/final.tex
initial import for the final report

--- 2009-11-12 report/paper.tex
@@
 \usepackage{longtable}
-\usepackage[latin1]{inputenc}
+\usepackage[utf8]{inputenc}
 \usepackage{subfigure}
@@
 \begin{document}
-\title{Progress Report on the ``Adult Database'' Analysis}
+\title{A comparison of classification algorithms applied to the Adult Database}
 \author{Flávio Cruz \and João Azevedo}
@@
 
-Classification in data mining is the activity of identifying a given object as 
-belonging to a pre-defined class \cite{1}. The ultimate goal is to learn a concept. This
-learning process is based on past observations, where the dataset performs a
-main role. This report gives a first overview on the ``Adult Database'' dataset,
-describing its format and identifying the problem it poses. Some data analysis
-is then done, in order to obtain some characteristics on the global set of
-values. Finally, the future work plan is presented and some conclusions on the
-subject are taken.
+\subsection{Problem Identification}
 
-\section{Data Format}
+Deducing from the previous section, this is a classification problem. Given an adult
+with the mentioned attributes we want to classify him into one of the following groups:
+one gains more than 50000 dollars per year, the other group does not.
+
+One important objective in classifying things is to understand why the given object
+is put under one class and not in another \cite{2}\cite{3}, that is, we want to understand what social characteristics
+allows one person to gain more than someone else, if the sex is important, the education level,
+the race and what factors are less important.
+
+Given a specific set of persons from a certain society and from a certain time we could then
+learn what contributes to a person's social standing and, as we all know, income is an important
+metric in that department. We could, for example, learn if the race affects the persons success and if
+that society discriminates against certain races.
+
+Understanding what affects the income can give a lot of information concerning society's behaviors.
+
+\section{Experimental Context}
 
@@
 
-\section{Problem Identification}
-
-Deducing from the previous section, this is a classification problem. Given an adult
-with the mentioned attributes we want to classify him into one of the following groups:
-one gains more than 50000 dollars per year, the other group does not.
-
-One important objective in classifying things is to understand why the given object
-is put under one class and not in another \cite{2}\cite{3}, that is, we want to understand what social characteristics
-allows one person to gain more than someone else, if the sex is important, the education level,
-the race and what factors are less important.
-
-Given a specific set of persons from a certain society and from a certain time we could then
-learn what contributes to a person's social standing and, as we all know, income is an important
-metric in that department. We could, for example, learn if the race affects the persons success and if
-that society discriminates against certain races.
-
-Understanding what affects the income can give a lot of information concerning society's behaviors.
-
-\section{Data Analysis}
-
-The table \ref{tab:summary} summarizes the distribution of values on the studied
-dataset. From a first observation, one can see that the amount of missing values
-is low. On the other side, in some of the attributes, some classes have a
-much higher frequency than others, such as the 'united-states' native country, 
-the 'private' workclass and the 'white' race. On the hours of work per week
-attribute, there is a great amount of outliers, with some unrealistic
-values, namely with amounts close to the 100 hours or less than 10 in 
-individuals with a defined occupation, as can be seen on figure 
-\ref{fig:box_hours}. Variables such has the capital gain and capital loss have
-few values above 0, being quite specific to certain classes of individuals. The
-sex is also unbalanced in the sample data, as the amount of men is more than
-twice the amount of women.
-
-It is also important to verify some correlations between variables. To that
-matter, we've tested some apparent situations, such as the influence of the age,
-hours of work per week and educational level on the amount gained by an
-individual. The figures \ref{fig:age_plus50}, \ref{fig:education_plus50} and
-\ref{fig:hpw_plus50} show that there is a tendency for older people, people with
-a higher educational level and people who work a great amount of hours per week
-to gain a larger amount of money.
-
-\subsection{Association Rules}
-
-Using association rules one can find frequent patterns and associations from a
-set of transactions. These rules can easily be used to make data analysis and
-to understand casual structures.
-
-To build a set of association rules we used the apriori algorithm that is featured
-in the R arules package. Before running the algorithm we made some data preprocessing
-for continuous attributes that are needed to build better and more general rules.
-
-The following attributes were transformed:
-
-\begin{itemize}
-  \item {\textbf{Age}: turned into categories. 15-25 (young), 25-45 (middle\_aged), 45-65 (senior), $>$ 65 (old).}
-  \item {\textbf{Hours-per-week}: turned into categories. 0-25 (part\_time), 25-40 (full\_time), 40-60 (overtime), $>$60 (workaholic).}
-  \item {\textbf{Capital-gain}: turned into categories. $<=$ 0 (none), $<$ median value (low), $>$ median value (high).}
-  \item {\textbf{Capital-loss}: turned into categories. $<=$ 0 (none), $<$ median value (low), $>$ median value (high).}
-\end{itemize}
-
-Next, we plotted the relative frequency of items in the transaction database
-with support greater than 10\%. The results are shown in figure \ref{fig:arules_item_frequency}.
-We can learn various things from that plot: the majority of people are from the united states,
-the white race is predominant, males outnumber females and that more people work in full time.
-
-Finally we run the apriori algorithm using the test data set to generate rules
-with support greater than 2\% and at least 60\% of confidence. The algorithm built
-50487 rules.
-
-We then built two sets of rules from the original set: one with the consequent component
-with plus\_50=0 (the person gains less than 50K a year) and the other with plus\_50=1
-(the person gains more than 50K a year).
-
-Each rule set was sorted by the confidence level, from more confident to less confident.
-
-The first 5 rules from the plus\_50=0 set are shown in listing \ref{arules_small_income} and
-are related to small incomes. We can easily make some observations
-about those rules: young people with a part-time are probably low income and having a child
-without being married is also not a very good signal.
-
-The second set, contrary to the ones above, shows us the top 5 rules for having large incomes.
-These rules are shown in listing \ref{arules_large_income}. A conclusion can readily be made:
-being a white husband with a high capital gain probably means we have a large income.
-
-\subsection{Clusters}
-
-Another method to analyze the dataset is based on the generation of clusters. 
-Clusters are sets of data in which objects in the same cluster are similar, and
-objects in different clusters are dissimilar. The first attempt at clustering
-used only three variables: age, education\_num and hours\_per\_week, the same
-variables we saw that had a greater impact on the amount gained by the
-individual. The data was normalized using the Z-Transformation method and the 
-K-Means method was used to obtain the final clusters, with a k of 6, using RapidMiner \cite{4}. Figures 
-\ref{fig:cluster_1_age_education}, \ref{fig:cluster_1_education_hpw} and \ref{fig:cluster_1_age_hpw} show
-scatter plots of the variables, colored by clusters. One thing one is able to
-note is that people with younger age belong to cluster \#2. In terms of
-education, people with higher education belong to cluster \#0, while people with
-lower education belong to cluster \#1. In terms of hours of work per week, 
-distribution is rather similar, but cluster \#2 and \#3 have a greater amount of
-lower and higher values, respectively. Figure \ref{fig:cluster_1_cluster_plus50}
-shows how clusters relate to the class values. Although the distribution is 
-balanced, clusters \#1 and \#2 have a greater amount of people gaining below 
-50000 dollars. Notice, as stated before, that this were the clusters having 
-people with lower education and younger age, respectively.
-
-Clustering with the remaining attributes leads to results more difficult to
-analyze. The same process was applied, using the Z-Transformation method for
-normalization and the K-Means algorithm with a k of 6, but this time with all
-the attributes. The distribution of values is more balanced, although a very
-high value of capital gain, very unusual on the data set, leads to the creation
-of a very well identified cluster, as can be seen on table \ref{tab:centroids}.
-Particular values of race, workclass and native country also identify clusters
-\#2, \#3 and \#2, respectively. Notice that this attributes were the ones
-refered to before as having a given value more frequent than others, thus
-leading to this particular composition of clusters.
-
-\section{Work plan}
-
-Given the problem at hand, we want to generate a classification mechanism
-that can tell if a specific person has a small or large income.
-
-For this we will use various kinds of classification algorithms, from classical
-decision trees to rule induction.
-
-We will try to compare the error rate among the various algorithms. We'll be using
-the train set to "train" the algorithm and then the test set to assess which algorithm
-is best suited to make income predictions.
-
-We'll also use some data-processing techniques to eliminate unknown values. It will
-also be useful to categorize continuous values for more generalized rules as we have
-done when generating association rules.
-
-The algorithms we will be using are: ID3, C4.5, CN2 and AQ. Depending on our progress
-we will experiment with neural networks and genetic algorithms, which are other
-mechanisms for building classifiers.
-
 \section{Conclusions}
@@
 
-\begin{table}
-\centering
-\subtable[Age]{
-       \begin{tabular}{| l | l |}
-       \hline
-       Min. & 17 \\
-       \hline
-       1st Qu. & 28 \\
-       \hline
-       Median & 37 \\
-       \hline
-       Mean & 38.58 \\
-       \hline
-       3rd Qu. & 48 \\
-       \hline
-       Max & 90 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Workclass]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       private & 22696 \\
-       \hline
-       self\_emp\_not\_inc & 2541 \\
-       \hline
-       local\_gov & 2093 \\
-       \hline
-       state\_gov & 1298 \\
-       \hline
-       self\_emp\_inc & 1116 \\
-       \hline
-       (Other) & 981 \\
-       \hline
-       NA's & 1836 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Education]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       hs\_grad & 10501 \\
-       \hline
-       some\_college & 7291 \\
-       \hline
-       bachelors & 5535 \\
-       \hline
-       masters & 1723 \\
-       \hline
-       assoc\_voc & 1382 \\
-       \hline
-       11th & 1175 \\
-       \hline
-       (Other) & 5134 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Marital Status]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       divorced & 4443 \\
-       \hline
-       married\_af\_spouse & 23 \\
-       \hline
-       married\_civ\_spouse & 14976 \\
-       \hline
-       married\_spouse\_absent & 418 \\
-       \hline
-       never\_married & 10683 \\
-       \hline
-       separated & 1025 \\
-       \hline
-       widowed & 993 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Occupation]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       prof\_specialty & 4140 \\
-       \hline
-       craft\_repair & 4099 \\
-       \hline
-       exec\_managerial & 4066 \\
-       \hline
-       adm\_clerical & 3770 \\
-       \hline
-       sales & 3650 \\
-       \hline
-       (Other) & 10993 \\
-       \hline
-       NA's & 1843 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Relatonship]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       husband & 13193 \\
-       \hline
-       not\_in\_family & 8305 \\
-       \hline
-       other\_relative & 981 \\
-       \hline
-       own\_child & 5068 \\
-       \hline
-       unmarried & 3446 \\
-       \hline
-       wife & 1568 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Race]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       amer\_indian\_eskimo & 311 \\
-       \hline
-       asian\_pac\_islander & 1039 \\
-       \hline
-       black & 3124 \\
-       \hline
-       other & 271 \\
-       \hline
-       white & 27816 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Sex]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       female & 10771 \\
-       \hline
-       male & 21790 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Capital Gain]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       Min. & 0 \\
-       \hline
-       1st Qu. & 0 \\
-       \hline
-       Median & 0 \\
-       \hline
-       Mean & 1078 \\
-       \hline
-       3rd Qu. & 0 \\
-       \hline
-       Max. & 99999 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Capital Loss]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       Min. & 0 \\
-       \hline
-       1st Qu. & 0 \\
-       \hline
-       Median & 0 \\
-       \hline
-       Mean & 87.3 \\
-       \hline
-       3rd Qu. & 0 \\
-       \hline
-       Max. & 4356 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Hours Per Week]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       Min. & 1 \\
-       \hline
-       1st Qu. & 40 \\
-       \hline
-       Median & 40 \\
-       \hline
-       Mean & 40.44 \\
-       \hline
-       3rd Qu. & 45 \\
-       \hline
-       Max. & 99 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Native Country]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       United States & 29170 \\
-       \hline
-       Mexico & 643 \\
-       \hline
-       Philippines & 198 \\
-       \hline
-       Germany & 137 \\
-       \hline
-       Canada & 121 \\
-       \hline
-       (Other) & 1709 \\
-       \hline
-       NA's & 583 \\
-       \hline
-       \end{tabular}
-}
-\qquad\qquad
-\subtable[Plus 50]{        
-       \begin{tabular}{| l | l |}
-       \hline
-       No & 24720 \\
-       \hline
-       Yes & 7841 \\
-       \hline
-       \end{tabular}
-}
-\caption{Summary of the value distribution in the database}
-\label{tab:summary}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{ | l | l | l | l | l | l | l | }
-\hline
-\textbf{Attribute} & \textbf{Centroid \#0} & \textbf{Centroid \#1} & \textbf{Centroid \#2} & \textbf{Centroid \#3} & \textbf{Centroid \#4} & \textbf{Centroid \#5} \\
-\hline
-age	& -0.874 & 0.692 & -0.166 & 0.246 & 0.570 & 0.190 \\
-\hline
-workclass & -0.272 & -0.063 & -0.036 & 2.208 & 0.455 & -0.364 \\
-\hline
-education\_num & -0.101 & -0.172 & -0.177 & 0.0309 & 1.103 & 0.102 \\
-\hline
-marital\_status & -0.679 & 1.693 & -0.083 & -0.150 & -0.007 & -0.108 \\
-\hline
-occupation & -0.305 & -0.303 & 0.091 & 0.952 & -0.434 & 0.010 \\
-\hline
-relationship & 0.703 & 0.526 & 0.189 & -0.053 & -0.351 & -0.507 \\
-\hline
-race & -0.086 & -0.020 & 2.630 & -0.147 & -0.002 & -0.230 \\
-\hline
-sex & 0.662 & 1.199 & 0.019 & -0.161 & -0.409 & -0.635 \\
-\hline
-capital\_gain & -0.116 & -0.091 & -0.086 & -0.029 & 13.394 & -0.040 \\
-\hline
-capital\_loss & -0.115 & -0.079 & -0.021 & 0.025 & -0.217 & 0.079 \\
-\hline
-hours\_per\_week & -0.523 & -0.231 & -0.031 &-0.071 & 0.758 & 0.339 \\
-\hline
-native\_country & -0.186 & -0.154 & 2.951 & -0.181 & -0.075 & -0.171 \\
-\hline
-\end{tabular}
-\caption{Centroid table for the results of the application of the K-Means algorithm with a k of 6}
-\label{tab:centroids}
-\end{table}
-
-\clearpage
-
-\section{Figures}
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm]{boxplot_hours_per_week.png}
-\caption{Boxplot of the hours per week variable}
-\label{fig:box_hours}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm]{plot_plus_50_age.png}
-\caption{Boxplot of the individuals' age, for each class of gain}
-\label{fig:age_plus50}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm]{plot_plus_50_education_num.png}
-\caption{Boxplot of the individuals' education level, for each class of gain}
-\label{fig:education_plus50}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm]{plot_plus_50_hpw.png}
-\caption{Boxplot of the individuals' hours of work per week, for each class of 
-gain}
-\label{fig:hpw_plus50}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm]{cluster_1_age_education.png}
-\caption{Scatter plot of age and education}
-\label{fig:cluster_1_age_education}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm]{cluster_1_education_hpw.png}
-\caption{Scatter plot of education and hours of work per week}
-\label{fig:cluster_1_education_hpw}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm]{cluster_1_age_hpw.png}
-\caption{Scatter plot of age and hours of work per week}
-\label{fig:cluster_1_age_hpw}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=10cm]{cluster_1_cluster_plus50.png}
-\caption{Scatter plot of cluster and class values of amount gained}
-\label{fig:cluster_1_cluster_plus50}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=20cm,angle=90]{item_frequency.png}
-\caption{Item frequencies of items with at least 10\% support.}
-\label{fig:arules_item_frequency}
-\end{figure}
-
-\clearpage
-\section{Listings}
-
-\begin{lstlisting}[language=r,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Association rules for small income.},label=arules_small_income]
-  lhs                               rhs            support confidence     lift
-1 {age=young,                                                                 
-   hours_per_week=part_time}     => {plus_50=0} 0.05730782          1 1.317193
-2 {age=young,                                                                 
-   occupation=unknown,                                                        
-   capital_gain=none}            => {plus_50=0} 0.02002396          1 1.317193
-3 {age=young,                                                                 
-   relationship=own_child,                                                    
-   hours_per_week=part_time}     => {plus_50=0} 0.04404042          1 1.317193
-4 {education=some_college,                                                    
-   relationship=own_child,                                                    
-   hours_per_week=part_time}     => {plus_50=0} 0.02069961          1 1.317193
-5 {marital_status=never_married,                                              
-   relationship=own_child,                                                    
-   hours_per_week=part_time}     => {plus_50=0} 0.04689659          1 1.317193
-\end{lstlisting}
-
-\begin{lstlisting}[language=r,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Association rules for large income.},label=arules_large_income]
-  lhs                                    rhs            support confidence     lift
-1 {relationship=husband,                                                           
-   race=white,                                                                     
-   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
-2 {marital_status=married_civ_spouse,                                              
-   relationship=husband,                                                           
-   race=white,                                                                     
-   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
-3 {relationship=husband,                                                           
-   race=white,                                                                     
-   sex=male,                                                                       
-   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
-4 {relationship=husband,                                                           
-   race=white,                                                                     
-   capital_gain=high,                                                              
-   capital_loss=none}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
-5 {relationship=husband,                                                           
-   capital_gain=high}                 => {plus_50=1} 0.02251159  0.9918809 4.118943
-\end{lstlisting}
-
 \end{document}

### 2009-12-08 report/final.tex
algorithms

@@
 Understanding what affects the income can give a lot of information concerning society's behaviors.
+
+\section{Algorithms}
+
+The algorithms we used to build classifiers can be grouped into four main
+groups: decision trees, rule induction \cite{rule_induction},
+inductive logic programming and bayesian classifiers \cite{bayes}.
+
+\subsection{Decision trees}
+
+For decision trees, the algorithms used are ID3 \cite{id3}, C4.5 \cite{c45}, CART \cite{cart},
+Random Forest \cite{random_forest}, and alternating decision trees (adtree) \cite{adtree}.
+
+\subsection{Rule induction}
+
+To experiment with rule induction algorithms we chose the CN2 \cite{cn2} algorithm.
+
+\subsection{Inductive Logic Programming}
+
+The Aleph ILP system \cite{aleph} was the targeted inductive logic programming system, Aleph uses
+the inverse entailment algorithm to construct hypotheses.
+
+\subsection{Bayesian classifiers}
+
+For bayesian classifiers we tested different kinds of search algorithms to build
+the network structure.
+
+The construction of the network structured is also known as
+\textit{local score metrics} and various algorithms can be employed for this optimization problem.
+
+The experiments conducted will use genetic based search, hill climbing \cite{bayes3}, simulated annealing,
+and a modified hill climbing algorithm, named K2 \cite{bayes2}.
+
+For the second step in building a bayesian network classifier, we used a
+simple estimator to estimate the conditional probability distributions.
+This simple estimator produces
+direct estimates of the conditional probabilities. \cite{bayes_weka}
 
@@
 
+\section{Experimental Evaluation}
+
+\section{Related Work}
+
 \section{Conclusions}
@@
 RapidMiner; \url{http://www.rapidminer.com/}, November 2009
-\bibitem{5}
+\bibitem{c45}
 Ross Quinlan, J.; C4.5: Programs for Machine Learning; Morgan Kaufmann, 1992
@@
 Hahsler, M., Gruen, B.; Hornik, K.; arules -- A Computational Environment for Mining Association Rules and Frequent Item Sets; Journal of Statistical Software 14/15, 2005
+
+\bibitem{id3}
+Quinlan, J. R. 1986. Induction of Decision Trees. Mach. Learn. 1, 1 (Mar. 1986), 81-106.
+
+\bibitem{cart}
+Leo Breiman, Jerome H. Friedman, Richard A. Olshen, Charles J. Stone (1984). Classification and Regression Trees. Wadsworth International Group, Belmont, California.
+
+\bibitem{random_forest}
+Liaw, Andy and Wiener, Matthew. Classification and Regression by randomForest. R News (2002) Vol. 2/3 p. 18
+
+\bibitem{adtree}
+Freund, Y., Mason, L.: The alternating decision tree learning algorithm. In: Proceeding of the Sixteenth International Conference on Machine Learning, Bled, Slovenia, 124-133, 1999
+
+\bibitem{bayes}
+N. Friedman, D. Geiger, M. Goldszmidt. Bayesian Network Classifiers. Machine Learning, 29: 131–163, 1997
+
+\bibitem{rule_induction}
+Quinlan, J.
+R.
+(1987).
+"Generating production rules from decision trees".
+in McDermott, John.
+Proceedings of the Tenth International Joint Conference on Artificial Intelligence (IJCAI-87).
+Milan, Italy.
+pp.
+304–307
+
+\bibitem{aleph}
+Ashwin Srinivasan. The Aleph Manual. 2007
+
+\bibitem{bayes2}
+G. Cooper, E. Herskovits. A Bayesian method for the induction of probabilistic networks from 
+data. Machine Learning, 9: 309–347, 1992.
+
+\bibitem{bayes3}
+W.L. Buntine. A guide to the literature on learning probabilistic networks from data. IEEE 
+Transactions on Knowledge and Data Engineering, 8:195–210, 1996.
+
+\bibitem{bayes_weka}
+Remco R. Bouckaert. Bayesian Network Classifiers in Weka. 2004
+
 \end{thebibliography}

### 2009-12-08 report/final.tex
more algorithms

@@
 
+Decision tree learning is a very commonly used method in classification. The main goal here is to
+create a tree model that can predict an object's class. Each interior node corresponds
+to one test for the object's attributes and each edge links to new interiors nodes based on the corresponding
+results. Each leaf in the tree represents the object class for the tests made
+across the respective path, from root to leaf.
+
+The ID3 algorithm (Iterative Dichotomiser 3) was one of the first decision trees algorithms.
+It was designed by Ross Quinlan.
+
+The C4.5 algorithm is the improved version of the ID3 algorithm, also designed by Quinlan.
+Contrary to ID3, this algorithm can handle both continuous and discrete attributes,
+it can use data with missing values and can also prune its decision trees.
+
+The CART algorithm builds classification and regression trees for predicting continuous dependent variables (regression)
+and categorical predictor variables (classification). The classic CART algorithm was popularized by Breiman et al. \cite{cart}
+
+The alternating decision tree is another type of decision tree algorithms. This tree
+consists of decision nodes, which specify a predicate condition, and prediction nodes, containing a single number.
+Each object is classified by following all paths for which all decision nodes are true and summing any
+prediction nodes along the way. Contrary to binary classification like ID3 ou C4.5, in adtrees one
+object can follow multiple paths.
+
+Finally, the Random Forest algorithm is an ensemble classifier which consists of many decision trees.
+Each object is classified multiple times by all decision trees and the mode class is used as the final class.
+Like the CART algorithm, Random Forest was also developed by Leo Breiman.
+
 \subsection{Rule induction}
@@
 
+Rule induction works by extracting formal rules from a set of observations.
+Usually rules are expressed in the form \textit{if condition1 and condition2 ... then (decision, value)}.
+
+CN2 is an algorithm for inducing 
+propositional classification rules. CN2 consists of two main procedures: the 
+search procedure that performs beam search in order to find a single rule and 
+the control procedure that repeatedly executes the search. 
+The search procedure performs beam search using classification accuracy 
+of the rule as a heuristic function.
+Two different control procedures are used in CN2: one for inducing an ordered 
+list of rules and the other for the unordered case.
+
 \subsection{Inductive Logic Programming}
 
-The Aleph ILP system \cite{aleph} was the targeted inductive logic programming system, Aleph uses
-the inverse entailment algorithm to construct hypotheses.
+Hypothesis construction is the core of any ILP system and there exists many different techniques for constructing them, including: Inverse Resolution, Relative Least General Generalisations, Inverse Implication, and Inverse Entailment.
+
+We will be using the Aleph system \cite{aleph} to experiment with inductive logic programming. This system
+uses the Inverse Entailment algorithm and has also incorporated features from many of its predecessors based on other techniques.
 

### 2009-12-08 report/final.tex
more experimental context.

@@
 
-\section{Algorithms}
+\section{Algorithms} \label{sec:algs}
 
@@
 
-\section{Experimental Context}
+\section{Data Format}
 
@@
 This data set is available in the following URL: \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}.
+
+\subsection{Experimental Context}
+
+In this section we will describe the experimental context for each
+algorithm described in the section \ref{sec:algs}, namely, programs we used,
+preprocessing tasks conducted, the parameterization that is available for each
+algorithm and so forth.
+
+The data set available in \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}
+is composed of two CSV files, one for the training set and the other for the test set.
+Each file was imported into a MySQL database and grouped into two tables: adult and adult\_test.
+These tables follow the same structure (Table \ref{tbl:mysql_table}).
+
+\begin{table}[ht]
+  \begin{tabular}{ | l | p{14cm} |}
+    \hline
+    \textbf{Field} & \textbf{Type} \\ \hline
+    id & SERIAL \\ \hline
+    age & INT \\ \hline
+    workclass & enum('private', 'self\_emp\_not\_inc', 'self\_emp\_inc','federal\_gov', 'local\_gov', 'state\_gov', 'without\_pay', 'never\_worked', 'unknown') \\ \hline
+    fnlwgt & INT \\ \hline
+    education & enum('bachelors', 'some\_college', '11th', 'hs\_grad', 'prof\_school', 'assoc\_acdm', 'assoc\_voc', '9th', '7th\_8th', '12th', 'masters', '1st\_4th', '10th', 'doctorate', '5th\_6th', 'preschool') \\ \hline
+    education\_num & INT \\ \hline
+    marital\_status & enum('married\_civ\_spouse', 'divorced', 'never\_married', 'separated', 'widowed', 'married\_spouse\_absent', 'married\_af\_spouse') \\ \hline
+    occupation & enum('tech\_support', 'craft\_repair', 'other\_service', 'sales', 'exec\_managerial', 'prof\_specialty', 'handlers\_cleaners', 'machine\_op\_inspct', 'adm\_clerical', 'farming\_fishing', 'transport\_moving', 'priv\_house\_serv', 'protective\_serv', 'armed\_forces', 'unknown') \\ \hline
+    relationship & enum('wife', 'own\_child', 'husband', 'not\_in\_family', 'other\_relative', 'unmarried') \\ \hline
+    race & enum('white', 'asian\_pac\_islander', 'amer\_indian\_eskimo', 'other', 'black') \\ \hline
+    sex & enum('male', 'female') \\ \hline
+    capital\_gain & INT \\ \hline
+    capital\_loss & INT \\ \hline
+    hours\_per\_week & INT \\ \hline
+    native\_country & enum('united\_states', 'cambodia', 'england', 'puerto\_rico', 'canada', 'germany', 'outlying\_us', 'india', 'japan', 'greece', 'south', 'china', 'cuba', 'iran', 'honduras', 'philippines', 'italy', 'poland', 'jamaica', 'vietnam', 'mexico', 'portugal', 'ireland', 'france', 'dominican\_republic', 'laos', 'ecuador', 'taiwan', 'haiti', 'columbia', 'hungary', 'guatemala', 'nicaragua', 'scotland', 'thailand', 'yugoslavia', 'el\_salvador', 'trinadad\_tobago', 'peru', 'hong', 'holand\_netherlands', 'unknown') \\ \hline
+    plus\_50 & BOOLEAN \\ \hline
+  \end{tabular}
+  \caption{MySQL table structure.}
+  \label{tbl:mysql_table}
+\end{table}
+
+For the majority of the proposed algorithms we tried to make use of the database, instead
+of using the CSV files, as it allows a more flexible approach to data querying using SQL.
+
+\subsection{ID3}
+
+\subsection{}
 

### 2009-12-08 report/final.tex
more experimental context

@@
 
-\subsection{Experimental Context}
+\section{Experimental Context}
 
@@
 
-\subsection{ID3}
+\subsection{Weka}
 
-\subsection{}
+Weka is a collection of machine learning algorithms for data mining tasks. \cite{weka}
+It can be applied directly to a dataset using command line tools or through Java code, using the Weka Java APIs.
+Weka supports tools for data pre-processing, classification, regression, clustering, association rules,
+and visualization.
 
-\section{Experimental Evaluation}
+This tool was used to apply the following algorithms: C4.5, ADTree, Random Forest, CART and Bayesian Networks.
+Instead of using the command line tools, we used the Java API to build classifiers.
+
+To use the data set we fetch it directly from the database and then apply various data pre-processing
+techniques that are available within Weka:
+
+\begin{itemize}
+  \item The \textbf{plus\_50} was converted from a numerical format to categories: $<$50K and $>$50K.
+  \item Every field marked as \textbf{unknown} was marked as \textbf{missing}.
+  \item The \textbf{age} was discretized into various categories: \textbf{young} ($<$25),
+    \textbf{middle\_aged} (25-45), \textbf{senior} (45-65), and \textbf{old} ($>$65).
+  \item The attribute \textbf{hours\_per\_week} was also discretized: \textbf{part\_time} ($<$25),
+    \textbf{full\_time} (25-40), \textbf{over\_time} (40-60), and \textbf{workaholic} ($>$60).
+  \item Each capital attribute (\textbf{capital\_gain} and \textbf{capital\_loss}) was also discretized:
+    \textbf{none} (0), \textbf{very\_low} (0-100), \textbf{low} (100-1000), \textbf{medium} (1000-5000),
+    \textbf{high} (5000-10000), and \textbf{very\_high} ($>$10000).
+\end{itemize}
+
+Only the following values are feed into the algorithms: age, workclass, education,
+marital\_status, occupation, relationship, race, sex, capital\_gain, capital\_loss,
+hours\_per\_week, native\_country, and plus\_50. The attributes education\_num
+and fnlwgt are derived from other attributes and so we keep them out.
+
+For missing values we used the \textit{ReplaceMissingValues} filter available in Weka that
+replaces unknown values with the modes and means from the data set.
+
+Each algorithm has various configuration parameters that will be explored
+and analyzed in section \ref{sec:experimental_eval}.
+For example, the Bayesian Network
+algorithm allows the parameterization of the search algorithm that optimizes the network structure.
+
+\section{Experimental Evaluation} \label{sec:experimental_eval}
 
@@
 
+\bibitem{weka}
+Mark Hall, Eibe Frank, Geoffrey Holmes, Bernhard Pfahringer, Peter Reutemann, Ian H. Witten (2009); The WEKA Data Mining Software: An Update; SIGKDD Explorations, Volume 11, Issue 1.
+
 \end{thebibliography}

### 2009-12-08 report/final.tex
CN2 paper.
some rework

@@
 
+After the data as been pre-processed,
+we will build the classifier using the train data set.
+Once the classifier is trained, it will be used to classify both the train and test data.
+
+Various performance metrics for both classification results will be registered:
+error rate, mean absolute error, kappa statistic, etc.
+
 \section{Experimental Evaluation} \label{sec:experimental_eval}
@@
 
+\bibitem{cn2}
+Peter Clark and Tim Niblett. The CN2 induction algorithm. 1988.
+
 \end{thebibliography}

### 2009-12-09 weka/C45.java
More parameters to C4.5

@@ -5,13 +5,14 @@ import weka.core.Instances;
 public class C45 {
 	
 	public static void main(String[] args) throws Exception {
-		if(args.length != 2) {
-			System.err.println("usage: C45 <db user> <db password>");
+		if(args.length != 3) {
+			System.err.println("usage: C45 <db user> <db password> <options>");
 			System.exit(1);
 		}
 		
 		String user = args[0];
 		String password = args[1];
+		String options = args[2];
 		
 		Instances train_data = Utils.getTrainSet(user, password);
 		System.out.println("====================================");
@@ -27,7 +28,7 @@ public class C45 {
 		
 		// run c4.5
 		J48 tree = new J48();
-		tree.setOptions(weka.core.Utils.splitOptions("-S -C 0.3"));
+		tree.setOptions(weka.core.Utils.splitOptions(options));
 		tree.buildClassifier(train_data);
 		
 		System.out.println(tree.toString());

### 2009-12-09 report/final.tex
Experimental evaluation for c4.5

@@
 \usepackage{a4wide}
+\usepackage{amsmath}
 \usepackage{listings}
@@
 
-Various performance metrics for both classification results will be registered:
-error rate, mean absolute error, kappa statistic, etc.
+Performance metrics for both classification results will be registered: error rate and time spent
+building the classifier and classifying all the examples.
 
 \section{Experimental Evaluation} \label{sec:experimental_eval}
+
+This section shows performance results for the various algorithms.
+
+\subsection{C4.5}
+
+C4.5 in Weka provides a number of parameters to tune the algorithm.
+One of these parameters allows the creation of unpruned trees $<$\textbf{-U}$>$.
+For pruned trees we can use the confidence threshold parameter that dictates how
+much pruning the algorithm will do $<$\textbf{-C float}$>$.
+
+We executed the algorithm using different combinations of the parameters described above.
+Table \ref{tbl:results_c45} displays the obtained results. From the table
+we note that once the confidence threshold reaches a certain level ($\approx$0.55) the results
+are very similar for the unpruned classifier and the over-fitting effect starts to get worse,
+where the train data gets very good results ($\approx$11.19 error rate) and the test set gets a
+large error rate ($\approx$24).
+
+The confidence threshold for best test error rates appears to be in the interval 0.05-0.15.
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    -U & 11.18 & 23.74 & 0m24.378s \\ \hline
+    -C 0.05 & 15.67 & 22.86 & 0m24.414s \\ \hline
+    -C 0.1 & 15.22 & 21.69 & 0m25.417s \\ \hline
+    -C 0.15 & 14.85 & 22.25 & 0m25.086s \\ \hline
+    -C 0.2 & 14.52 & 22.05 & 0m24.436s \\ \hline
+    -C 0.25 & 14.40 & 22.08 & 0m24.628s \\ \hline
+    -C 0.3 & 14.17 & 23.26 & 0m24.765s \\ \hline
+    -C 0.35 & 13.84 & 22.42 & 0m25.154s \\ \hline
+    -C 0.4 & 13.68 & 22.49 & 0m24.325s \\ \hline
+    -C 0.45 & 13.26 & 22.43 & 0m25.086s \\ \hline
+    -C 0.5 & 12.95 & 22.51 & 0m24.874s \\ \hline
+    -C 0.55 & 11.19 & 24.94 & 0m25.899s \\ \hline
+    -C 0.6 & 11.19 & 23.94 & 0m26.882s \\ \hline
+    -C 0.7 & 11.19 & 23.94 & 0m26.638s \\ \hline
+    -C 0.8 & 11.19 & 23.94 & 0m26.818s \\ \hline
+    -C 0.9 & 11.19 & 23.94 & 0m26.634s \\ \hline
+  \end{tabular}
+  \caption{Results for C4.5.}
+  \label{tbl:results_c45}
+  \end{center}
+\end{table}
 

### 2009-12-09 report/final.tex
added preliminary version of rapidminer's description.

@@
 \subsection{Weka}
+\label{sec:weka}
 
@@
 building the classifier and classifying all the examples.
+
+\subsection{RapidMiner}
+
+RapidMiner is an environment for machine learning and data mining experiments 
+\cite{4}. Experiments are done by composing a process made up of a large amount
+of nestable operators. This kind of approach makes it very suitable to create
+and modify different kinds of experiments in a quick fashion.
+
+RapidMiner provides both a GUI to design the process pipeline and a command-line
+tool to load XML-based configuration files. When using the GUI, users can easily
+switch between the pipeline visualization and the corresponding XML file,
+allowing for a better understanding of the configuration file internals. A large
+number of different outputs can be configured on a RapidMiner's process 
+pipeline, providing a broad range of visualizations for the final results.
+RapidMiner also integrates learning schemes and attribute evaluators from the 
+Weka \cite{weka} learning environment (see section \ref{sec:weka} for more 
+details) and provides a programming API for use in different environments.
+
+This tool was used to apply the ID3 \cite{id3}\cite{1} algorithm, mainly because
+of its support for numerical attributes (not seen in Weka), wich enabled a
+better comparison of approaches.
 

### 2009-12-09 report/final.tex
added preliminary CN2 experimental context description.

@@
 
+\subsection{CN2}
+
+CN2 \cite{cn2} is a learning algorithm for rule induction \cite{rule_induction}
+that isn't available in Weka nor in RapidMiner. CN2 inductively learns a set of
+propositional if...then.. rules from a set of training examples, performing a
+general-to-specific beam search through the rule space for the "best" rule, 
+removing training examples covered by it, and repeating it until no more "good"
+rules can be found. The original version of the algorithm used entropy and
+significance test to decide the best rule. Later, it was improved to use the
+Laplace estimate \cite{cn2_laplace}, which is implemented in Peter Clark's 
+software used for our testing purposes.
+
+Peter Clark's CN2 software is built in C and provides a command-line environment
+for conducting experiments. The input format is limited to two files, one
+defining the attributes and other defining the training examples. The software 
+also provides an environment for performance evaluation of the produced set of 
+rules, by accepting a test file with examples.
+
+Only the following values are feed into the algorithms: age, workclass, 
+education, marital\_status, occupation, relationship, race, sex, capital\_gain,
+capital\_loss, hours\_per\_week, native\_country, and plus\_50. The attributes
+education\_num and fnlwgt are derived from other attributes and so are kept out.
+In terms of data pre-processing for missing values, continuous values were 
+replaced by the average value, while discrete values were replaced by the mode.
+
+In terms of performance metris, we took into account the error rate and the time
+spent building the rule set.
+
 \section{Experimental Evaluation} \label{sec:experimental_eval}
@@
 
+\bibitem{cn2_laplace}
+Peter Clark and Robin Boswell. Rule induction with CN2: Some recent improvements. In Y. Kodratoff, editor, Machine Learning - EWSL-91, pages 151-163, Berlin, 1991. Springer-Verlag.
+
 \end{thebibliography}

### 2009-12-09 report/final.tex
added preliminary description of aleph.

@@
 
-In terms of performance metris, we took into account the error rate and the time
-spent building the rule set.
+In terms of performance metrics, we took into account the error rate and the 
+time spent building the rule set.
+
+\subsection{Aleph}
+
+Aleph \cite{aleph} is currently one of the most widely used ILP \cite{ilp} 
+systems. Developed by Ashwin Srinivasan is, as of 2007, in its 5th version. The
+basic Aleph algorithm (used on our experiments) consists of 4 steps:
+
+\begin{enumerate}
+    \item{Select example}
+    \item{Build most-specific clause}
+    \item{Search}
+    \item{Remove redundant}
+\end{enumerate}
+
+Aleph uses inverse entailment to build the most-specific clause and its basic
+algorithm uses a best-first search at the search step. Advanced use of Aleph 
+allows tweaks in every step of the algorithm, namely:
+
+\begin{itemize}
+    \item{\textbf{Select example}: The size of the sample can be configured.}
+    \item{\textbf{Build most-specific clause}: The approach on building the 
+    bottom-clause can be configured. The evaluation mechanism can also be 
+    tweaked.}
+    \item{\textbf{Search}: A huge range of search strategies, evaluation 
+    functions and refinement operators can be applied.}
+    \item{\textbf{Remove redundant}: A strategy of retaining covered examples
+    can be applied, for a better estimation of clause scores.}
+\end{itemize}
+
+Aleph is built entirely in prolog and is best suited to run on the YAP Prolog 
+engine.
+
+In order to use ILP to build a classifier for our data set, some data
+pre-processing needs to be run, namely on converting the examples to logic 
+predicates. Despite the fact that Aleph is able to deal with numerical values,
+discretization has been run on the variables ``age'', ``capital\_gain'', 
+``capital\_loss'' and ``hours\_per\_week''. The value distribution was the same
+as described in section \ref{sec:weka}. The main classification clause is given
+by \verb$more_than_50K(+person).$ Those persons winning more than 50K per year
+are the positive examples, the rest are the negative examples. Each of the
+attributes was converted to unary prolog predicates (for example 
+\verb$young(+person).$, \verb$work_private(+person).$, \verb$male(+person).$,
+\verb$is_native_of_portugal(+person).$). Missing values were simply ommited from
+the background knowledge.
+
+In terms of performance metrics, we took into account the error rate and the 
+time spent building the hypothesis set. The expressiveness of the results is
+also analyzed in section \ref{sec:experimental_eval}.
 
@@
 
+\bibitem{ilp}
+Nada Lavrac and Saso Dzeroski. Inductive Logic Programming: Techniques and Applications. Ellis Horwood, New York, 1994.
+
 \end{thebibliography}

### 2009-12-09 weka/run_c45.sh
script to run c45 tests

@@ -1,0 +1,28 @@
+#!/bin/sh
+
+run_c45()
+{
+  OPTIONS=$1
+  FILE=$2
+  echo "C45 $OPTIONS"
+  { time java C45 root nutshell111 "$OPTIONS"; } &> C45/$FILE.txt
+}
+
+mkdir -p C45
+run_c45 "-U" unpruned
+run_c45 "-C 0.05" 005
+run_c45 "-C 0.1" 010
+run_c45 "-C 0.15" 015
+run_c45 "-C 0.2" 020
+run_c45 "-C 0.25" 025
+run_c45 "-C 0.3" 030
+run_c45 "-C 0.35" 035
+run_c45 "-C 0.4" 040
+run_c45 "-C 0.45" 045
+run_c45 "-C 0.5" 050
+run_c45 "-C 0.55" 055
+run_c45 "-C 0.6" 060
+run_c45 "-C 0.7" 070
+run_c45 "-C 0.8" 080
+run_c45 "-C 0.9" 090
+run_c45 "-C 0.95" 095

### 2009-12-09 report/final.tex
nr of leaves and nodes for results.

@@
   \begin{center}
-  \begin{tabular}{ | l | c | c | c |}
+  \begin{tabular}{ | l | c | c | c | c | c |}
     \hline
-    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
-    -U & 11.18 & 23.74 & 0m24.378s \\ \hline
-    -C 0.05 & 15.67 & 22.86 & 0m24.414s \\ \hline
-    -C 0.1 & 15.22 & 21.69 & 0m25.417s \\ \hline
-    -C 0.15 & 14.85 & 22.25 & 0m25.086s \\ \hline
-    -C 0.2 & 14.52 & 22.05 & 0m24.436s \\ \hline
-    -C 0.25 & 14.40 & 22.08 & 0m24.628s \\ \hline
-    -C 0.3 & 14.17 & 23.26 & 0m24.765s \\ \hline
-    -C 0.35 & 13.84 & 22.42 & 0m25.154s \\ \hline
-    -C 0.4 & 13.68 & 22.49 & 0m24.325s \\ \hline
-    -C 0.45 & 13.26 & 22.43 & 0m25.086s \\ \hline
-    -C 0.5 & 12.95 & 22.51 & 0m24.874s \\ \hline
-    -C 0.55 & 11.19 & 24.94 & 0m25.899s \\ \hline
-    -C 0.6 & 11.19 & 23.94 & 0m26.882s \\ \hline
-    -C 0.7 & 11.19 & 23.94 & 0m26.638s \\ \hline
-    -C 0.8 & 11.19 & 23.94 & 0m26.818s \\ \hline
-    -C 0.9 & 11.19 & 23.94 & 0m26.634s \\ \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
+    -U & 11.18 & 23.74 & 0m24.378s & 6919 & 6198 \\ \hline
+    -C 0.05 & 15.67 & 22.86 & 0m24.414s & 46 & 38 \\ \hline
+    -C 0.1 & 15.22 & 21.69 & 0m25.417s & 106 & 89 \\ \hline
+    -C 0.15 & 14.85 & 22.25 & 0m25.086s & 186 & 158 \\ \hline
+    -C 0.2 & 14.52 & 22.05 & 0m24.436s & 298 & 254 \\ \hline
+    -C 0.25 & 14.40 & 22.08 & 0m24.628s & 376 & 319 \\ \hline
+    -C 0.3 & 14.17 & 23.26 & 0m24.765s & 494 & 423 \\ \hline
+    -C 0.35 & 13.84 & 22.42 & 0m25.154s & 692 & 593 \\ \hline
+    -C 0.4 & 13.68 & 22.49 & 0m24.325s & 835 & 721 \\ \hline
+    -C 0.45 & 13.26 & 22.43 & 0m25.086s & 1143 & 992 \\ \hline
+    -C 0.5 & 12.95 & 22.51 & 0m24.874s & 1483 & 1292 \\ \hline
+    -C 0.55 & 11.19 & 24.94 & 0m25.899s & 6727 & 6036 \\ \hline
+    -C 0.6 & 11.19 & 23.94 & 0m26.882s & 6727 & 6036 \\ \hline
+    -C 0.7 & 11.19 & 23.94 & 0m26.638s & 6727 & 6036 \\ \hline
+    -C 0.8 & 11.19 & 23.94 & 0m26.818s & 6727 & 6036 \\ \hline
+    -C 0.9 & 11.19 & 23.94 & 0m26.634s & 6727 & 6036 \\ \hline
   \end{tabular}

### 2009-12-09 report/final.tex
paste decision tree by C4.5

@@
 
+A decision tree built by this algorithm with the confidence threshold set to 0.05 is shown in Listing \ref{decision_tree_c45}.
+
 \section{Related Work}
@@
 
+\clearpage
+\section{Decision trees}
+
+\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Decision tree built by C4.5 with -C 0.05},label=decision_tree_c45]
+capital_gain = none
+|   marital_status = never_married: <50K (10228.0/340.0)
+|   marital_status = married_civ_spouse
+|   |   education = bachelors
+|   |   |   capital_loss = none
+|   |   |   |   age = young: <50K (33.0/4.0)
+|   |   |   |   age = middle_aged: >50K (1287.0/515.0)
+|   |   |   |   age = senior: >50K (707.0/243.0)
+|   |   |   |   age = old: <50K (77.0/25.0)
+|   |   |   capital_loss = low: >50K (0.0)
+|   |   |   capital_loss = medium: >50K (222.0/26.0)
+|   |   education = hs_grad
+|   |   |   capital_loss = none: <50K (4167.0/1130.0)
+|   |   |   capital_loss = low: <50K (1.0)
+|   |   |   capital_loss = medium: >50K (235.0/105.0)
+|   |   education = 11th: <50K (328.0/29.0)
+|   |   education = masters: >50K (823.0/214.0)
+|   |   education = 9th: <50K (212.0/18.0)
+|   |   education = some_college: <50K (2508.0/1000.0)
+|   |   education = assoc_acdm: <50K (412.0/186.0)
+|   |   education = assoc_voc: <50K (596.0/251.0)
+|   |   education = 7th_8th: <50K (325.0/29.0)
+|   |   education = doctorate: >50K (230.0/42.0)
+|   |   education = prof_school: >50K (295.0/62.0)
+|   |   education = 5th_6th: <50K (161.0/11.0)
+|   |   education = 10th
+|   |   |   capital_loss = none: <50K (314.0/41.0)
+|   |   |   capital_loss = low: <50K (0.0)
+|   |   |   capital_loss = medium: >50K (9.0/2.0)
+|   |   education = 1st_4th: <50K (78.0/5.0)
+|   |   education = preschool: <50K (18.0)
+|   |   education = 12th: <50K (119.0/21.0)
+|   marital_status = divorced: <50K (4155.0/333.0)
+|   marital_status = married_spouse_absent: <50K (397.0/25.0)
+|   marital_status = separated: <50K (973.0/46.0)
+|   marital_status = married_af_spouse
+|   |   hours_per_week = part_time: <50K (2.0)
+|   |   hours_per_week = full_time: <50K (3.0)
+|   |   hours_per_week = over_time: >50K (13.0/4.0)
+|   |   hours_per_week = workaholic: <50K (3.0)
+|   marital_status = widowed: <50K (918.0/62.0)
+capital_gain = low: <50K (55.0)
+capital_gain = medium: <50K (1009.0/181.0)
+capital_gain = high: >50K (878.0/138.0)
+capital_gain = very_high: >50K (770.0/14.0)
+\end{lstlisting}
+
 \end{document}

### 2009-12-09 weka/run_c45.sh
oops

@@ -1,11 +1,14 @@
 #!/bin/sh
 
+USER=$1
+PASSWORD=$2
+
 run_c45()
 {
   OPTIONS=$1
   FILE=$2
   echo "C45 $OPTIONS"
-  { time java C45 root nutshell111 "$OPTIONS"; } &> C45/$FILE.txt
+  { time java C45 $USER $PASSWORD "$OPTIONS"; } &> C45/$FILE.txt
 }
 
 mkdir -p C45

### 2009-12-09 weka/run_adtree.sh
script to run adtree in batch

@@ -1,0 +1,40 @@
+#!/bin/sh
+
+USER=$1
+PASSWORD=$2
+DIR=ADTree
+
+run_adtree()
+{
+  TYPE=$1
+  NAME=$2
+  ITER=$3
+
+  OPTIONS="-E $TYPE -B $ITER"
+  echo "ADTree $OPTIONS"
+  { time java ADTree $USER $PASSWORD "$OPTIONS"; } &> $DIR/$NAME$ITER.txt
+}
+
+run_expand()
+{
+  TYPE=$1
+  NAME=$2
+
+  run_adtree $TYPE $NAME 2
+  run_adtree $TYPE $NAME 5
+  run_adtree $TYPE $NAME 10
+  run_adtree $TYPE $NAME 15
+  run_adtree $TYPE $NAME 20
+  run_adtree $TYPE $NAME 25
+  run_adtree $TYPE $NAME 30
+  run_adtree $TYPE $NAME 35
+  run_adtree $TYPE $NAME 40
+  run_adtree $TYPE $NAME 45
+}
+
+mkdir -p $DIR
+run_expand "-3" "all"
+run_expand "-2" "weight"
+run_expand "-1" "z_pure"
+run_expand "0" "random"
+

### 2009-12-09 report/final.tex
report adtree

@@
 
+\subsection{Alternating Decision Trees}
+
+Alternating Decision Trees are related to a concept named \textit{boosting}.
+Boosting is based on the observation that finding many rough rules of thumb can be a lot easier 
+than finding a single, highly accurate prediction rule. To apply the boosting approach,
+we start with a method or algorithm for finding the rough rules of thumb. 
+The boosting algorithm calls this “weak” or “base” learning algorithm repeatedly, 
+each time feeding it a different subset of the training examples. Each time 
+it is called, the base learning algorithm generates a new weak prediction rule, and 
+after many rounds, the boosting algorithm must combine these weak rules into a 
+single prediction rule that, hopefully, will be much more accurate than any one of 
+the weak rules. \cite{boosting}
+
+The ADTree's Weka algorithm can be parametrized by the number of boosting iterations
+\textbf{$<$-B integer$>$} and by the way the algorithm will expand the nodes,
+\textbf{$<$-E [-3,-2,-1,$>=$0]$>$},
+-3 for \textit{all}, -2 for \textit{weight} based expanding,
+-1 for \textit{z\_pure} and a number
+equal or greater than zero as a seed for random walks.
+
+We experimented the algorithm using various combinations of those two parameters.
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
+    -E -3 -B 2 & 21.80 & 21.44 & 0m24.641s & 7 & 5 \\ \hline
+    -E -3 -B 5 & 18.20 & 20.99 & 0m24.571s & 16 & 11 \\ \hline
+    -E -3 -B 10 & 16.13 & 20.40 & 0m26.539s & 31 & 21 \\ \hline
+    -E -3 -B 15 & 15.91 & 20.19 & 0m28.565s & 46 & 31 \\ \hline
+    -E -3 -B 20 & 15.25 & 20.39 & 0m32.336s & 61 & 41 \\ \hline
+    -E -3 -B 25 & 15.08 & 20.37 & 0m37.223s & 76 & 51 \\ \hline
+    -E -3 -B 30 & 14.88 & 20.25 & 0m41.062s & 91 & 61 \\ \hline
+    -E -3 -B 35 & 14.78 & 20.01 & 0m45.760s & 106 & 71 \\ \hline
+    -E -3 -B 40 & 14.68 & 20.58 & 0m52.976s & 121 & 81 \\ \hline
+
+    -E -2 -B 2 & 21.80 & 21.44 & 0m25.219s & 7 & 5 \\ \hline
+    -E -2 -B 5 & 18.20 & 20.99 & 0m24.445s & 16 & 11 \\ \hline
+    -E -2 -B 10 & 16.34 & 20.71 & 0m25.390s & 31 & 21 \\ \hline
+    -E -2 -B 15 & 15.80 & 20.42 & 0m26.982s & 46 & 31 \\ \hline
+    -E -2 -B 20 & 15.13 & 19.80 & 0m29.050s & 58 & 39 \\ \hline
+    -E -2 -B 25 & 14.95 & 20.25 & 0m28.956s & 73 & 49 \\ \hline
+    -E -2 -B 30 & 14.93 & 20.07 & 0m31.118s & 88 & 59 \\ \hline
+    -E -2 -B 35 & 14.90 & 20.20 & 0m32.565s & 100 & 67 \\ \hline
+    -E -2 -B 40 & 14.65 & 20.73 & 0m34.003s & 115 & 77 \\ \hline
+
+    -E -1 -B 2 & 21.80 & 21.436 & 0m24.149s & 7 & 5 \\ \hline
+    -E -1 -B 5 & 18.20 & 20.99 & 0m24.061s & 16 & 11 \\ \hline
+    -E -1 -B 10 & 16.34 & 20.71 & 0m24.967s & 31 & 21 \\ \hline
+    -E -1 -B 15 & 15.80 & 20.42 & 0m27.095s & 46 & 31 \\ \hline
+    -E -1 -B 20 & 15.13 & 19.80 & 0m28.813s & 58 & 38 \\ \hline
+    -E -1 -B 25 & 14.95 & 20.25 & 0m29.205s & 73 & 49 \\ \hline
+    -E -1 -B 30 & 14.93 & 20.07 & 0m30.569s & 88 & 59 \\ \hline
+    -E -1 -B 35 & 14.93 & 20.20 & 0m32.708s & 100 & 67 \\ \hline
+    -E -1 -B 40 & 14.65 & 20.73 & 0m33.205s & 115 & 77 \\ \hline
+
+    -E 0 -B 2 & 21.80 & 21.44 & 0m23.717s & 7 & 4 \\ \hline
+    -E 0 -B 5 & 18.21 & 20.99 & 0m24.619s & 16 & 11 \\ \hline
+    -E 0 -B 10 & 16.37 & 20.71 & 0m24.512s & 31 & 21 \\ \hline
+    -E 0 -B 15 & 15.81 & 20.42 & 0m25.240s & 46 & 31 \\ \hline
+    -E 0 -B 20 & 14.94 & 20.07 & 0m25.589s & 61 & 41 \\ \hline
+    -E 0 -B 25 & 15.04 & 20.15 & 0m25.868s & 76 & 51 \\ \hline
+    -E 0 -B 30 & 15.01 & 20.50 & 0m26.492s & 91 & 61 \\ \hline
+    -E 0 -B 35 & 14.96 & 20.53 & 0m26.294s & 106 & 71 \\ \hline
+    -E 0 -B 40 & 14.90 & 20.54 & 0m27.248s & 118 & 79 \\ \hline
+
+  \end{tabular}
+  \caption{Results for ADTree.}
+  \label{tbl:results_adtree}
+  \end{center}
+\end{table}
+
+From Table \ref{tbl:results_adtree} we can see that the number of boosting iterations affects
+both the execution time (directly proportional) and the error rate.
+The best error rate is for the parameters \textbf{-E -2 -B 20} giving $\approx$19.80\% for the test set.
+
+Another good observation is that this algorithm can build small trees that give satisfactory results, both for
+the train set and for the test set.
+
 \section{Related Work}
@@
 
+\bibitem{boosting}
+Robert E. Schapire, The Boosting Approach to Machine Learning: An Overview, 2001.
+
 \end{thebibliography}

### 2009-12-09 weka/run_rf.sh
add run forest script

@@ -1,0 +1,33 @@
+#!/bin/sh
+
+USER=$1
+PASSWORD=$2
+DIR=RandomForest
+
+run_rf()
+{
+  NTREES=$1
+  FILE=$NTREES
+
+  OPTIONS="-I $NTREES"
+  echo "RandomForest $OPTIONS"
+
+  { time java -Xmx1G RandomForest $USER $PASSWORD "$OPTIONS"; } &> $DIR/$FILE.txt
+}
+
+mkdir -p $DIR
+#run_rf 2
+#run_rf 3
+#run_rf 4
+#run_rf 5
+#run_rf 7
+#run_rf 10
+#run_rf 15
+#run_rf 20
+#run_rf 40
+#run_rf 60
+#run_rf 80
+#run_rf 100
+#run_rf 150
+#run_rf 200
+run_rf 400

### 2009-12-09 report/final.tex

@@
 
+\subsection{Random Forest}
+
+In Weka, the Random Forest algorithm can be parametrized by the number
+of trees to build \textbf{$<$-I integer$>$}. Table \ref{tbl:results_random_forest} shows results for
+a varying number of trees.
+
+From the table we conclude that we need a certain amount of trees ($\approx$80) to
+reach a satisfactory error rate ($<$22\%) for the test set. Interestingly
+the error rate for the train set is low ($\approx$9\%), which indicates over-fitting.
+
+One drawback of using a big number of trees is the use of large
+quantities of memory (we needed 1GB to compute with 400 trees) and the time needed to compute everything.
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    -I 2 & 10.52 & 34.03 & 0m25.861s \\ \hline
+    -I 3 & 9.83 & 34.00 & 0m26.318s \\ \hline
+    -I 4 & 9.69 & 30.91 & 0m26.154s \\ \hline
+    -I 5 & 9.50 & 29.62 & 0m27.108s \\ \hline
+    -I 7 & 9.32 & 28.03 & 0m27.158s \\ \hline
+    -I 10 & 9.24 & 29.80 & 0m28.107s \\ \hline
+    -I 15 & 9.18 & 28.80 & 0m30.716s \\ \hline
+    -I 20 & 9.15 & 28.09 & 0m36.340s \\ \hline
+    -I 40 & 9.15 & 23.43 & 0m39.687s \\ \hline
+    -I 60 & 9.15 & 23.65 & 0m49.443s \\ \hline
+    -I 80 & 9.15 & 22.92 & 0m57.652s \\ \hline
+    -I 100 & 9.15 & 22.83 & 1m28.024s \\ \hline
+    -I 400 & 9.15 & 22.71 & 4m58.842s \\ \hline
+  \end{tabular}
+  \caption{Results for Random Forest.}
+  \label{tbl:results_random_forest}
+  \end{center}
+\end{table}
+
+\subsection{CART}
+
+The simplified version of the CART algorithm is available
+in the \textit{SimpleCart} Weka class.
+
+This algorithm allows the disabling of various optimizations
+and heuristics \textbf{$<$-U and -H $>$} as explained by the Weka manual.
+We will keep these optimizations in.
+
+Other interesting parameters that we will be changing, include:
+
+\begin{itemize}
+  \item \textbf{$<$-M int$>$}: to set the minimal number of instances at the terminal nodes.
+  \item \textbf{$<$-N int$>$}: to set the number of folds used in the minimal cost-complexity pruning.
+\end{itemize}
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
+    -M 2 -N 5 & 13.96 & 25.37 & 3m6.315s & 119 & 60 \\ \hline
+    -M 2 -N 10 & 13.85 & 25.38 & 5m36.925s & 139 & 70 \\ \hline
+    -M 2 -N 15 & 13.93 & 25.41 & 8m1.214s & 123 & 62 \\ \hline
+
+  \end{tabular}
+  \caption{Results for CART.}
+  \label{tbl:results_cart}
+  \end{center}
+\end{table}
+
+CART results are displayed in Table \ref{tbl:results_cart}.
+
 \section{Related Work}

### 2009-12-09 report/final.tex
experimenta evaluation of CART.

@@
     -M 2 -N 15 & 13.93 & 25.41 & 8m1.214s & 123 & 62 \\ \hline
+    -M 2 -N 20 & 13.88 & 25.39 & 10m27.982s & 133 & 67 \\ \hline
 
+    -M 3 -N 5 & 13.96 & 25.37 & 2m51.491s & 119 & 60 \\ \hline
+    -M 3 -N 10 & 13.85 & 25.39 & 5m18.642s & 139 & 70 \\ \hline
+    -M 3 -N 15 & 13.94 & 25.41 & 7m53.984s & 123 & 62 \\ \hline
+    -M 3 -N 20 & 13.88 & 25.39 & 10m19.665s & 133 & 67 \\ \hline
+
+    -M 5 -N 5 & 13.94 & 25.41 & 2m43.363s & 123 & 62 \\ \hline
+    -M 5 -N 10 & 13.82 & 25.34 & 4m54.547s & 145 & 73 \\ \hline
+    -M 5 -N 15 & 13.94 & 25.41 & 7m9.537s & 123 & 62 \\ \hline
+    -M 5 -N 20 & 13.88 & 25.39 & 9m18.095s & 133 & 67 \\ \hline
   \end{tabular}
@@
 CART results are displayed in Table \ref{tbl:results_cart}.
+From these results we can see that there isn't a lot of variation in the results.
+The algorithm also performs badly when compared to C4.5 or ADTree.
 

### 2009-12-09 report/final.tex
initial description of cn2 experimental results.

--- 2009-12-09 report/final.tex
@@
 
+\subsection{CN2}
+
+The CN2 software used supports different algorithms for building either ordered
+or unordered rule sets and with different error estimations: laplacian, naive,
+information gain and modified entropy. We've tested each combination of
+parameters and the overall results are presented in table \ref{tbl:results_cn2}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l | }
+    \hline
+    \textbf{Algorithm} & \textbf{Error Estimate} & \textbf{Train Error Rate} & \textbf{Test Error Rate} & \textbf{Time}
+\end{tabular}
+\caption{Results for CN2.}
+\label{tbl:results_cn2}
+\end{center}
+\end{table}
+
 \section{Related Work}

### 2009-12-10 report/final.tex
fixing more cn2 data pre-processing scripts.

@@
 \begin{center}
-\begin{tabular}{ | l | l | l | l | l | }
+\begin{tabular}{ | l | l | l | l | l |}
     \hline
-    \textbf{Algorithm} & \textbf{Error Estimate} & \textbf{Train Error Rate} & \textbf{Test Error Rate} & \textbf{Time}
+    \textbf{Algorithm} & \textbf{Error estimate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    Unordered & Laplacian & 13.3 & 15.2 & 3min29sec \\ \hline
+    Unordered & Naive & - & - & - \\ \hline
+    Ordered & Laplacian & 0 & 17.6 & 1min45sec \\ \hline
+    Ordered & Naive &
 \end{tabular}

### 2009-12-10 generate_examples.py

@@ -48,8 +48,8 @@ def convert_line(line, outputfile):
         workclass = 'private'
     if occupation == '?':
         occupation = 'prof_specialty'
-    outputfile.write(age + " " + workclass + " " + fnlwgt + " " + education +
-                     " " + education_num + " " + marital_status + " " + 
+    outputfile.write(age + " " + workclass + " " + education +
+                     " " + marital_status + " " + 
                      occupation + " " + relationship + " " + race + " " + sex +
                      " " + capital_gain + " " + capital_loss + " " + 
                      hours_per_week + " " + native_country + " " + plus_50 +

### 2009-12-10 report/final.tex
more bayesian networks

--- 2009-12-09 report/final.tex
@@
 
-The experiments conducted will use genetic based search, hill climbing \cite{bayes3}, simulated annealing,
-and a modified hill climbing algorithm, named K2 \cite{bayes2}.
+The experiments conducted will use genetic based search, hill climbing \cite{bayes3}, and simulated annealing.
 
@@
 
-This section shows performance results for the various algorithms.
+This section shows performance results for the various algorithms described in section \ref{sec:algs}.
 
@@
 reach a satisfactory error rate ($<$22\%) for the test set. Interestingly
-the error rate for the train set is low ($\approx$9\%), which indicates over-fitting.
+the error rate for the train set is low ($\approx$9\%), which indicates over-fitting
+when comparing the difference to the test error rate.
 
 One drawback of using a big number of trees is the use of large
-quantities of memory (we needed 1GB to compute with 400 trees) and the time needed to compute everything.
+quantities of memory (we needed 1GB to compute 400 trees) and the long time needed to construct the classifier.
 
@@
 
-\subsection{CN2}
+\subsection{Bayesian Networks}
 
-The CN2 software used supports different algorithms for building either ordered
-or unordered rule sets and with different error estimations: laplacian, naive,
-information gain and modified entropy. We've tested each combination of
-parameters and the overall results are presented in table \ref{tbl:results_cn2}.
+Our strategy for bayesian networks revolve around changing the search algorithm
+and its parameters. Weka easily allows the parameterization of this algorithm (\textbf{$<$-Q$>$})
+that optimizes the network structure.
 
-\begin{table}
+The search algorithms experimented include:
+
+\begin{itemize}
+  \item Genetic based search (\textit{weka.classifiers.bayes.net.search.local.GeneticSearch}) \\
+  The following options are available:
+  \begin{itemize}
+    \item -L integer: Population size. 
+    \item -A integer: Descendant population size.
+    \item -U integer: Number of runs/generations.
+    \item -M: Use mutation (true by default).
+    \item -C: Use cross-over (true by default).
+  \end{itemize}
+
+  \item Hill climbing (\textit{weka.classifiers.bayes.net.search.local.HillClimber}) \\
+  The following options are available:
+  \begin{itemize}
+    \item -P $<$nr of parents$>$: Maximum number of parents.
+    \item -R: Use arc reversal operation. (default false)
+  \end{itemize}
+
+  \item Simulated annealing (\textit{weka.classifiers.bayes.net.search.local.SimulatedAnnealing}) \\
+  The following options are available:
+  \begin{itemize}
+    \item -A float: Start temperature.
+    \item -U integer: Number of runs.
+    \item -D float: Delta temperature.
+  \end{itemize}
+\end{itemize}
+
+Weka also allows the parameterization of the estimator algorithm (\textbf{$<$-E$>$}).
+We will use the \textit{SimpleEstimator} estimator (\textit{weka.classifiers.bayes.net.estimate.SimpleEstimator -- -A 1.0})
+for all the experiments.
+
+\subsubsection{Genetic Search}
+
+The Genetic Search algorithm is a genetic based algorithm to optimize the bayesian network structure.
+We can configure the population size and number of generations. Mutation and cross-over can also
+be set or kept out.
+
+For this experiment we will only use a portion of the train data to speed up the process.
+
+\begin{table}[ht]
   \begin{center}
-\begin{tabular}{ | l | l | l | l | l | }
+  \begin{tabular}{ | l | c | c | c | c | c | c |}
     \hline
-    \textbf{Algorithm} & \textbf{Error Estimate} & \textbf{Train Error Rate} & \textbf{Test Error Rate} & \textbf{Time}
+    \textbf{Parameter} & \textbf{Portion} & \textbf{Portion error rate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    GeneticSearch -C -M -L 3 -A 3 -U 3 & 5 & & & & \\ \hline
+    GeneticSearch -C -M -L 3 -A 3 -U 5 & 5 & & & & \\ \hline
+    GeneticSearch -C -M -L 3 -A 3 -U 15 & 5 & & & & \\ \hline
   \end{tabular}
-\caption{Results for CN2.}
-\label{tbl:results_cn2}
+  \caption{Results for Bayesian Networks / Genetic Search.}
+  \label{tbl:results_bayesian_networks_gs}
   \end{center}
 \end{table}
+
+\subsubsection{Hill Climbing}
+
+The best configuration for hill climbing (Table \ref{tbl:results_bayesian_networks_hc})
+is when the number of parents is exactly 1, still, the error rate is somewhat unsatisfactory when
+comparing to other results.
+
+Incrementing the number of parents forces the classifier to over-fit the train set, as it
+directly increases the test set error rate while decreasing the train set error rate.
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    HillClimber -P 1 & 19.22 & 22.93 & 0m29.512s \\ \hline
+    HillClimber -P 2 & 15.15 & 26.34 & 0m30.340s \\ \hline
+    HillClimber -P 3 & 14.51 & 28.11 & 0m29.963s \\ \hline
+    HillClimber -P 4 & 14.43 & 27.92 & 0m29.945s \\ \hline
+    HillClimber -P 10 & 14.43 & 27.92 & 0m30.638s \\ \hline
+  \end{tabular}
+  \caption{Results for Bayesian Networks / Hill Climbing.}
+  \label{tbl:results_bayesian_networks_hc}
+  \end{center}
+\end{table}
+
+\subsubsection{Simulated Annealing}
+
+The simulated annealing algorithm provides lots of configuration parameters
+and it is difficult to input the best values of start temperature, delta
+temperature and number of runs to archive the best error rates.
+
+We tried various temperatures values and came up with the Table \ref{tbl:results_bayesian_networks_sa}.
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    SimulatedAnnealing -U 75 -D 0.4 -A 20 & 14.32 & 26.80 & 0m29.841s \\ \hline
+    SimulatedAnnealing -U 55 -D 0.4 -A 20 & 14.40 & 21.45 & 0m29.77s \\ \hline
+    SimulatedAnnealing -U 52 -D 0.4 -A 20 & 14.49 & 21.09 & 0m29.407s \\ \hline
+    SimulatedAnnealing -U 50 -D 0.4 -A 20 & 15.2 & 21.35 & 0m30.884s \\ \hline
+    SimulatedAnnealing -U 50 -D 0.4 -A 40 & 15.2 & 21.35 & 0m30.884s \\ \hline
+    SimulatedAnnealing -U 25 -D 0.4 -A 20 & 18.25 & 22.31 & 0m29.602s \\ \hline
+    SimulatedAnnealing -U 10 -D 0.4 -A 20 & 19.82 & 22.96 & 0m29.131s \\ \hline
+  \end{tabular}
+  \caption{Results for Bayesian Networks / Simulated Annealing.}
+  \label{tbl:results_bayesian_networks_sa}
+  \end{center}
+\end{table}
+
+The diminish the error rates more research will be required to find the best
+tuple of configuration values. Anyway, with just a few guesses we came up
+with $\approx21$\% error rate, which is quite good when comparing to other algorithms.
 
@@
 Ashwin Srinivasan. The Aleph Manual. 2007
-
-\bibitem{bayes2}
-G. Cooper, E. Herskovits. A Bayesian method for the induction of probabilistic networks from 
-data. Machine Learning, 9: 309–347, 1992.
 

### 2009-12-10 report/final.tex
more CN2 results.

--- 2009-12-09 report/final.tex
@@
 \begin{center}
-\begin{tabular}{ | l | l | l | l | l | }
+\begin{tabular}{ | l | l | l | l | l |}
     \hline
-    \textbf{Algorithm} & \textbf{Error Estimate} & \textbf{Train Error Rate} & \textbf{Test Error Rate} & \textbf{Time}
+    \textbf{Algorithm} & \textbf{Error estimate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    Unordered & Laplacian & 13.3 & 15.2 & 3min29sec \\ \hline
+    Unordered & Naive & - & - & - \\ \hline
+    Ordered & Laplacian & 0 & 17.6 & 1min45sec \\ \hline
+    Ordered & Naive & 0 & 19.9 & 12min21sec \\ \hline
 \end{tabular}

### 2009-12-10 report/final.tex

--- 2009-12-10 report/final.tex
@@
 
+\subsection{CN2}
+
+The CN2 software used supports different algorithms for building either ordered
+or unordered rule sets and with different error estimations: laplacian, naive,
+information gain and modified entropy. We've tested each combination of
+parameters and the overall results are presented in table \ref{tbl:results_cn2}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l |}
+    \hline
+    \textbf{Algorithm} & \textbf{Error estimate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    Unordered & Laplacian & 13.3 & 15.2 & 3min29sec \\ \hline
+    Unordered & Naive & - & - & - \\ \hline
+    Ordered & Laplacian & 0 & 17.6 & 1min45sec \\ \hline
+    Ordered & Naive & 0 & 19.9 & 12min21sec \\ \hline
+\end{tabular}
+\caption{Results for CN2.}
+\label{tbl:results_cn2}
+\end{center}
+\end{table}
+
 \subsection{Bayesian Networks}

### 2009-12-10 report/final.tex
finishing CN2 experimental results description.

@@
     Ordered & Naive & 0 & 19.9 & 12min21sec \\ \hline
+    Ordered & Information Gain & 18.9 & 18.9 & 5sec \\ \hline
+    Ordered & Modified Entropy & 18.9 & 19.1 & 20sec \\ \hline
 \end{tabular}
@@
 \end{table}
+
+The information gain and modified entropy approaches to error estimations are
+only available for the algorithm producing ordered set of rules. The approach
+building an unordered set of rules with a naive error estimation lead to a stack
+overflow error in the application. One thing one can note is that the process of
+building classification rules takes, in average, longer than the one of building
+classification trees. Nevertheless, the results are arguably more expressive and
+the error rates are low (the best being 15.2\% for a previously unkown set of 
+examples). Also note that the variant of building an ordered set of rules with
+an information gain error estimate produces rules faster and with an acceptable
+error rate (18.9\%). To analyze the expressiveness of the rules produced, a set
+of rules that cover the most examples of the training data are described in 
+listing \ref{cn2_rules} (using the ordered algorithm with the laplacian error 
+estimation).
 
@@
 
+\clearpage
+\section{Classification Rules}
+\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Subset of the ordered set of rules induced by CN2 with the laplacian error estimation method},label=cn2_rules]
+IF    age < 20.50
+  AND capital_gain < 9720.00
+  AND hours_per_week < 59.00
+THEN  class = less_or_equal_than_50K  [0 2366]
+ELSE
+IF    19.50 < age < 21.50
+  AND marital_status = never_married
+  AND capital_gain < 67047.00
+THEN  class = less_or_equal_than_50K  [0 664]
+ELSE
+IF    age < 54.50
+  AND capital_gain > 10585.50
+  AND hours_per_week > 33.00
+THEN  class = more_than_50K  [520 0]
+ELSE
+IF    age < 23.50
+  AND hours_per_week < 31.00
+THEN  class = less_or_equal_than_50K  [0 492]
+ELSE
+IF    age < 24.50
+  AND workclass = private
+  AND relationship = own_child
+  AND capital_loss < 2129.50
+  AND hours_per_week > 32.50
+  AND native_country = united_states
+THEN  class = less_or_equal_than_50K  [0 562]
+ELSE
+IF    age < 60.50
+  AND 7565.50 < capital_gain < 30961.50
+THEN  class = more_than_50K  [445 0]
+ELSE
+IF    relationship = own_child
+  AND race = black
+  AND hours_per_week < 49.00
+THEN  class = less_or_equal_than_50K  [0 261]
+ELSE
+IF    age < 38.50
+  AND education = hs_grad
+  AND marital_status = never_married
+  AND relationship = own_child
+  AND hours_per_week < 75.00
+THEN  class = less_or_equal_than_50K  [0 378]
+ELSE
+IF    20.50 < age < 23.50
+  AND workclass = private
+  AND marital_status = never_married
+  AND capital_loss < 2116.00
+  AND hours_per_week > 32.50
+THEN  class = less_or_equal_than_50K  [0 434]
+\end{lstlisting}
+
 \end{document}

### 2009-12-10 report/final.tex
added ILP experimental description.

@@
 \usepackage{listings}
+\usepackage{multirow}
 \usepackage[hyphens]{url}
@@
 
+\subsection{ILP}
+
+There are a lot of techniques for building ILP classifiers. Due to time
+constraints (as building an ILP classifier is usually a very time-consuming 
+task), we only tested the basic Aleph system, which uses inverse entailment for
+building the bottom clause and a best-first search method to find the best
+hypothesis. The results obtained are described in table \ref{tbl:results_ilp}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{ | c | c | c | c | c |}
+\hline
+\vspace{1pt} & \multicolumn{4}{| c |}{\textbf{Actual}} \\ \hline
+\multirow{3}{*}{\textbf{Predicted}} & \vspace{1pt} & \textbf{+} & \textbf{-} & \textbf{Total} \\ \hline
+& \textbf{+} & 647 & 133 & 780 \\ \hline
+& \textbf{-} & 3199 & 12302 & 15501 \\ \hline
+& \textbf{Total} & 3846 & 12435 & 16281 \\ \hline
+\end{tabular}
+\caption{Results for ILP (Aleph).}
+\label{tbl:results_ilp}
+\end{center}
+\end{table}
+
+The error rate produced by Aleph when evaluating the test data was 20.45\%.
+18881882 clauses were built, taking around 5 hours for the whole training
+data. Despite the number of clauses, most of them are irrelevant, directly
+classifying examples where no more general rule is available. Listing 
+\ref{ilp_clauses} lists the most relevant clauses produced.
+
 \subsection{Bayesian Networks}
@@
 \end{lstlisting}
+\clearpage
+
+\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Subset of the prolog clauses produced by Aleph},label=ilp_clauses]
+more_than_50K(A) :-
+   middle_aged(A), very_high_capital_gain(A), is_native_of_united_states(A).
+more_than_50K(A) :-
+   education_prof_school(A), wife(A), workaholic(A).
+more_than_50K(A) :-
+   senior(A), work_private(A), very_high_capital_gain(A).
+more_than_50K(A) :-
+   senior(A), work_federal_gov(A), education_doctorate(A).
+more_than_50K(A) :-
+   education_doctorate(A), medium_capital_loss(A), workaholic(A).
+more_than_50K(A) :-
+   work_self_emp_inc(A), high_capital_gain(A), full_time_worker(A).
+more_than_50K(A) :-
+   middle_aged(A), education_masters(A), is_native_of_iran(A).
+more_than_50K(A) :-
+   senior(A), very_high_capital_gain(A), over_time_worker(A).
+more_than_50K(A) :-
+   other_relative(A), high_capital_gain(A).
+more_than_50K(A) :-
+   education_bachelors(A), occupation_other_service(A), high_capital_gain(A).
+more_than_50K(A) :-
+   senior(A), married_civ_spouse(A), very_high_capital_gain(A).
+more_than_50K(A) :-
+   senior(A), full_time_worker(A), is_native_of_cambodia(A).
+more_than_50K(A) :-
+   middle_aged(A), married_civ_spouse(A), is_native_of_thailand(A).
+more_than_50K(A) :-
+   young(A), high_capital_gain(A), over_time_worker(A).
+more_than_50K(A) :-
+   middle_aged(A), asian_pac_islander(A), high_capital_gain(A).
+more_than_50K(A) :-
+   occupation_protective_serv(A), high_capital_gain(A), full_time_worker(A).
+more_than_50K(A) :-
+   work_state_gov(A), occupation_exec_managerial(A), high_capital_gain(A).
+more_than_50K(A) :-
+   education_bachelors(A), occupation_tech_support(A), wife(A).
+more_than_50K(A) :-
+   education_prof_school(A), very_high_capital_gain(A).
+more_than_50K(A) :-
+   high_capital_gain(A), full_time_worker(A), is_native_of_canada(A).
+more_than_50K(A) :-
+   education_prof_school(A), medium_capital_loss(A), over_time_worker(A).
+more_than_50K(A) :-
+   work_private(A), education_doctorate(A), is_native_of_taiwan(A).
+\end{lstlisting}
 

### 2009-12-10 report/final.tex
added ID3 initial description.

@@
 This tool was used to apply the ID3 \cite{id3}\cite{1} algorithm, mainly because
-of its support for numerical attributes (not seen in Weka), wich enabled a
+of its support for numerical attributes (not seen in Weka), which enabled a
 better comparison of approaches.
+
+The data pre-processing conducted consisted on a missing value replacement by 
+the average on numerical values and by the mode on the nominal/ordinal values. 
+An absolute discretization was applied to the numerical values for testing
+purposes of the default ID3 implementation.
 
@@
 This section shows performance results for the various algorithms described in section \ref{sec:algs}.
+
+\subsection{ID3}
+
+ID3 in RapidMiner is implemented in two different flavours, one that supports
+numerical values, one that doesn't. For the one that doesn't support numerical
+values, an absolute discretization was conducted. Table \ref{tbl:results_id3}
+summarizes the results with the ID3 algorithm.
+
+\begin{table}
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c | c |}
+    \hline
+    \textbf{Type} & \textbf{Criterion} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    ID3 & Gain Ratio & 0 & 0 & 0 \\ \hline
+    ID3 & Information Gain & 0 & 0 & 0 \\ \hline
+    ID3 & Gini Index & 0 & 0 & 0 \\ \hline
+    ID3 & Accuracy & 0 & 0 & 0 \\ \hline
+    ID3 Numerical & Gain Ratio & 0 & 0 & 0 \\ \hline
+    ID3 Numerical & Information Gain & 0 & 0 & 0 \\ \hline
+    ID3 Numerical & Gini Index & 0 & 0 & 0 \\ \hline
+    ID3 Numerical & Accuracy & 0 & 0 & 0 \\ \hline
+  \end{tabular}
+  \caption{Results for ID3.}
+  \label{tbl:results_id3}
+  \end{center}
+\end{table}
 
@@
     \textbf{Algorithm} & \textbf{Error estimate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
-    Unordered & Laplacian & 13.3 & 15.2 & 3min29sec \\ \hline
+    Unordered & Laplacian & 13.3 & 15.2 & 3m29s \\ \hline
     Unordered & Naive & - & - & - \\ \hline
-    Ordered & Laplacian & 0 & 17.6 & 1min45sec \\ \hline
-    Ordered & Naive & 0 & 19.9 & 12min21sec \\ \hline
-    Ordered & Information Gain & 18.9 & 18.9 & 5sec \\ \hline
-    Ordered & Modified Entropy & 18.9 & 19.1 & 20sec \\ \hline
+    Ordered & Laplacian & 0 & 17.6 & 1m45s \\ \hline
+    Ordered & Naive & 0 & 19.9 & 12m21s \\ \hline
+    Ordered & Information Gain & 18.9 & 18.9 & 0m5s \\ \hline
+    Ordered & Modified Entropy & 18.9 & 19.1 & 0m20s \\ \hline
 \end{tabular}

### 2009-12-10 report/final.tex
added id3 results.

@@
     \textbf{Type} & \textbf{Criterion} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
-    ID3 & Gain Ratio & 0 & 0 & 0 \\ \hline
-    ID3 & Information Gain & 0 & 0 & 0 \\ \hline
-    ID3 & Gini Index & 0 & 0 & 0 \\ \hline
-    ID3 & Accuracy & 0 & 0 & 0 \\ \hline
-    ID3 Numerical & Gain Ratio & 0 & 0 & 0 \\ \hline
-    ID3 Numerical & Information Gain & 0 & 0 & 0 \\ \hline
-    ID3 Numerical & Gini Index & 0 & 0 & 0 \\ \hline
-    ID3 Numerical & Accuracy & 0 & 0 & 0 \\ \hline
+    ID3 & Gain Ratio & 18.38 & 18.61 & 0m04s \\ \hline
+    ID3 & Information Gain & 18.48 & 19.05 & 0m03s \\ \hline
+    ID3 & Gini Index & 18.94 & 19.15 & 0m04s \\ \hline
+    ID3 & Accuracy & 20.41 & 20.24 & 0m04s \\ \hline
+    ID3 Numerical & Gain Ratio & 14.93 & 15.68 & 0m14s \\ \hline
+    ID3 Numerical & Information Gain & 15.68 & 16.42 & 0m06s \\ \hline
+    ID3 Numerical & Gini Index & 16.10 & 16.96 & 0m07s \\ \hline
+    ID3 Numerical & Accuracy & 17.99 & 17.92 & 0m12s \\ \hline
   \end{tabular}

### 2009-12-10 report/final.tex
minor fixes on the report.

@@
 
+As can be seen from the table, the numerical ID3 outperforms the standard ID3 in
+accuracy but takes longer to compute the decision tree. The decision tree
+produced is very large, as ID3 doesn't support pruning (see section 
+\ref{sec:4.5}), and so is omitted from this report.
+
 \subsection{C4.5}
+\label{sec:c4.5}
 
@@
 data. Despite the number of clauses, most of them are irrelevant, directly
-classifying examples where no more general rule is available. Listing 
+classifying examples when no more general rule is available. Listing 
 \ref{ilp_clauses} lists the most relevant clauses produced.

### 2009-12-10 report/final.tex
preliminary vesion of the introduction.

@@
 \section{Introduction}
+
+Current real-world databases are rich with information that can be used for
+decision making. Classification is a form of data analysis used to extract 
+models that describe important data classes \cite{2}. Classification is, thus, 
+the process of predicting categorical labels. There has been active research on 
+the fields of machine learning, pattern recognition and statistics regarding 
+classification. Data classification is a two-step process, involving both a 
+learning step and a training step. The learning step builds a model, or 
+classifier, from training data, to which the class is already known (therefore
+this is called supervised learning, as opposed to unsupervised learning used in
+clustering, for example). The training step evaluates the accuracy of the 
+classifier by providing data not used on the learning step. If the accuracy is
+considered acceptable, then the rules can be applied to the classification of
+new data tuples.
+
+The models built can assume different forms, ranging from decision trees to 
+classification rules, including prolog clauses (as in ILP \cite{ilp}). This
+report describes the application of different classification algorithms applied
+to the adult database. It starts by identifying the problem, describing the
+algorithms to be applied, the data format, the experimental context and the 
+experimental evaluation. Some conclusions are then taken on the subject and 
+related work is described.
 

### 2009-12-10 report/final.tex
added abstract.

@@
 \begin{abstract}
-Highly frequent in experimental tests related to data mining, the Adult Database
-has some interesting properties which make it quite useful to test the 
-efficiency of classification algorithms. This report aims to describe the Adult
-Database from a statistical perspective and taking into account its support for
-data mining tasks.
+Classification is the process of predicting categorical labels in data. It has a
+particular interest in a wide range of real world situations, where efficient
+data analysis can provide a higher understanding of data at large. Taking that 
+into account, the adult database, highly frequent in experimental tests related
+to data mining, provides a solid data background to test different types of 
+classification and to obtain interesting and relevant models to real world 
+situations. This report aims to describe different approaches on building 
+classifiers with the adult database and to conclude both on the accuracy of the
+models built and on the properties obtained by them.
 \end{abstract}

### 2009-12-10 report/final.tex
added TODO for conclusions.
added data analysis to the final paper.

@@
 
+\subsection{Data Analysis}
+
+The table \ref{tab:summary} summarizes the distribution of values on the studied
+dataset. From a first observation, one can see that the amount of missing values
+is low. On the other side, in some of the attributes, some classes have a
+much higher frequency than others, such as the 'united-states' native country, 
+the 'private' workclass and the 'white' race. On the hours of work per week
+attribute, there is a great amount of outliers, with some unrealistic
+values, namely with amounts close to the 100 hours or less than 10 in 
+individuals with a defined occupation, as can be seen on figure 
+\ref{fig:box_hours}. Variables such has the capital gain and capital loss have
+few values above 0, being quite specific to certain classes of individuals. The
+sex is also unbalanced in the sample data, as the amount of men is more than
+twice the amount of women.
+
+It is also important to verify some correlations between variables. To that
+matter, we've tested some apparent situations, such as the influence of the age,
+hours of work per week and educational level on the amount gained by an
+individual. The figures \ref{fig:age_plus50}, \ref{fig:education_plus50} and
+\ref{fig:hpw_plus50} show that there is a tendency for older people, people with
+a higher educational level and people who work a great amount of hours per week
+to gain a larger amount of money.
+
+\subsubsection{Association Rules}
+
+Using association rules one can find frequent patterns and associations from a
+set of transactions. These rules can easily be used to make data analysis and
+to understand casual structures.
+
+To build a set of association rules we used the apriori algorithm that is featured
+in the R \cite{6} arules \cite{7} package. Before running the algorithm we made some data preprocessing
+for continuous attributes that are needed to build better and more general rules.
+
+The following attributes were transformed:
+
+\begin{itemize}
+  \item {\textbf{Age}: turned into categories. 15-25 (young), 25-45 (middle\_aged), 45-65 (senior), $>$ 65 (old).}
+  \item {\textbf{Hours-per-week}: turned into categories. 0-25 (part\_time), 25-40 (full\_time), 40-60 (overtime), $>$60 (workaholic).}
+  \item {\textbf{Capital-gain}: turned into categories. $<=$ 0 (none), $<$ median value (low), $>$ median value (high).}
+  \item {\textbf{Capital-loss}: turned into categories. $<=$ 0 (none), $<$ median value (low), $>$ median value (high).}
+\end{itemize}
+
+Next, we plotted the relative frequency of items in the transaction database
+with support greater than 10\%. The results are shown in figure \ref{fig:arules_item_frequency}.
+We can learn various things from that plot: the majority of people are from the united states,
+the white race is predominant, males outnumber females and that more people work in full time.
+
+Finally we run the apriori algorithm using the test data set to generate rules
+with support greater than 2\% and at least 60\% of confidence. The algorithm built
+50487 rules.
+
+We then built two sets of rules from the original set: one with the consequent component
+with plus\_50=0 (the person gains less than 50K a year) and the other with plus\_50=1
+(the person gains more than 50K a year).
+
+Each rule set was sorted by the confidence level, from more confident to less confident.
+
+The first 5 rules from the plus\_50=0 set are shown in listing \ref{arules_small_income} and
+are related to small incomes. We can easily make some observations
+about those rules: young people with a part-time are probably low income and having a child
+without being married is also not a very good signal.
+
+The second set, contrary to the ones above, shows us the top 5 rules for having large incomes.
+These rules are shown in listing \ref{arules_large_income}. A conclusion can readily be made:
+being a white husband with a high capital gain probably means we have a large income.
+
+\subsubsection{Clusters}
+
+Another method to analyze the dataset is based on the generation of clusters. 
+Clusters are sets of data in which objects in the same cluster are similar, and
+objects in different clusters are dissimilar. The first attempt at clustering
+used only three variables: age, education\_num and hours\_per\_week, the same
+variables we saw that had a greater impact on the amount gained by the
+individual. The data was normalized using the Z-Transformation method and the 
+K-Means method was used to obtain the final clusters, with a k of 6, using RapidMiner \cite{4}. Figures 
+\ref{fig:cluster_1_age_education}, \ref{fig:cluster_1_education_hpw} and \ref{fig:cluster_1_age_hpw} show
+scatter plots of the variables, colored by clusters. One thing one is able to
+note is that people with younger age belong to cluster \#2. In terms of
+education, people with higher education belong to cluster \#0, while people with
+lower education belong to cluster \#1. In terms of hours of work per week, 
+distribution is rather similar, but cluster \#2 and \#3 have a greater amount of
+lower and higher values, respectively. Figure \ref{fig:cluster_1_cluster_plus50}
+shows how clusters relate to the class values. Although the distribution is 
+balanced, clusters \#1 and \#2 have a greater amount of people gaining below 
+50000 dollars. Notice, as stated before, that this were the clusters having 
+people with lower education and younger age, respectively.
+
+Clustering with the remaining attributes leads to results more difficult to
+analyze. The same process was applied, using the Z-Transformation method for
+normalization and the K-Means algorithm with a k of 6, but this time with all
+the attributes. The distribution of values is more balanced, although a very
+high value of capital gain, very unusual on the data set, leads to the creation
+of a very well identified cluster, as can be seen on table \ref{tab:centroids}.
+Particular values of race, workclass and native country also identify clusters
+\#2, \#3 and \#2, respectively. Notice that this attributes were the ones
+refered to before as having a given value more frequent than others, thus
+leading to this particular composition of clusters.
+
 \section{Experimental Context}
@@
 produced is very large, as ID3 doesn't support pruning (see section 
-\ref{sec:4.5}), and so is omitted from this report.
+\ref{sec:c4.5}), and so is omitted from this report.
 
@@
 
+// TODO: Tabela e descrição das diferentes taxas de erro
+
+// TODO: Descrição de alguns padrões interessantes que sobressaiam nos modelos
+
 \begin{thebibliography}{1}
@@
 \section{Tables}
+
+\begin{table}
+\centering
+\subtable[Age]{
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 17 \\
+       \hline
+       1st Qu. & 28 \\
+       \hline
+       Median & 37 \\
+       \hline
+       Mean & 38.58 \\
+       \hline
+       3rd Qu. & 48 \\
+       \hline
+       Max & 90 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Workclass]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       private & 22696 \\
+       \hline
+       self\_emp\_not\_inc & 2541 \\
+       \hline
+       local\_gov & 2093 \\
+       \hline
+       state\_gov & 1298 \\
+       \hline
+       self\_emp\_inc & 1116 \\
+       \hline
+       (Other) & 981 \\
+       \hline
+       NA's & 1836 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Education]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       hs\_grad & 10501 \\
+       \hline
+       some\_college & 7291 \\
+       \hline
+       bachelors & 5535 \\
+       \hline
+       masters & 1723 \\
+       \hline
+       assoc\_voc & 1382 \\
+       \hline
+       11th & 1175 \\
+       \hline
+       (Other) & 5134 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Marital Status]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       divorced & 4443 \\
+       \hline
+       married\_af\_spouse & 23 \\
+       \hline
+       married\_civ\_spouse & 14976 \\
+       \hline
+       married\_spouse\_absent & 418 \\
+       \hline
+       never\_married & 10683 \\
+       \hline
+       separated & 1025 \\
+       \hline
+       widowed & 993 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Occupation]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       prof\_specialty & 4140 \\
+       \hline
+       craft\_repair & 4099 \\
+       \hline
+       exec\_managerial & 4066 \\
+       \hline
+       adm\_clerical & 3770 \\
+       \hline
+       sales & 3650 \\
+       \hline
+       (Other) & 10993 \\
+       \hline
+       NA's & 1843 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Relatonship]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       husband & 13193 \\
+       \hline
+       not\_in\_family & 8305 \\
+       \hline
+       other\_relative & 981 \\
+       \hline
+       own\_child & 5068 \\
+       \hline
+       unmarried & 3446 \\
+       \hline
+       wife & 1568 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Race]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       amer\_indian\_eskimo & 311 \\
+       \hline
+       asian\_pac\_islander & 1039 \\
+       \hline
+       black & 3124 \\
+       \hline
+       other & 271 \\
+       \hline
+       white & 27816 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Sex]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       female & 10771 \\
+       \hline
+       male & 21790 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Capital Gain]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 0 \\
+       \hline
+       1st Qu. & 0 \\
+       \hline
+       Median & 0 \\
+       \hline
+       Mean & 1078 \\
+       \hline
+       3rd Qu. & 0 \\
+       \hline
+       Max. & 99999 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Capital Loss]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 0 \\
+       \hline
+       1st Qu. & 0 \\
+       \hline
+       Median & 0 \\
+       \hline
+       Mean & 87.3 \\
+       \hline
+       3rd Qu. & 0 \\
+       \hline
+       Max. & 4356 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Hours Per Week]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 1 \\
+       \hline
+       1st Qu. & 40 \\
+       \hline
+       Median & 40 \\
+       \hline
+       Mean & 40.44 \\
+       \hline
+       3rd Qu. & 45 \\
+       \hline
+       Max. & 99 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Native Country]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       United States & 29170 \\
+       \hline
+       Mexico & 643 \\
+       \hline
+       Philippines & 198 \\
+       \hline
+       Germany & 137 \\
+       \hline
+       Canada & 121 \\
+       \hline
+       (Other) & 1709 \\
+       \hline
+       NA's & 583 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Plus 50]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       No & 24720 \\
+       \hline
+       Yes & 7841 \\
+       \hline
+       \end{tabular}
+}
+\caption{Summary of the value distribution in the database}
+\label{tab:summary}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{ | l | l | l | l | l | l | l | }
+\hline
+\textbf{Attribute} & \textbf{Centroid \#0} & \textbf{Centroid \#1} & \textbf{Centroid \#2} & \textbf{Centroid \#3} & \textbf{Centroid \#4} & \textbf{Centroid \#5} \\
+\hline
+age	& -0.874 & 0.692 & -0.166 & 0.246 & 0.570 & 0.190 \\
+\hline
+workclass & -0.272 & -0.063 & -0.036 & 2.208 & 0.455 & -0.364 \\
+\hline
+education\_num & -0.101 & -0.172 & -0.177 & 0.0309 & 1.103 & 0.102 \\
+\hline
+marital\_status & -0.679 & 1.693 & -0.083 & -0.150 & -0.007 & -0.108 \\
+\hline
+occupation & -0.305 & -0.303 & 0.091 & 0.952 & -0.434 & 0.010 \\
+\hline
+relationship & 0.703 & 0.526 & 0.189 & -0.053 & -0.351 & -0.507 \\
+\hline
+race & -0.086 & -0.020 & 2.630 & -0.147 & -0.002 & -0.230 \\
+\hline
+sex & 0.662 & 1.199 & 0.019 & -0.161 & -0.409 & -0.635 \\
+\hline
+capital\_gain & -0.116 & -0.091 & -0.086 & -0.029 & 13.394 & -0.040 \\
+\hline
+capital\_loss & -0.115 & -0.079 & -0.021 & 0.025 & -0.217 & 0.079 \\
+\hline
+hours\_per\_week & -0.523 & -0.231 & -0.031 &-0.071 & 0.758 & 0.339 \\
+\hline
+native\_country & -0.186 & -0.154 & 2.951 & -0.181 & -0.075 & -0.171 \\
+\hline
+\end{tabular}
+\caption{Centroid table for the results of the application of the K-Means algorithm with a k of 6}
+\label{tab:centroids}
+\end{table}
+
+\clearpage
+\section{Figures}
+
+\begin{figure}
+\centering
+\includegraphics[width=9cm]{boxplot_hours_per_week.png}
+\caption{Boxplot of the hours per week variable}
+\label{fig:box_hours}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=8cm]{plot_plus_50_age.png}
+\caption{Boxplot of the individuals' age, for each class of gain}
+\label{fig:age_plus50}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=9cm]{plot_plus_50_education_num.png}
+\caption{Boxplot of the individuals' education level, for each class of gain}
+\label{fig:education_plus50}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=9cm]{plot_plus_50_hpw.png}
+\caption{Boxplot of the individuals' hours of work per week, for each class of 
+gain}
+\label{fig:hpw_plus50}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{cluster_1_age_education.png}
+\caption{Scatter plot of age and education}
+\label{fig:cluster_1_age_education}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{cluster_1_education_hpw.png}
+\caption{Scatter plot of education and hours of work per week}
+\label{fig:cluster_1_education_hpw}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{cluster_1_age_hpw.png}
+\caption{Scatter plot of age and hours of work per week}
+\label{fig:cluster_1_age_hpw}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=10cm]{cluster_1_cluster_plus50.png}
+\caption{Scatter plot of cluster and class values of amount gained}
+\label{fig:cluster_1_cluster_plus50}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=20cm,angle=90]{item_frequency.png}
+\caption{Item frequencies of items with at least 10\% support.}
+\label{fig:arules_item_frequency}
+\end{figure}
+
+\clearpage
+\section{Association rules}
+
+\begin{lstlisting}[language=r,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Association rules for small income.},label=arules_small_income]
+  lhs                               rhs            support confidence     lift
+1 {age=young,                                                                 
+   hours_per_week=part_time}     => {plus_50=0} 0.05730782          1 1.317193
+2 {age=young,                                                                 
+   occupation=unknown,                                                        
+   capital_gain=none}            => {plus_50=0} 0.02002396          1 1.317193
+3 {age=young,                                                                 
+   relationship=own_child,                                                    
+   hours_per_week=part_time}     => {plus_50=0} 0.04404042          1 1.317193
+4 {education=some_college,                                                    
+   relationship=own_child,                                                    
+   hours_per_week=part_time}     => {plus_50=0} 0.02069961          1 1.317193
+5 {marital_status=never_married,                                              
+   relationship=own_child,                                                    
+   hours_per_week=part_time}     => {plus_50=0} 0.04689659          1 1.317193
+\end{lstlisting}
+
+\begin{lstlisting}[language=r,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Association rules for large income.},label=arules_large_income]
+  lhs                                    rhs            support confidence     lift
+1 {relationship=husband,                                                           
+   race=white,                                                                     
+   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
+2 {marital_status=married_civ_spouse,                                              
+   relationship=husband,                                                           
+   race=white,                                                                     
+   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
+3 {relationship=husband,                                                           
+   race=white,                                                                     
+   sex=male,                                                                       
+   capital_gain=high}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
+4 {relationship=husband,                                                           
+   race=white,                                                                     
+   capital_gain=high,                                                              
+   capital_loss=none}                 => {plus_50=1} 0.02063819  0.9926145 4.121990
+5 {relationship=husband,                                                           
+   capital_gain=high}                 => {plus_50=1} 0.02251159  0.9918809 4.118943
+\end{lstlisting}
 

### 2009-12-10 report/final.tex
related work

--- 2009-12-10 report/final.tex
@@
 
+The adult database has been widely used to benchmark novel learning algorithms.
+
+In the article \textit{Scaling Up the Accuracy of Naive-Bayes Classifiers: a Decision Tree hybrid},
+the author presents the nbtree algorithm, which is an hybrid approach suitable in learning scenarios
+when many attributes are likely to be relevant for a classification task, yet the attributes
+are not necessarily conditionally independent given the class \cite{nbtree}. It builds
+a decision tree with bayesian classifiers at the leaves.
+
+In the paper the author used various datasets, including the adult database. Records with unknown values
+were removed from the experiments. C4.5 reached an error rate of $\approx$15\% and NBTree $\approx$16.5\%,
+which are identical to our ID3 results, but worse than our C4.5 experiments. One reason for this is
+the incapability of Weka to use numeric attributes. Every other algorithm we used fared worse ($\approx$20\%).
+
+Another paper, \textit{Learning Bayesian Belief Network Classifiers: Algorithms and System}, explores
+various kinds of search algorithms used in Bayesian Networks and also comes up with very good error rates ($\approx$ 15\%).
+Other papers using this dataset can be found in \cite{dataset_papers}.
+
+The location from where we downloaded the dataset also contains results
+using various algorithms (Table \ref{tbl:dataset_error_rates}).
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c |}
+    \hline
+    \textbf{Algorithm} & \textbf{Error rate} \\ \hline
+    C4.5 & 15.54 \\ \hline
+    C4.5-auto & 14.46 \\ \hline
+    C4.5 rules & 14.94 \\ \hline
+    Voted ID3 (0.6) & 15.64 \\ \hline
+    Voted ID3 (0.8) & 16.47 \\ \hline
+    T2 & 16.84 \\ \hline
+    1R & 19.54 \\ \hline
+    NBTree & 14.10 \\ \hline
+    CN2 & 16.00 \\ \hline
+    HOODG & 14.82 \\ \hline
+    FSS Naive Bayes & 14.05 \\ \hline
+    IDTM (Decision table) & 14.05 \\ \hline
+    IDTM (Decision table) & 14.46 \\ \hline
+    Naive-Bayes & 16.12 \\ \hline
+    Nearest-neighbor (1) & 21.42 \\ \hline
+    Nearest-neighbor (3) & 20.35 \\ \hline
+    OC1 & 15.04 \\ \hline
+  \end{tabular}
+  \caption{Error rates from the dataset.}
+  \label{tbl:dataset_error_rates}
+  \end{center}
+\end{table}
+
 \section{Conclusions}
+
+// TODO: Tabela e descrição das diferentes taxas de erro
+
+// TODO: Descrição de alguns padrões interessantes que sobressaiam nos modelos
 
@@
 Robert E. Schapire, The Boosting Approach to Machine Learning: An Overview, 2001.
+
+\bibitem{nbtree}
+Ron Kohavi. Scaling Up the Accuracy of Naive-Bayes Classifiers: a Decision-Tree Hybrid. 1996
+
+\bibitem{dataset_papers}
+Adult Data Set, Machine Learning Repository, \url{http://archive.ics.uci.edu/ml/datasets/Adult}
+
+\bibitem{bayes4}
+Jie Cheng, Russell Greiner. Learning Bayesian Belief Network Classifiers: Algorithms and System. 2001
 

### 2009-12-10 report/final.tex

--- 2009-12-10 report/final.tex
@@
 
+The adult database has been widely used to benchmark novel learning algorithms.
+
+In the article \textit{Scaling Up the Accuracy of Naive-Bayes Classifiers: a Decision Tree hybrid},
+the author presents the nbtree algorithm, which is an hybrid approach suitable in learning scenarios
+when many attributes are likely to be relevant for a classification task, yet the attributes
+are not necessarily conditionally independent given the class \cite{nbtree}. It builds
+a decision tree with bayesian classifiers at the leaves.
+
+In the paper the author used various datasets, including the adult database. Records with unknown values
+were removed from the experiments. C4.5 reached an error rate of $\approx$15\% and NBTree $\approx$16.5\%,
+which are identical to our ID3 results, but worse than our C4.5 experiments. One reason for this is
+the incapability of Weka to use numeric attributes. Every other algorithm we used fared worse ($\approx$20\%).
+
+Another paper, \textit{Learning Bayesian Belief Network Classifiers: Algorithms and System}, explores
+various kinds of search algorithms used in Bayesian Networks and also comes up with very good error rates ($\approx$ 15\%).
+Other papers using this dataset can be found in \cite{dataset_papers}.
+
+The location from where we downloaded the dataset also contains results
+using various algorithms (Table \ref{tbl:dataset_error_rates}).
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c |}
+    \hline
+    \textbf{Algorithm} & \textbf{Error rate} \\ \hline
+    C4.5 & 15.54 \\ \hline
+    C4.5-auto & 14.46 \\ \hline
+    C4.5 rules & 14.94 \\ \hline
+    Voted ID3 (0.6) & 15.64 \\ \hline
+    Voted ID3 (0.8) & 16.47 \\ \hline
+    T2 & 16.84 \\ \hline
+    1R & 19.54 \\ \hline
+    NBTree & 14.10 \\ \hline
+    CN2 & 16.00 \\ \hline
+    HOODG & 14.82 \\ \hline
+    FSS Naive Bayes & 14.05 \\ \hline
+    IDTM (Decision table) & 14.05 \\ \hline
+    IDTM (Decision table) & 14.46 \\ \hline
+    Naive-Bayes & 16.12 \\ \hline
+    Nearest-neighbor (1) & 21.42 \\ \hline
+    Nearest-neighbor (3) & 20.35 \\ \hline
+    OC1 & 15.04 \\ \hline
+  \end{tabular}
+  \caption{Error rates from the dataset.}
+  \label{tbl:dataset_error_rates}
+  \end{center}
+\end{table}
+
 \section{Conclusions}

### 2009-12-10 report/final.tex
added conclusions on the classification methods.

--- 2009-12-10 report/final.tex
@@
 
-// TODO: Tabela e descrição das diferentes taxas de erro
-
-// TODO: Descrição de alguns padrões interessantes que sobressaiam nos modelos
+\begin{table}
+    \begin{center}
+    \begin{tabular}{ | l | l | }
+    \hline
+    \textbf{Algorithm} & \textbf{Error rate} \\ \hline
+    CN2 & 15.20 \\ \hline
+    ID3 & 15.68 \\ \hline
+    ADTree & 19.80 \\ \hline
+    ILP (Aleph) & 20.45 \\ \hline
+    Bayesian Networks (Simulated Annealing) & 21.09 \\ \hline
+    C4.5 & 21.69 \\ \hline
+    Random Forest & 22.71 \\ \hline
+    Bayesian Networks (Hill Climbing) & 22.93 \\ \hline
+    CART & 25.37 \\ \hline
+    Bayesian Networks (Genetic Search) & - \\ \hline
+    \end{tabular}
+    \caption{Error rates for the different algorithms used in the classification process}
+    \label{tbl:overall_results}
+    \end{center}
+\end{table}
+
+Table \ref{tbl:overall_results} summarizes the best error rates for the 
+different algorithms used in the classification process. Each algorithm has 
+different parameters to tweak it and the table only shows the best error rates
+obtained. CN2 has the best results for this particular dataset, although it
+takes longer than the majority of the other algorithms to build a classifier. 
+ID3 and C4.5 are usually quick to process and build classifiers within a decent
+error rate. ILP and CN2 build more expressive classifiers, capable of a better
+interpretation by a human. One special note goes to Inductive Logic Programming,
+a technique that allows the use of logic predicates to describe both the 
+background knowledge and the produced model. The use of prolog enables the 
+output produced by this technique to be better analyzed by a human agent. 
+Another important conclusion of the conducted experimental process is that the 
+worst error rate of the algorithms used is near 25\%, which denotes the power
+of the current classification algorithms, capable of building relevant models in
+a short period of time for relatively large datasets.
+
+Analyzing the models produced, one can see that having a high capital gain 
+clearly contributes to gaining more than 50K dollars per year. On the other
+hand, part time workers usually don't gain more than 50K dollars per year, like
+the younger and not married people.
 

### 2009-12-10 report/final.tex
correccoes + raw data no c4.5

--- 2009-12-10 report/final.tex
@@
 Instead of using the command line tools, we used the Java API to build classifiers.
-
-To use the data set we fetch it directly from the database and then apply various data pre-processing
-techniques that are available within Weka:
+We will build classifiers using only nominal attributes, with numeric values being
+discretized into categories, and classifiers with the data unchanged, to test
+the effects of using numeric attributes within the algorithms.
+
+To use the data set we fetch it directly from the database and then optionally
+apply various data pre-processing techniques that are available within Weka:
 
@@
 building the classifier and classifying all the examples.
+We will separate the results for the discretized data set and for the raw data set.
 
@@
 We executed the algorithm using different combinations of the parameters described above.
-Table \ref{tbl:results_c45} displays the obtained results. From the table
-we note that once the confidence threshold reaches a certain level ($\approx$0.55) the results
-are very similar for the unpruned classifier and the over-fitting effect starts to get worse,
-where the train data gets very good results ($\approx$11.19 error rate) and the test set gets a
-large error rate ($\approx$24).
-
-The confidence threshold for best test error rates appears to be in the interval 0.05-0.15.
+Table \ref{tbl:results_c45_discr} and \ref{tbl:results_c45_raw} displays the results
+with discretized and raw data, respectively. From the tables we can see that
+using numeric attributes increases the number of nodes and leaves, and
+that once the confidence threshold increases, the test error rate
+increases while the train error rate decreases, which indicates over-fitting.
+This effect is clearly visible for unpruned trees.
 
@@
     -C 0.4 & 13.68 & 22.49 & 0m24.325s & 835 & 721 \\ \hline
-    -C 0.45 & 13.26 & 22.43 & 0m25.086s & 1143 & 992 \\ \hline
     -C 0.5 & 12.95 & 22.51 & 0m24.874s & 1483 & 1292 \\ \hline
-    -C 0.55 & 11.19 & 24.94 & 0m25.899s & 6727 & 6036 \\ \hline
-    -C 0.6 & 11.19 & 23.94 & 0m26.882s & 6727 & 6036 \\ \hline
-    -C 0.7 & 11.19 & 23.94 & 0m26.638s & 6727 & 6036 \\ \hline
-    -C 0.8 & 11.19 & 23.94 & 0m26.818s & 6727 & 6036 \\ \hline
-    -C 0.9 & 11.19 & 23.94 & 0m26.634s & 6727 & 6036 \\ \hline
-  \end{tabular}
-  \caption{Results for C4.5.}
-  \label{tbl:results_c45}
+  \end{tabular}
+  \caption{Results for C4.5 with discretized attributes.}
+  \label{tbl:results_c45_discr}
   \end{center}
@@
 
-A decision tree built by this algorithm with the confidence threshold set to 0.05 is shown in Listing \ref{decision_tree_c45}.
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
+    -U & 9.48 & 21.58 & 0m7.525s & 6422 & 5491 \\ \hline
+    -C 0.05 & 13.78 & 19.49 & 0m7.667s & 149 & 105 \\ \hline
+    -C 0.1 & 13.37 & 19.06 & 0m7.969s & 278 & 212 \\ \hline
+    -C 0.15 & 12.99 & 18.84 & 0m7.708s & 361 & 278 \\ \hline
+    -C 0.2 & 12.95 & 18.89 & 0m7.812s & 373 & 284 \\ \hline
+    -C 0.25 & 12.86 & 18.86 & 0m7.974s & 433 & 331 \\ \hline
+    -C 0.3 & 12.11 & 19.73 & 0m8.064s & 806 & 624 \\ \hline
+    -C 0.35 & 12.02 & 19.77 & 0m8.035s & 878 & 678 \\ \hline
+    -C 0.4 & 11.54 & 20.47 & 0m8.255s & 1264 & 995 \\ \hline
+    -C 0.5 & 11.22 & 20.28 & 0m8.117s & 1603 & 1257 \\ \hline
+  \end{tabular}
+  \caption{Results for C4.5 with non-discretized attributes.}
+  \label{tbl:results_c45_raw}
+  \end{center}
+\end{table}
+
+A decision tree built by this algorithm with the confidence threshold set to 0.05 and
+discretized data is shown in Listing \ref{decision_tree_c45}.
 

### 2009-12-10 report/final.tex
added more conclusions on classification.

--- 2009-12-10 report/final.tex
@@
 hand, part time workers usually don't gain more than 50K dollars per year, like
-the younger and not married people.
+the younger and not married people. Having a higher education clearly 
+contributes to gaining more money, as does a stable life (married people
+gain more). People who work in private businesses or in the government are also
+more suitable to gain above 50K dollars per year. Working more hours per week
+obviously contributes to getting more money. Some less obvious conclusions show 
+that being a black own child, even though working more than 40 hours per week 
+doesn't lead to winning more than 50K dollars per year. In fact, being an own 
+child appears on a lot of rules describing people who win less than 50K. ILP
+also builds some interesting clauses where the nationality is associated with.
+This shows that being a native of some asian countries, such as taiwan, thailand
+or cambodia usually lead to gaining more than 50K dollars per year.
+
+Classification is a very active area on the machine learning research field, as
+the great amount of software tools and algorithms to tackle it show. It has
+arrived to a state where efficient models are built fast and with an expression
+suitable for analysis by a human agent. Progresses in relational data mining are
+allowing models to infer relations between entities, expanding their range of 
+applications by a lot. This work enabled a broad study of different types of 
+classification algorithms, analyzed both in terms of efficiency and 
+expressiveness of the models produced. Taking that into account, classification
+is nowadays a very reliable process for data analysis at large.
 

### 2009-12-10 report/final.tex

--- 2009-12-10 report/final.tex
@@
 
-// TODO: Tabela e descrição das diferentes taxas de erro
-
-// TODO: Descrição de alguns padrões interessantes que sobressaiam nos modelos
+\begin{table}
+    \begin{center}
+    \begin{tabular}{ | l | l | }
+    \hline
+    \textbf{Algorithm} & \textbf{Error rate} \\ \hline
+    CN2 & 15.20 \\ \hline
+    ID3 & 15.68 \\ \hline
+    ADTree & 19.80 \\ \hline
+    ILP (Aleph) & 20.45 \\ \hline
+    Bayesian Networks (Simulated Annealing) & 21.09 \\ \hline
+    C4.5 & 21.69 \\ \hline
+    Random Forest & 22.71 \\ \hline
+    Bayesian Networks (Hill Climbing) & 22.93 \\ \hline
+    CART & 25.37 \\ \hline
+    Bayesian Networks (Genetic Search) & - \\ \hline
+    \end{tabular}
+    \caption{Error rates for the different algorithms used in the classification process}
+    \label{tbl:overall_results}
+    \end{center}
+\end{table}
+
+Table \ref{tbl:overall_results} summarizes the best error rates for the 
+different algorithms used in the classification process. Each algorithm has 
+different parameters to tweak it and the table only shows the best error rates
+obtained. CN2 has the best results for this particular dataset, although it
+takes longer than the majority of the other algorithms to build a classifier. 
+ID3 and C4.5 are usually quick to process and build classifiers within a decent
+error rate. ILP and CN2 build more expressive classifiers, capable of a better
+interpretation by a human. One special note goes to Inductive Logic Programming,
+a technique that allows the use of logic predicates to describe both the 
+background knowledge and the produced model. The use of prolog enables the 
+output produced by this technique to be better analyzed by a human agent. 
+Another important conclusion of the conducted experimental process is that the 
+worst error rate of the algorithms used is near 25\%, which denotes the power
+of the current classification algorithms, capable of building relevant models in
+a short period of time for relatively large datasets.
+
+Analyzing the models produced, one can see that having a high capital gain 
+clearly contributes to gaining more than 50K dollars per year. On the other
+hand, part time workers usually don't gain more than 50K dollars per year, like
+the younger and not married people. Having a higher education clearly 
+contributes to gaining more money, as does a stable life (married people
+gain more). People who work in private businesses or in the government are also
+more suitable to gain above 50K dollars per year. Working more hours per week
+obviously contributes to getting more money. Some less obvious conclusions show 
+that being a black own child, even though working more than 40 hours per week 
+doesn't lead to winning more than 50K dollars per year. In fact, being an own 
+child appears on a lot of rules describing people who win less than 50K. ILP
+also builds some interesting clauses where the nationality is associated with.
+This shows that being a native of some asian countries, such as taiwan, thailand
+or cambodia usually lead to gaining more than 50K dollars per year.
+
+Classification is a very active area on the machine learning research field, as
+the great amount of software tools and algorithms to tackle it show. It has
+arrived to a state where efficient models are built fast and with an expression
+suitable for analysis by a human agent. Progresses in relational data mining are
+allowing models to infer relations between entities, expanding their range of 
+applications by a lot. This work enabled a broad study of different types of 
+classification algorithms, analyzed both in terms of efficiency and 
+expressiveness of the models produced. Taking that into account, classification
+is nowadays a very reliable process for data analysis at large.
 

### 2009-12-10 report/final.tex
better adtree experimental section

--- 2009-12-10 report/final.tex
@@
 We experimented the algorithm using various combinations of those two parameters.
+Tables \ref{tbl:results_adtree_discr} and \ref{tbl:results_adtree_raw} show
+results for both discretized data and raw data, respectively.
+An example tree built using non-discretized data and with the parameters \textbf{-E -2 -B 10}
+is displayed in Listing \ref{decision_tree_adtree}.
 
@@
   \end{tabular}
-  \caption{Results for ADTree.}
-  \label{tbl:results_adtree}
+  \caption{Results for ADTree for discretized data.}
+  \label{tbl:results_adtree_discr}
   \end{center}
@@
 
-From Table \ref{tbl:results_adtree} we can see that the number of boosting iterations affects
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
+    -E -3 -B 2 & 19.90 & 19.69 & 0m4.166s & 7 & 5 \\ \hline
+    -E -3 -B 5 & 18.43 & 18.58 & 0m7.070s & 16 & 11 \\ \hline
+    -E -3 -B 10 & 14.98 & 19.68 & 0m14.188s & 31 & 21 \\ \hline
+    -E -3 -B 15 & 14.78 & 20.18 & 0m24.777s & 46 & 31 \\ \hline
+    -E -3 -B 20 & 14.48 & 19.88 & 0m36.194s & 61 & 41 \\ \hline
+    -E -3 -B 25 & 14.08 & 18.87 & 0m51.563s & 76 & 51 \\ \hline
+    -E -3 -B 30 & 13.80 & 18.49 & 1m11.919s & 91 & 61 \\ \hline
+    -E -3 -B 35 & 13.74 & 18.65 & 1m32.886s & 106 & 71 \\ \hline
+    -E -3 -B 40 & 13.65 & 18.98 & 1m54.094s & 121 & 81 \\ \hline
+
+    -E -2 -B 2 & 19.90 & 19.69 & 0m3.995s & 7 & 5 \\ \hline
+    -E -2 -B 5 & 18.43 & 18.58 & 0m5.502s & 16 & 11 \\ \hline
+    -E -2 -B 10 & 15.14 & 20.07 & 0m8.662s & 31 & 21 \\ \hline
+    -E -2 -B 15 & 14.67 & 20.26 & 0m11.982s & 46 & 31 \\ \hline
+    -E -2 -B 20 & 14.67 & 19.95 & 0m15.543s & 61 & 41 \\ \hline
+    -E -2 -B 25 & 14.43 & 19.88 & 0m19.500s & 76 & 51 \\ \hline
+    -E -2 -B 30 & 14.34 & 19.85 & 0m22.929s & 91 & 61 \\ \hline
+    -E -2 -B 35 & 14.26 & 20.00 & 0m27.735s & 106 & 71 \\ \hline
+    -E -2 -B 40 & 14.05 & 19.71 & 0m32.832s & 121 & 81 \\ \hline
+
+    -E -1 -B 2 & 19.90 & 19.69 & 0m4.093s & 7 & 5 \\ \hline
+    -E -1 -B 5 & 18.43 & 18.58 & 0m5.414s & 16 & 11 \\ \hline
+    -E -1 -B 10 & 15.14 & 20.07 & 0m8.459s & 31 & 21 \\ \hline
+    -E -1 -B 15 & 14.67 & 20.26 & 0m11.937s & 46 & 31 \\ \hline
+    -E -1 -B 20 & 14.67 & 19.95 & 0m15.430s & 61 & 41 \\ \hline
+    -E -1 -B 25 & 14.43 & 19.88 & 0m19.413s & 76 & 51 \\ \hline
+    -E -1 -B 30 & 14.34 & 19.85 & 0m22.973s & 91 & 61 \\ \hline
+    -E -1 -B 35 & 14.26 & 20.00 & 0m27.246s & 106 & 71 \\ \hline
+    -E -1 -B 40 & 14.05 & 19.72 & 0m32.555s & 121 & 81 \\ \hline
+
+    -E 0 -B 2 & 19.90 & 19.69 & 0m4.039s & 7 & 5 \\ \hline
+    -E 0 -B 5 & 18.43 & 18.58 & 0m4.901s & 16 & 11 \\ \hline
+    -E 0 -B 10 & 15.15 & 20.10 & 0m6.701s & 31 & 21 \\ \hline
+    -E 0 -B 15 & 14.80 & 20.65 & 0m8.409s & 46 & 31 \\ \hline
+    -E 0 -B 20 & 14.50 & 20.27 & 0m9.858s & 61 & 41 \\ \hline
+    -E 0 -B 25 & 14.37 & 20.41 & 0m11.057s & 73 & 49 \\ \hline
+    -E 0 -B 30 & 14.21 & 20.28 & 0m12.712s & 88 & 59 \\ \hline
+    -E 0 -B 35 & 13.99 & 19.45 & 0m14.145s & 103 & 69 \\ \hline
+    -E 0 -B 40 & 14.04 & 20.21 & 0m15.273s & 118 & 79 \\ \hline
+
+  \end{tabular}
+  \caption{Results for ADTree for non-discretized data.}
+  \label{tbl:results_adtree_raw}
+  \end{center}
+\end{table}
+
+From tables \ref{tbl:results_adtree_discr} and \ref{tbl:results_adtree_raw}
+we can see that the number of boosting iterations affects
 both the execution time (directly proportional) and the error rate.
-The best error rate is for the parameters \textbf{-E -2 -B 20} giving $\approx$19.80\% for the test set.
-
-Another good observation is that this algorithm can build small trees that give satisfactory results, both for
-the train set and for the test set.
+The best error rate for raw data is using the parameters \textbf{-E -1 -B 5} giving $\approx$18.58\% for the test set.
+
+For nearly all parameters the algorithm gives best results when using numeric attributes instead
+of categories. Non-discretized data also generates larger trees.
+
+Another good observation is that this algorithm can build small trees that give
+satisfactory results, both for the train set and for the test set.
 
@@
 capital_gain = very_high: >50K (770.0/14.0)
+\end{lstlisting}
+
+\clearpage
+\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Decision tree built by ADTree with -E -2 -B 10},label=decision_tree_adtree]
+: -0.574
+|  (1)marital_status = married_civ_spouse: 0.467
+|  (1)marital_status != married_civ_spouse: -0.756
+|  (2)capital_gain < 7073.5: -0.11
+|  |  (5)capital_loss < 1748: -0.047
+|  |  (5)capital_loss >= 1748: 0.996
+|  |  (7)education = masters: 0.714
+|  |  (7)education != masters: -0.055
+|  (2)capital_gain >= 7073.5: 2.636
+|  (3)age < 27.5: -0.891
+|  (3)age >= 27.5: 0.11
+|  (4)hours_per_week < 41.5: -0.183
+|  (4)hours_per_week >= 41.5: 0.291
+|  (6)education = bachelors: 0.418
+|  (6)education != bachelors: -0.101
+|  (8)education = prof_school: 0.986
+|  (8)education != prof_school: -0.023
+|  (9)education = doctorate: 1.067
+|  (9)education != doctorate: -0.021
+|  (10)occupation = exec_managerial: 0.335
+|  (10)occupation != exec_managerial: -0.062
 \end{lstlisting}

### 2009-12-10 report/final.tex
update best values

@@
 both the execution time (directly proportional) and the error rate.
-The best error rate for raw data is using the parameters \textbf{-E -1 -B 5} giving $\approx$18.58\% for the test set.
+The best error rate for raw data is using the parameters \textbf{-E -3 -B 30} giving $\approx$18.49\% for the test set.
 
@@
     ID3 & 15.68 \\ \hline
-    ADTree & 19.80 \\ \hline
+    ADTree & 18.49 \\ \hline
     ILP (Aleph) & 20.45 \\ \hline
     Bayesian Networks (Simulated Annealing) & 21.09 \\ \hline
-    C4.5 & 21.69 \\ \hline
+    C4.5 & 18.84 \\ \hline
     Random Forest & 22.71 \\ \hline

### 2009-12-10 report/final.tex
minor fixes in the report.

--- 2009-12-10 report/final.tex
@@
 
-Deducing from the previous section, this is a classification problem. Given an adult
-with the mentioned attributes we want to classify him into one of the following groups:
+The adult database poses a classification problem. Given an adult
+with some attributes we want to classify him into one of the following groups:
 one gains more than 50000 dollars per year, the other group does not.
@@
 Understanding what affects the income can give a lot of information concerning society's behaviors.
-
-\section{Algorithms} \label{sec:algs}
-
-The algorithms we used to build classifiers can be grouped into four main
-groups: decision trees, rule induction \cite{rule_induction},
-inductive logic programming and bayesian classifiers \cite{bayes}.
-
-\subsection{Decision trees}
-
-For decision trees, the algorithms used are ID3 \cite{id3}, C4.5 \cite{c45}, CART \cite{cart},
-Random Forest \cite{random_forest}, and alternating decision trees (adtree) \cite{adtree}.
-
-Decision tree learning is a very commonly used method in classification. The main goal here is to
-create a tree model that can predict an object's class. Each interior node corresponds
-to one test for the object's attributes and each edge links to new interiors nodes based on the corresponding
-results. Each leaf in the tree represents the object class for the tests made
-across the respective path, from root to leaf.
-
-The ID3 algorithm (Iterative Dichotomiser 3) was one of the first decision trees algorithms.
-It was designed by Ross Quinlan.
-
-The C4.5 algorithm is the improved version of the ID3 algorithm, also designed by Quinlan.
-Contrary to ID3, this algorithm can handle both continuous and discrete attributes,
-it can use data with missing values and can also prune its decision trees.
-
-The CART algorithm builds classification and regression trees for predicting continuous dependent variables (regression)
-and categorical predictor variables (classification). The classic CART algorithm was popularized by Breiman et al. \cite{cart}
-
-The alternating decision tree is another type of decision tree algorithms. This tree
-consists of decision nodes, which specify a predicate condition, and prediction nodes, containing a single number.
-Each object is classified by following all paths for which all decision nodes are true and summing any
-prediction nodes along the way. Contrary to binary classification like ID3 ou C4.5, in adtrees one
-object can follow multiple paths.
-
-Finally, the Random Forest algorithm is an ensemble classifier which consists of many decision trees.
-Each object is classified multiple times by all decision trees and the mode class is used as the final class.
-Like the CART algorithm, Random Forest was also developed by Leo Breiman.
-
-\subsection{Rule induction}
-
-To experiment with rule induction algorithms we chose the CN2 \cite{cn2} algorithm.
-
-Rule induction works by extracting formal rules from a set of observations.
-Usually rules are expressed in the form \textit{if condition1 and condition2 ... then (decision, value)}.
-
-CN2 is an algorithm for inducing 
-propositional classification rules. CN2 consists of two main procedures: the 
-search procedure that performs beam search in order to find a single rule and 
-the control procedure that repeatedly executes the search. 
-The search procedure performs beam search using classification accuracy 
-of the rule as a heuristic function.
-Two different control procedures are used in CN2: one for inducing an ordered 
-list of rules and the other for the unordered case.
-
-\subsection{Inductive Logic Programming}
-
-Hypothesis construction is the core of any ILP system and there exists many different techniques for constructing them, including: Inverse Resolution, Relative Least General Generalisations, Inverse Implication, and Inverse Entailment.
-
-We will be using the Aleph system \cite{aleph} to experiment with inductive logic programming. This system
-uses the Inverse Entailment algorithm and has also incorporated features from many of its predecessors based on other techniques.
-
-\subsection{Bayesian classifiers}
-
-For bayesian classifiers we tested different kinds of search algorithms to build
-the network structure.
-
-The construction of the network structured is also known as
-\textit{local score metrics} and various algorithms can be employed for this optimization problem.
-
-The experiments conducted will use genetic based search, hill climbing \cite{bayes3}, and simulated annealing.
-
-For the second step in building a bayesian network classifier, we used a
-simple estimator to estimate the conditional probability distributions.
-This simple estimator produces
-direct estimates of the conditional probabilities. \cite{bayes_weka}
 
@@
 
+\section{Algorithms} \label{sec:algs}
+
+The algorithms we used to build classifiers can be grouped into four main
+groups: decision trees, rule induction \cite{rule_induction},
+inductive logic programming and bayesian classifiers \cite{bayes}.
+
+\subsection{Decision trees}
+
+For decision trees, the algorithms used are ID3 \cite{id3}, C4.5 \cite{c45}, CART \cite{cart},
+Random Forest \cite{random_forest}, and alternating decision trees (adtree) \cite{adtree}.
+
+Decision tree learning is a very commonly used method in classification. The main goal here is to
+create a tree model that can predict an object's class. Each interior node corresponds
+to one test for the object's attributes and each edge links to new interiors nodes based on the corresponding
+results. Each leaf in the tree represents the object class for the tests made
+across the respective path, from root to leaf.
+
+The ID3 algorithm (Iterative Dichotomiser 3) was one of the first decision trees algorithms.
+It was designed by Ross Quinlan.
+
+The C4.5 algorithm is the improved version of the ID3 algorithm, also designed by Quinlan.
+Contrary to ID3, this algorithm can handle both continuous and discrete attributes,
+it can use data with missing values and can also prune its decision trees.
+
+The CART algorithm builds classification and regression trees for predicting continuous dependent variables (regression)
+and categorical predictor variables (classification). The classic CART algorithm was popularized by Breiman et al. \cite{cart}
+
+The alternating decision tree is another type of decision tree algorithms. This tree
+consists of decision nodes, which specify a predicate condition, and prediction nodes, containing a single number.
+Each object is classified by following all paths for which all decision nodes are true and summing any
+prediction nodes along the way. Contrary to binary classification like ID3 ou C4.5, in adtrees one
+object can follow multiple paths.
+
+Finally, the Random Forest algorithm is an ensemble classifier which consists of many decision trees.
+Each object is classified multiple times by all decision trees and the mode class is used as the final class.
+Like the CART algorithm, Random Forest was also developed by Leo Breiman.
+
+\subsection{Rule induction}
+
+To experiment with rule induction algorithms we chose the CN2 \cite{cn2} algorithm.
+
+Rule induction works by extracting formal rules from a set of observations.
+Usually rules are expressed in the form \textit{if condition1 and condition2 ... then (decision, value)}.
+
+CN2 is an algorithm for inducing 
+propositional classification rules. CN2 consists of two main procedures: the 
+search procedure that performs beam search in order to find a single rule and 
+the control procedure that repeatedly executes the search. 
+The search procedure performs beam search using classification accuracy 
+of the rule as a heuristic function.
+Two different control procedures are used in CN2: one for inducing an ordered 
+list of rules and the other for the unordered case.
+
+\subsection{Inductive Logic Programming}
+
+Hypothesis construction is the core of any ILP system and there exists many different techniques for constructing them, including: Inverse Resolution, Relative Least General Generalisations, Inverse Implication, and Inverse Entailment.
+
+We will be using the Aleph system \cite{aleph} to experiment with inductive logic programming. This system
+uses the Inverse Entailment algorithm and has also incorporated features from many of its predecessors based on other techniques.
+
+\subsection{Bayesian classifiers}
+
+For bayesian classifiers we tested different kinds of search algorithms to build
+the network structure.
+
+The construction of the network structured is also known as
+\textit{local score metrics} and various algorithms can be employed for this optimization problem.
+
+The experiments conducted will use genetic based search, hill climbing \cite{bayes3}, and simulated annealing.
+
+For the second step in building a bayesian network classifier, we used a
+simple estimator to estimate the conditional probability distributions.
+This simple estimator produces
+direct estimates of the conditional probabilities \cite{bayes_weka}.
+
 \section{Experimental Context}
@@
 We experimented the algorithm using various combinations of those two parameters.
-Tables \ref{tbl:results_adtree_discr} and \ref{tbl:results_adtree_raw} show
-results for both discretized data and raw data, respectively.
-An example tree built using non-discretized data and with the parameters \textbf{-E -2 -B 10}
-is displayed in Listing \ref{decision_tree_adtree}.
 
@@
   \end{tabular}
-  \caption{Results for ADTree for discretized data.}
-  \label{tbl:results_adtree_discr}
+  \caption{Results for ADTree.}
+  \label{tbl:results_adtree}
   \end{center}
@@
 
-\begin{table}[ht]
-  \begin{center}
-  \begin{tabular}{ | l | c | c | c | c | c |}
-    \hline
-    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
-    -E -3 -B 2 & 19.90 & 19.69 & 0m4.166s & 7 & 5 \\ \hline
-    -E -3 -B 5 & 18.43 & 18.58 & 0m7.070s & 16 & 11 \\ \hline
-    -E -3 -B 10 & 14.98 & 19.68 & 0m14.188s & 31 & 21 \\ \hline
-    -E -3 -B 15 & 14.78 & 20.18 & 0m24.777s & 46 & 31 \\ \hline
-    -E -3 -B 20 & 14.48 & 19.88 & 0m36.194s & 61 & 41 \\ \hline
-    -E -3 -B 25 & 14.08 & 18.87 & 0m51.563s & 76 & 51 \\ \hline
-    -E -3 -B 30 & 13.80 & 18.49 & 1m11.919s & 91 & 61 \\ \hline
-    -E -3 -B 35 & 13.74 & 18.65 & 1m32.886s & 106 & 71 \\ \hline
-    -E -3 -B 40 & 13.65 & 18.98 & 1m54.094s & 121 & 81 \\ \hline
-
-    -E -2 -B 2 & 19.90 & 19.69 & 0m3.995s & 7 & 5 \\ \hline
-    -E -2 -B 5 & 18.43 & 18.58 & 0m5.502s & 16 & 11 \\ \hline
-    -E -2 -B 10 & 15.14 & 20.07 & 0m8.662s & 31 & 21 \\ \hline
-    -E -2 -B 15 & 14.67 & 20.26 & 0m11.982s & 46 & 31 \\ \hline
-    -E -2 -B 20 & 14.67 & 19.95 & 0m15.543s & 61 & 41 \\ \hline
-    -E -2 -B 25 & 14.43 & 19.88 & 0m19.500s & 76 & 51 \\ \hline
-    -E -2 -B 30 & 14.34 & 19.85 & 0m22.929s & 91 & 61 \\ \hline
-    -E -2 -B 35 & 14.26 & 20.00 & 0m27.735s & 106 & 71 \\ \hline
-    -E -2 -B 40 & 14.05 & 19.71 & 0m32.832s & 121 & 81 \\ \hline
-
-    -E -1 -B 2 & 19.90 & 19.69 & 0m4.093s & 7 & 5 \\ \hline
-    -E -1 -B 5 & 18.43 & 18.58 & 0m5.414s & 16 & 11 \\ \hline
-    -E -1 -B 10 & 15.14 & 20.07 & 0m8.459s & 31 & 21 \\ \hline
-    -E -1 -B 15 & 14.67 & 20.26 & 0m11.937s & 46 & 31 \\ \hline
-    -E -1 -B 20 & 14.67 & 19.95 & 0m15.430s & 61 & 41 \\ \hline
-    -E -1 -B 25 & 14.43 & 19.88 & 0m19.413s & 76 & 51 \\ \hline
-    -E -1 -B 30 & 14.34 & 19.85 & 0m22.973s & 91 & 61 \\ \hline
-    -E -1 -B 35 & 14.26 & 20.00 & 0m27.246s & 106 & 71 \\ \hline
-    -E -1 -B 40 & 14.05 & 19.72 & 0m32.555s & 121 & 81 \\ \hline
-
-    -E 0 -B 2 & 19.90 & 19.69 & 0m4.039s & 7 & 5 \\ \hline
-    -E 0 -B 5 & 18.43 & 18.58 & 0m4.901s & 16 & 11 \\ \hline
-    -E 0 -B 10 & 15.15 & 20.10 & 0m6.701s & 31 & 21 \\ \hline
-    -E 0 -B 15 & 14.80 & 20.65 & 0m8.409s & 46 & 31 \\ \hline
-    -E 0 -B 20 & 14.50 & 20.27 & 0m9.858s & 61 & 41 \\ \hline
-    -E 0 -B 25 & 14.37 & 20.41 & 0m11.057s & 73 & 49 \\ \hline
-    -E 0 -B 30 & 14.21 & 20.28 & 0m12.712s & 88 & 59 \\ \hline
-    -E 0 -B 35 & 13.99 & 19.45 & 0m14.145s & 103 & 69 \\ \hline
-    -E 0 -B 40 & 14.04 & 20.21 & 0m15.273s & 118 & 79 \\ \hline
-
-  \end{tabular}
-  \caption{Results for ADTree for non-discretized data.}
-  \label{tbl:results_adtree_raw}
-  \end{center}
-\end{table}
-
-From tables \ref{tbl:results_adtree_discr} and \ref{tbl:results_adtree_raw}
-we can see that the number of boosting iterations affects
+From Table \ref{tbl:results_adtree} we can see that the number of boosting iterations affects
 both the execution time (directly proportional) and the error rate.
-The best error rate for raw data is using the parameters \textbf{-E -1 -B 5} giving $\approx$18.58\% for the test set.
-
-For nearly all parameters the algorithm gives best results when using numeric attributes instead
-of categories. Non-discretized data also generates larger trees.
-
-Another good observation is that this algorithm can build small trees that give
-satisfactory results, both for the train set and for the test set.
+The best error rate is for the parameters \textbf{-E -2 -B 20} giving $\approx$19.80\% for the test set.
+
+Another good observation is that this algorithm can build small trees that give satisfactory results, both for
+the train set and for the test set.
 
@@
 an information gain error estimate produces rules faster and with an acceptable
-error rate (18.9\%). To analyze the expressiveness of the rules produced, a set
-of rules that cover the most examples of the training data are described in 
-listing \ref{cn2_rules} (using the ordered algorithm with the laplacian error 
-estimation).
+error rate (18.9\%). Models generating an ordered set of rules with the 
+laplacian or naive error estimate overfit the training data. To analyze the 
+expressiveness of the rules produced, a set of rules that cover the most 
+examples of the training data are described in listing \ref{cn2_rules} (using 
+the ordered algorithm with the laplacian error estimation).
 
@@
 capital_gain = very_high: >50K (770.0/14.0)
-\end{lstlisting}
-
-\clearpage
-\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Decision tree built by ADTree with -E -2 -B 10},label=decision_tree_adtree]
-: -0.574
-|  (1)marital_status = married_civ_spouse: 0.467
-|  (1)marital_status != married_civ_spouse: -0.756
-|  (2)capital_gain < 7073.5: -0.11
-|  |  (5)capital_loss < 1748: -0.047
-|  |  (5)capital_loss >= 1748: 0.996
-|  |  (7)education = masters: 0.714
-|  |  (7)education != masters: -0.055
-|  (2)capital_gain >= 7073.5: 2.636
-|  (3)age < 27.5: -0.891
-|  (3)age >= 27.5: 0.11
-|  (4)hours_per_week < 41.5: -0.183
-|  (4)hours_per_week >= 41.5: 0.291
-|  (6)education = bachelors: 0.418
-|  (6)education != bachelors: -0.101
-|  (8)education = prof_school: 0.986
-|  (8)education != prof_school: -0.023
-|  (9)education = doctorate: 1.067
-|  (9)education != doctorate: -0.021
-|  (10)occupation = exec_managerial: 0.335
-|  (10)occupation != exec_managerial: -0.062
 \end{lstlisting}

### 2009-12-10 report/final.tex
rework random forests

--- 2009-12-10 report/final.tex
@@
 In Weka, the Random Forest algorithm can be parametrized by the number
-of trees to build \textbf{$<$-I integer$>$}. Table \ref{tbl:results_random_forest} shows results for
-a varying number of trees.
-
-From the table we conclude that we need a certain amount of trees ($\approx$80) to
-reach a satisfactory error rate ($<$22\%) for the test set. Interestingly
-the error rate for the train set is low ($\approx$9\%), which indicates over-fitting
-when comparing the difference to the test error rate.
+of trees to build \textbf{$<$-I integer$>$}. Tables \ref{tbl:results_random_forest_discr}
+and \ref{tbl:results_random_forest_raw} show results for
+a varying number of trees, for discretized data and non-discretized data, respectively.
+
+The classifier built with non-discretized data can easily reach a satisfactory
+error rate with less trees ($\approx$4 against $\approx$40).
+
+For both cases, the error rate for the train set is low ($\approx$9\%), which indicates over-fitting
+when comparing to the test error rate.
 
@@
   \end{tabular}
-  \caption{Results for Random Forest.}
-  \label{tbl:results_random_forest}
+  \caption{Results for Random Forest using discretized data.}
+  \label{tbl:results_random_forest_discr}
+  \end{center}
+\end{table}
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    -I 2 & 5.72 & 25.70 & 0m4.882s \\ \hline
+    -I 3 & 3.89 & 24.58 & 0m5.588s \\ \hline
+    -I 4 & 3.81 & 23.40 & 0m6.086s \\ \hline
+    -I 5 & 3.20 & 23.43 & 0m6.839s \\ \hline
+    -I 7 & 2.86 & 21.01 & 0m8.190s \\ \hline
+    -I 10 & 2.69 & 21.15 & 0m10.162s \\ \hline
+    -I 15 & 2.52 & 21.01 & 0m13.597s \\ \hline
+    -I 20 & 2.46 & 20.83 & 0m16.970s \\ \hline
+    -I 40 & 2.44 & 21.26 & 0m30.488s \\ \hline
+    -I 60 & 2.44 & 21.67 & 0m46.093s \\ \hline
+    -I 80 & 2.44 & 21.70 & 1m1.347s \\ \hline
+    -I 100 & 2.44 & 21.75 & 1m15.592s \\ \hline
+    -I 400 & 2.44 & 21.62 & 7m51.624s \\ \hline
+  \end{tabular}
+  \caption{Results for Random Forest using non-discretized data.}
+  \label{tbl:results_random_forest_raw}
   \end{center}

### 2009-12-10 report/final.tex
better cart

@@
   \end{tabular}
-  \caption{Results for CART.}
-  \label{tbl:results_cart}
+  \caption{Results for CART with discretized data.}
+  \label{tbl:results_cart_discr}
   \end{center}
@@
 
-CART results are displayed in Table \ref{tbl:results_cart}.
-From these results we can see that there isn't a lot of variation in the results.
-The algorithm also performs badly when compared to C4.5 or ADTree.
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
+    -M 2 -N 5 & 12.68 & 22.26 & 2m34.703s & 157 & 79 \\ \hline
+    -M 2 -N 10 & 12.72 & 22.26 & 4m49.261s & 151 & 76 \\ \hline
+    -M 2 -N 15 & 12.22 & 21.34 & 7m4.560s & 237 & 119 \\ \hline
+    -M 2 -N 20 & 11.75 & 21.70 & 9m14.989s & 341 & 171 \\ \hline
+
+    -M 3 -N 5 & 12.68 & 22.26 & 2m21.439s & 157 & 79 \\ \hline
+    -M 3 -N 10 & 12.72 & 22.26 & 4m35.048s & 151 & 76 \\ \hline
+    -M 3 -N 15 & 12.22 & 21.34 & 6m45.982s & 237 & 119 \\ \hline
+    -M 3 -N 20 & 11.84 & 21.70 & 7m44.483s & 321 & 161 \\ \hline
+
+    -M 5 -N 5 & 12.72 & 22.26 & 2m18.255s & 151 & 76 \\ \hline
+    -M 5 -N 10 & 12.60 & 22.78 & 4m28.169s & 171 & 86 \\ \hline
+    -M 5 -N 15 & 13.29 & 22.01 & 6m45.657s & 83 & 42 \\ \hline
+    -M 5 -N 20 & 12.22 & 21.43 & 8m53.232s & 259 & 130 \\ \hline
+  \end{tabular}
+  \caption{Results for CART with non-discretized data.}
+  \label{tbl:results_cart_raw}
+  \end{center}
+\end{table}
+
+CART results are displayed in tables \ref{tbl:results_cart_discr} and
+\ref{tbl:results_cart_raw}.
+From these results we can see that there isn't a lot of error rate variation
+for each type of data.
+The algorithm also performs worse compared to C4.5 or ADTree
+and takes a much longer time to build the classifier.
 

### 2009-12-10 report/final.tex
more results

@@
 This section shows performance results for the various algorithms described in section \ref{sec:algs}.
+For every algorithm we will considerate the following aspects: train error rate, test error rate and time.
+Each error rate is displayed as a percentage.
 
@@
 
+With the discretized data set this algorithm performs quite badly
+against other algorithms with the same data, like C4.5, with $\approx$25\% error rate.
+
 \subsection{CN2}
@@
 
+Again, we will use two train sets: a discretized set and a raw set with numeric values.
+The bayesian algorithm in Weka, when fed with numeric values,
+it auto discretizes the values, as the nature of bayesian algorithms demands.
+
 \subsubsection{Genetic Search}
@@
 
-For this experiment we will only use a portion of the train data to speed up the process.
+For this experiment we will only use 1\% of the train data to speed up the process.
 
@@
     \hline
-    \textbf{Parameter} & \textbf{Portion} & \textbf{Portion error rate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
-    GeneticSearch -C -M -L 3 -A 3 -U 3 & 5 & & & & \\ \hline
-    GeneticSearch -C -M -L 3 -A 3 -U 5 & 5 & & & & \\ \hline
-    GeneticSearch -C -M -L 3 -A 3 -U 15 & 5 & & & & \\ \hline
+    \textbf{Parameter} & \textbf{Data} & \textbf{Portion error rate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    GeneticSearch -C -M -L 3 -A 3 -U 3 & discretized & 13.50 & 18.72 & 22.69 & 0m25.115s \\ \hline
+    GeneticSearch -C -M -L 3 -A 3 -U 5 & raw & 13.80 & 18.23 & 21.26 & 0m3.601s \\ \hline
   \end{tabular}
@@
 
-The best configuration for hill climbing (Table \ref{tbl:results_bayesian_networks_hc})
-is when the number of parents is exactly 1, still, the error rate is somewhat unsatisfactory when
-comparing to other results.
+The best configuration for hill climbing (tables \ref{tbl:results_bayesian_networks_hc_discr}
+and \ref{tbl:results_bayesian_networks_hc_raw})
+is when the number of parents is exactly 1.
 
@@
   \end{tabular}
-  \caption{Results for Bayesian Networks / Hill Climbing.}
-  \label{tbl:results_bayesian_networks_hc}
+  \caption{Results for Bayesian Networks / Hill Climbing with discretized data.}
+  \label{tbl:results_bayesian_networks_hc_discr}
+  \end{center}
+\end{table}
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    HillClimber -P 1 & 16.61 & 20.03 & 0m6.873s \\ \hline
+    HillClimber -P 2 & 13.79 & 23.20 & 0m7.023s \\ \hline
+    HillClimber -P 3 & 12.86 & 25.14 & 0m7.166s \\ \hline
+    HillClimber -P 4 & 12.54 & 25.60 & 0m8.326s \\ \hline
+    HillClimber -P 10 & 12.54 & 25.60 & 0m9.236s \\ \hline
+  \end{tabular}
+  \caption{Results for Bayesian Networks / Hill Climbing with raw data.}
+  \label{tbl:results_bayesian_networks_hc_raw}
   \end{center}
@@
 
-We tried various temperatures values and came up with the Table \ref{tbl:results_bayesian_networks_sa}.
+We tried various temperatures values and came up with the tables \ref{tbl:results_bayesian_networks_sa_discr} and
+\ref{tbl:results_bayesian_networks_sa_raw}.
 
@@
   \end{tabular}
-  \caption{Results for Bayesian Networks / Simulated Annealing.}
-  \label{tbl:results_bayesian_networks_sa}
+  \caption{Results for Bayesian Networks / Simulated Annealing with discretized data.}
+  \label{tbl:results_bayesian_networks_sa_discr}
+  \end{center}
+\end{table}
+
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c |}
+    \hline
+    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    SimulatedAnnealing -U 75 -D 0.4 -A 20 & 12.90 & 22.74 & 0m7.270s \\ \hline
+    SimulatedAnnealing -U 55 -D 0.4 -A 20 & 13.53 & 19.15 & 0m7.223s \\ \hline
+    SimulatedAnnealing -U 52 -D 0.4 -A 20 & 13.55 & 19.15 & 0m7.712s \\ \hline
+    SimulatedAnnealing -U 50 -D 0.4 -A 20 & 13.55 & 19.15 & 0m7.164s \\ \hline
+    SimulatedAnnealing -U 50 -D 0.4 -A 40 & 13.55 & 19.15 & 0m7.252s \\ \hline
+    SimulatedAnnealing -U 25 -D 0.4 -A 20 & 15.85 & 19.90 & 0m7.239s \\ \hline
+    SimulatedAnnealing -U 10 -D 0.4 -A 20 & 16.60 & 19.91 & 0m7.168s \\ \hline
+  \end{tabular}
+  \caption{Results for Bayesian Networks / Simulated Annealing with raw data.}
+  \label{tbl:results_bayesian_networks_sa_raw}
   \end{center}
@@
 tuple of configuration values. Anyway, with just a few guesses we came up
-with $\approx21$\% error rate, which is quite good when comparing to other algorithms.
+with $\approx19.15$\% error rate, which is quite good when comparing to other algorithms.
+The Bayesian network can be seen in Listing \ref{bayes_net_example}.
 
@@
 
+\clearpage
+\section{Bayesian Networks}
+
+\begin{lstlisting}[language=c,frame=single,breaklines=true,basicstyle=\footnotesize\ttfamily,caption={Bayesian Network built using a Simulated Annealing search algorithm},label=bayes_net_example]
+Bayes Network Classifier
+Using ADTree
+#attributes=13 #classindex=12
+Network structure (nodes followed by parents)
+age(9): plus_50 workclass 
+workclass(9): plus_50 
+education(16): age plus_50 
+marital_status(7): plus_50 
+occupation(15): plus_50 
+relationship(6): plus_50 
+race(5): plus_50 
+sex(2): plus_50 capital_loss 
+capital_gain(15): plus_50 
+capital_loss(15): plus_50 
+hours_per_week(6): plus_50 
+native_country(42): plus_50 
+plus_50(2): 
+LogScore Bayes: -454151.06415610015
+LogScore BDeu: -458613.8065210351
+LogScore MDL: -458373.7686675328
+LogScore ENTROPY: -454908.41334505216
+LogScore AIC: -455575.41334505216
+\end{lstlisting}
+
 \end{document}

### 2009-12-10 report/final.tex
put best values

@@
     ILP (Aleph) & 20.45 \\ \hline
-    Bayesian Networks (Simulated Annealing) & 21.09 \\ \hline
+    Bayesian Networks (Simulated Annealing) & 19.15 \\ \hline
     C4.5 & 18.84 \\ \hline
-    Random Forest & 22.71 \\ \hline
-    Bayesian Networks (Hill Climbing) & 22.93 \\ \hline
-    CART & 25.37 \\ \hline
-    Bayesian Networks (Genetic Search) & - \\ \hline
+    Random Forest & 21.01 \\ \hline
+    Bayesian Networks (Hill Climbing) & 20.03 \\ \hline
+    CART & 21.34 \\ \hline
+    Bayesian Networks (Genetic Search) & 21.26 \\ \hline
     \end{tabular}

### 2009-12-10 report/final.tex
updates.

--- 2009-12-10 report/final.tex
@@
 In Weka, the Random Forest algorithm can be parametrized by the number
-of trees to build \textbf{$<$-I integer$>$}. Tables \ref{tbl:results_random_forest_discr}
-and \ref{tbl:results_random_forest_raw} show results for
-a varying number of trees, for discretized data and non-discretized data, respectively.
-
-The classifier built with non-discretized data can easily reach a satisfactory
-error rate with less trees ($\approx$4 against $\approx$40).
-
-For both cases, the error rate for the train set is low ($\approx$9\%), which indicates over-fitting
-when comparing to the test error rate.
+of trees to build \textbf{$<$-I integer$>$}. Table \ref{tbl:results_random_forest} shows results for
+a varying number of trees.
+
+From the table we conclude that we need a certain amount of trees ($\approx$80) to
+reach a satisfactory error rate ($<$22\%) for the test set. Interestingly
+the error rate for the train set is low ($\approx$9\%), which indicates over-fitting
+when comparing the difference to the test error rate.
 
@@
   \end{tabular}
-  \caption{Results for Random Forest using discretized data.}
-  \label{tbl:results_random_forest_discr}
-  \end{center}
-\end{table}
-
-\begin{table}[ht]
-  \begin{center}
-  \begin{tabular}{ | l | c | c | c |}
-    \hline
-    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
-    -I 2 & 5.72 & 25.70 & 0m4.882s \\ \hline
-    -I 3 & 3.89 & 24.58 & 0m5.588s \\ \hline
-    -I 4 & 3.81 & 23.40 & 0m6.086s \\ \hline
-    -I 5 & 3.20 & 23.43 & 0m6.839s \\ \hline
-    -I 7 & 2.86 & 21.01 & 0m8.190s \\ \hline
-    -I 10 & 2.69 & 21.15 & 0m10.162s \\ \hline
-    -I 15 & 2.52 & 21.01 & 0m13.597s \\ \hline
-    -I 20 & 2.46 & 20.83 & 0m16.970s \\ \hline
-    -I 40 & 2.44 & 21.26 & 0m30.488s \\ \hline
-    -I 60 & 2.44 & 21.67 & 0m46.093s \\ \hline
-    -I 80 & 2.44 & 21.70 & 1m1.347s \\ \hline
-    -I 100 & 2.44 & 21.75 & 1m15.592s \\ \hline
-    -I 400 & 2.44 & 21.62 & 7m51.624s \\ \hline
-  \end{tabular}
-  \caption{Results for Random Forest using non-discretized data.}
-  \label{tbl:results_random_forest_raw}
+  \caption{Results for Random Forest.}
+  \label{tbl:results_random_forest}
   \end{center}
@@
   \end{tabular}
-  \caption{Results for CART with discretized data.}
-  \label{tbl:results_cart_discr}
+  \caption{Results for CART.}
+  \label{tbl:results_cart}
   \end{center}
@@
 
-\begin{table}[ht]
-  \begin{center}
-  \begin{tabular}{ | l | c | c | c | c | c |}
-    \hline
-    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
-    -M 2 -N 5 & 12.68 & 22.26 & 2m34.703s & 157 & 79 \\ \hline
-    -M 2 -N 10 & 12.72 & 22.26 & 4m49.261s & 151 & 76 \\ \hline
-    -M 2 -N 15 & 12.22 & 21.34 & 7m4.560s & 237 & 119 \\ \hline
-    -M 2 -N 20 & 11.75 & 21.70 & 9m14.989s & 341 & 171 \\ \hline
-
-    -M 3 -N 5 & 12.68 & 22.26 & 2m21.439s & 157 & 79 \\ \hline
-    -M 3 -N 10 & 12.72 & 22.26 & 4m35.048s & 151 & 76 \\ \hline
-    -M 3 -N 15 & 12.22 & 21.34 & 6m45.982s & 237 & 119 \\ \hline
-    -M 3 -N 20 & 11.84 & 21.70 & 7m44.483s & 321 & 161 \\ \hline
-
-    -M 5 -N 5 & 12.72 & 22.26 & 2m18.255s & 151 & 76 \\ \hline
-    -M 5 -N 10 & 12.60 & 22.78 & 4m28.169s & 171 & 86 \\ \hline
-    -M 5 -N 15 & 13.29 & 22.01 & 6m45.657s & 83 & 42 \\ \hline
-    -M 5 -N 20 & 12.22 & 21.43 & 8m53.232s & 259 & 130 \\ \hline
-  \end{tabular}
-  \caption{Results for CART with non-discretized data.}
-  \label{tbl:results_cart_raw}
-  \end{center}
-\end{table}
-
-CART results are displayed in tables \ref{tbl:results_cart_discr} and
-\ref{tbl:results_cart_raw}.
-From these results we can see that there isn't a lot of error rate variation
-for each type of data.
-The algorithm also performs worse compared to C4.5 or ADTree
-and takes a much longer time to build the classifier.
+CART results are displayed in Table \ref{tbl:results_cart}.
+From these results we can see that there isn't a lot of variation in the results.
+The algorithm also performs badly when compared to C4.5 or ADTree.
 

### 2009-12-10 report/final.tex
ordering of results in overall table.

--- 2009-12-10 report/final.tex
@@
 
-It is also important to verify some correlations between variables. To that
+It is also important to verify some relations between variables. To that
 matter, we've tested some apparent situations, such as the influence of the age,
@@
     ADTree & 18.49 \\ \hline
+    C4.5 & 18.84 \\ \hline
+    Bayesian Networks (Simulated Annealing) & 19.15 \\ \hline
+    Bayesian Networks (Hill Climbing) & 20.03 \\ \hline
     ILP (Aleph) & 20.45 \\ \hline
-    Bayesian Networks (Simulated Annealing) & 21.09 \\ \hline
-    C4.5 & 18.84 \\ \hline
-    Random Forest & 22.71 \\ \hline
-    Bayesian Networks (Hill Climbing) & 22.93 \\ \hline
-    CART & 25.37 \\ \hline
-    Bayesian Networks (Genetic Search) & - \\ \hline
+    Random Forest & 21.01 \\ \hline
+    Bayesian Networks (Genetic Search) & 21.26 \\ \hline
+    CART & 21.34 \\ \hline
     \end{tabular}

### 2009-12-14 report/final.tex
ordering of figures.
minor report updates.

@@
 
-There are 3620 records with missing values. The attributes that contain missing values are: \textbf{Workclass}, \textbf{Occupation} and \textbf{Native-country}.
+For the training set, there are 3620 records with missing values. The attributes
+that contain missing values are: \textbf{Workclass}, \textbf{Occupation} and 
+\textbf{Native-country}.
 
@@
 
-The table \ref{tab:summary} summarizes the distribution of values on the studied
-dataset. From a first observation, one can see that the amount of missing values
+The table \ref{tab:summary} summarizes the distribution of values in the training data on the studied
+dataset, while the table \ref{tab:summarytest} summarizes the same distribution but
+for the test set. A quick observation shows that the distribution of values is 
+similar in both the training and the test set. Taking that into account, some
+analysis was conducted on the training data, expecting some interesting 
+properties to arise. From a first observation, one can see that the amount of missing values
 is low. On the other side, in some of the attributes, some classes have a
@@
 Two different control procedures are used in CN2: one for inducing an ordered 
-list of rules and the other for the unordered case.
+list of rules and the other for an unordered one.
 
@@
 
-\begin{table}
+\begin{table}[ht]
   \begin{center}
@@
 
-\begin{table}
+\begin{table}[ht]
 \begin{center}
@@
 The error rate produced by Aleph when evaluating the test data was 20.45\%.
-18881882 clauses were built, taking around 5 hours for the whole training
+Around 7000 clauses were built, taking around 5 hours for the whole training
 data. Despite the number of clauses, most of them are irrelevant, directly
@@
 
-\begin{table}
+\begin{table}[ht]
     \begin{center}
@@
 }
-\caption{Summary of the value distribution in the database}
+\caption{Summary of the value distribution on the training data of the dataset}
 \label{tab:summary}
+\end{table}
+
+\begin{table}
+\centering
+\subtable[Age]{
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 17 \\
+       \hline
+       1st Qu. & 28 \\
+       \hline
+       Median & 37 \\
+       \hline
+       Mean & 38.77 \\
+       \hline
+       3rd Qu. & 48 \\
+       \hline
+       Max & 90 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Workclass]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       private & 11210 \\
+       \hline
+       self\_emp\_not\_inc & 1321 \\
+       \hline
+       local\_gov & 1043 \\
+       \hline
+       state\_gov & 683 \\
+       \hline
+       self\_emp\_inc & 579 \\
+       \hline
+       (Other) & 482 \\
+       \hline
+       NA's & 963 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Education]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       hs\_grad & 5283 \\
+       \hline
+       some\_college & 3587 \\
+       \hline
+       bachelors & 2670 \\
+       \hline
+       masters & 934 \\
+       \hline
+       assoc\_voc & 679 \\
+       \hline
+       11th & 637 \\
+       \hline
+       (Other) & 2491 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Marital Status]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       divorced & 2190 \\
+       \hline
+       married\_af\_spouse & 14 \\
+       \hline
+       married\_civ\_spouse & 7403 \\
+       \hline
+       married\_spouse\_absent & 210 \\
+       \hline
+       never\_married & 5434 \\
+       \hline
+       separated & 505 \\
+       \hline
+       widowed & 525 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Occupation]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       prof\_specialty & 2032 \\
+       \hline
+       exec\_managerial & 2020 \\
+       \hline
+       craft\_repair & 2013 \\
+       \hline
+       sales & 1854 \\
+       \hline
+       adm\_clerical & 1841 \\
+       \hline
+       (Other) & 5555 \\
+       \hline
+       NA's & 966 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Relatonship]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       husband & 6523 \\
+       \hline
+       not\_in\_family & 4278 \\
+       \hline
+       other\_relative & 525 \\
+       \hline
+       own\_child & 2513 \\
+       \hline
+       unmarried & 1679 \\
+       \hline
+       wife & 763 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Race]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       amer\_indian\_eskimo & 159 \\
+       \hline
+       asian\_pac\_islander & 480 \\
+       \hline
+       black & 1561 \\
+       \hline
+       other & 135 \\
+       \hline
+       white & 13946 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Sex]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       female & 5421 \\
+       \hline
+       male & 10860 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Capital Gain]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 0 \\
+       \hline
+       1st Qu. & 0 \\
+       \hline
+       Median & 0 \\
+       \hline
+       Mean & 1082 \\
+       \hline
+       3rd Qu. & 0 \\
+       \hline
+       Max. & 99999 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Capital Loss]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 0 \\
+       \hline
+       1st Qu. & 0 \\
+       \hline
+       Median & 0 \\
+       \hline
+       Mean & 87.9 \\
+       \hline
+       3rd Qu. & 0 \\
+       \hline
+       Max. & 3770 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Hours Per Week]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 1 \\
+       \hline
+       1st Qu. & 40 \\
+       \hline
+       Median & 40 \\
+       \hline
+       Mean & 40.39 \\
+       \hline
+       3rd Qu. & 45 \\
+       \hline
+       Max. & 99 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Native Country]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       United States & 14662 \\
+       \hline
+       Mexico & 308 \\
+       \hline
+       Philippines & 97 \\
+       \hline
+       Puerto Rico & 70 \\
+       \hline
+       Germany & 69 \\
+       \hline
+       (Other) & 801 \\
+       \hline
+       NA's & 274 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Plus 50]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       No & 12435 \\
+       \hline
+       Yes & 3846 \\
+       \hline
+       \end{tabular}
+}
+\caption{Summary of the value distribution on the test data of the dataset}
+\label{tab:summarytest}
 \end{table}
@@
 \centering
+\includegraphics[width=20cm,angle=90]{item_frequency.png}
+\caption{Item frequencies of items with at least 10\% support.}
+\label{fig:arules_item_frequency}
+\end{figure}
+
+\begin{figure}
+\centering
 \includegraphics[width=10cm]{cluster_1_age_education.png}
@@
 \label{fig:cluster_1_cluster_plus50}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=20cm,angle=90]{item_frequency.png}
-\caption{Item frequencies of items with at least 10\% support.}
-\label{fig:arules_item_frequency}
 \end{figure}

### 2009-12-16 demo/demo.tex
added slides.

\documentclass[xcolor=dvipsnames,professionalfonts,smaller,presentation]{beamer} 
\usecolortheme[named=Brown]{structure} 
\useoutertheme[subsection=false]{miniframes}
\usetheme{Malmoe}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage[portuguese]{babel}
\usepackage[latin1]{inputenc}
\usepackage{alltt}
\usepackage{moreverb}
\usepackage{hyperref}
\usepackage{indentfirst}
\usepackage{lastpage}
\usepackage{url}
\usepackage{float}

\urldef{\mails}\path|ei05011@fe.up.pt,ei05028@fe.up.pt|
\title[A comparison of classification algorithms applied to the Adult Database \insertframenumber/\inserttotalframenumber]{A comparison of classification algorithms applied to the Adult Database}
\author{Flávio Cruz and João Azevedo}
\institute{Faculdade de Engenharia da Universidade do Porto\\
\mails}
\date{\today}

\begin{document}

\frame{\titlepage}

\section[Outline]{}
\frame{\tableofcontents}

\section{Problem proposed by the Dataset}
\frame
{
}

\section{Dataset Characterization}
\frame
{
}

\section{Data Mining Problem}
\frame
{
}

\section{Algorithms Used}
\frame
{
}

\section{Conducted Experiments}
\frame
{
}

\section{Results}
\frame
{
}

\section{Conclusions}
\frame
{
}

\end{document}

### 2009-12-16 demo/demo.tex
added some data analysis to the presentation.

@@
 \usecolortheme[named=Brown]{structure} 
-\useoutertheme[subsection=false]{miniframes}
+\useoutertheme{miniframes}
 \usetheme{Malmoe}
@@
 \usepackage{float}
+\usepackage{subfigure}
 
@@
 {
+    \frametitle{Problem proposed by the Dataset}
+
+    \begin{itemize}
+    \item<1-> Given an adult with some attributes, identify whether he gains more than 50000 dollars per year, or not.
+    \item<1-> Identify what social characteristics allow one person to gain more than someone else.
+    \item<1-> Understanding what affects the income can give a lot of information concerning society's behaviour.
+    \end{itemize}
 }
@@
 \section{Dataset Characterization}
+
+\subsection{The Dataset}
 \frame
 {
+    \frametitle{The Dataset}
+
+    \begin{itemize}
+    \item<1-> 48842 entries of individual information of people gathered in a 1994 census on the United States of America.
+    \begin{itemize}
+        \item<1-> 32561 records for the training set (67\%).
+        \item<1-> 16281 records for the test set (33\%).
+    \end{itemize}
+    \item<1-> Each entry is composed by a set of 14 attributes and 2 classes (winning more or less than 50K dollars a year).
+    \item<1-> There are 3620 records with missing values. The only attributes containing missing values are \textbf{Workclass}, \textbf{Occupation} and \textbf{Native-country}.
+    \item<1-> Available in \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}.
+    \end{itemize}
+}
+
+\frame
+{
+    \frametitle{Attributes}
+
+    \begin{itemize}
+        \item<1-> \textbf{Age} (continuous).
+        \item<1-> \textbf{Workclass} (categorical).
+        \item<1-> \textbf{FNLWGT} (continuous).
+        \item<1-> \textbf{Education} (ordinal).
+        \item<1-> \textbf{Education-num} (continuous).
+        \item<1-> \textbf{Marital-status} (categorical).
+        \item<1-> \textbf{Occupation} (categorical).
+        \item<1-> \textbf{Relationship} (categorical).
+        \item<1-> \textbf{Race} (categorical).
+        \item<1-> \textbf{Sex} (categorical).
+        \item<1-> \textbf{Capital-gain} (continuous).
+        \item<1-> \textbf{Capital-loss} (continuous).
+        \item<1-> \textbf{Hours-per-week} (continuous).
+        \item<1-> \textbf{Native country} (categorical).
+    \end{itemize}   
+}
+
+\subsection{Data Analysis}
+\frame
+{
+    \frametitle{Value distribution in the training set}
+
+\footnotesize
+\begin{table}
+\subtable[Age]{
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 17 \\
+       \hline
+       1st Qu. & 28 \\
+       \hline
+       Median & 37 \\
+       \hline
+       Mean & 38.58 \\
+       \hline
+       3rd Qu. & 48 \\
+       \hline
+       Max & 90 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Workclass]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       private & 22696 \\
+       \hline
+       self\_emp\_not\_inc & 2541 \\
+       \hline
+       local\_gov & 2093 \\
+       \hline
+       state\_gov & 1298 \\
+       \hline
+       self\_emp\_inc & 1116 \\
+       \hline
+       (Other) & 981 \\
+       \hline
+       NA's & 1836 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Education]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       hs\_grad & 10501 \\
+       \hline
+       some\_college & 7291 \\
+       \hline
+       bachelors & 5535 \\
+       \hline
+       masters & 1723 \\
+       \hline
+       assoc\_voc & 1382 \\
+       \hline
+       11th & 1175 \\
+       \hline
+       (Other) & 5134 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Marital Status]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       divorced & 4443 \\
+       \hline
+       married\_af\_spouse & 23 \\
+       \hline
+       married\_civ\_spouse & 14976 \\
+       \hline
+       married\_spouse\_absent & 418 \\
+       \hline
+       never\_married & 10683 \\
+       \hline
+       separated & 1025 \\
+       \hline
+       widowed & 993 \\
+       \hline
+       \end{tabular}
+}
+\end{table}
+\normalsize
+}
+
+\frame
+{
+    \frametitle{Value distribution in the training set}
+
+\footnotesize
+\begin{table}
+\subtable[Occupation]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       prof\_specialty & 4140 \\
+       \hline
+       craft\_repair & 4099 \\
+       \hline
+       exec\_managerial & 4066 \\
+       \hline
+       adm\_clerical & 3770 \\
+       \hline
+       sales & 3650 \\
+       \hline
+       (Other) & 10993 \\
+       \hline
+       NA's & 1843 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Relatonship]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       husband & 13193 \\
+       \hline
+       not\_in\_family & 8305 \\
+       \hline
+       other\_relative & 981 \\
+       \hline
+       own\_child & 5068 \\
+       \hline
+       unmarried & 3446 \\
+       \hline
+       wife & 1568 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Race]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       amer\_indian\_eskimo & 311 \\
+       \hline
+       asian\_pac\_islander & 1039 \\
+       \hline
+       black & 3124 \\
+       \hline
+       other & 271 \\
+       \hline
+       white & 27816 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Sex]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       female & 10771 \\
+       \hline
+       male & 21790 \\
+       \hline
+       \end{tabular}
+}
+\end{table}
+\normalsize
+}
+
+\frame
+{
+    \frametitle{Value distribution in the training set}
+
+\footnotesize
+\begin{table}
+\subtable[Capital Gain]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 0 \\
+       \hline
+       1st Qu. & 0 \\
+       \hline
+       Median & 0 \\
+       \hline
+       Mean & 1078 \\
+       \hline
+       3rd Qu. & 0 \\
+       \hline
+       Max. & 99999 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Capital Loss]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 0 \\
+       \hline
+       1st Qu. & 0 \\
+       \hline
+       Median & 0 \\
+       \hline
+       Mean & 87.3 \\
+       \hline
+       3rd Qu. & 0 \\
+       \hline
+       Max. & 4356 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Hours Per Week]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       Min. & 1 \\
+       \hline
+       1st Qu. & 40 \\
+       \hline
+       Median & 40 \\
+       \hline
+       Mean & 40.44 \\
+       \hline
+       3rd Qu. & 45 \\
+       \hline
+       Max. & 99 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Native Country]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       United States & 29170 \\
+       \hline
+       Mexico & 643 \\
+       \hline
+       Philippines & 198 \\
+       \hline
+       Germany & 137 \\
+       \hline
+       Canada & 121 \\
+       \hline
+       (Other) & 1709 \\
+       \hline
+       NA's & 583 \\
+       \hline
+       \end{tabular}
+}
+\qquad\qquad
+\subtable[Plus 50]{        
+       \begin{tabular}{| l | l |}
+       \hline
+       No & 24720 \\
+       \hline
+       Yes & 7841 \\
+       \hline
+       \end{tabular}
+}
+\end{table}
+\normalsize
 }

### 2009-12-16 demo/demo.tex
added more sections in the slides.

@@
 
+\frame
+{
+    \frametitle{Some obvious relationships}
+
+    \begin{itemize}
+    \item<1-> There is a higher percentage of older people gaining more than 50K than younger ones.
+    \item<1-> There is a higher percentage of people with more education gaining more than 50K than people with less studies.
+    \item<1-> There is a higher percentage of workaholics gaining more than 50K than people who work in part-time.
+    \end{itemize}
+
+   \begin{minipage}{0.30\linewidth}
+        \begin{figure}
+        \centering
+            \includegraphics[width=3.5cm]{plot_plus_50_age.png}
+        \end{figure}
+    \end{minipage}
+    \hspace{0.1cm}
+    \begin{minipage}{0.30\linewidth}
+        \begin{figure}
+        \centering
+            \includegraphics[width=3.5cm]{plot_plus_50_education_num.png}
+        \end{figure}
+    \end{minipage}
+    \hspace{0.1cm}
+    \begin{minipage}{0.30\linewidth}
+        \begin{figure}
+        \centering
+            \includegraphics[width=3.5cm]{plot_plus_50_hpw.png}
+        \end{figure}
+    \end{minipage}
+}
+
+\frame
+{
+    \frametitle{Association Rules}
+}
+
+\frame
+{
+    \frametitle{Clusters}
+}
+
 \section{Data Mining Problem}
+\frame
+{
+    \frametitle{Data Mining Problem}
+
+    \begin{itemize}
+    \item<1-> The database poses a classification problem.
+    \item<1-> There is no single classifier that works best on all given problems:
+    \begin{itemize}
+        \item<1-> Decision trees.
+        \item<1-> Rule induction.
+        \item<1-> Inductive logic programming.
+        \item<1-> Bayesian classifiers.
+    \end{itemize}
+    \end{itemize}
+}
+
+\section{Algorithms Used}
+
+\subsection{Decision Trees}
 \frame
@@
 
-\section{Algorithms Used}
+\subsection{Rule Induction}
+\frame
+{
+}
+
+\subsection{Inductive Logic Programming}
+\frame
+{
+}
+
+\subsection{Bayesian Classifiers}
 \frame

### 2009-12-16 demo/demo.tex
added some conclusions.

@@
 {
+    \frametitle{Results}
+
+\begin{table}[ht]
+    \begin{center}
+    \begin{tabular}{ | l | l | }
+    \hline
+    \textbf{Algorithm} & \textbf{Error rate} \\ \hline
+    CN2 & 15.20 \\ \hline
+    ID3 & 15.68 \\ \hline
+    ADTree & 18.49 \\ \hline
+    C4.5 & 18.84 \\ \hline
+    Bayesian Networks (Simulated Annealing) & 19.15 \\ \hline
+    Bayesian Networks (Hill Climbing) & 20.03 \\ \hline
+    ILP (Aleph) & 20.45 \\ \hline
+    Random Forest & 21.01 \\ \hline
+    Bayesian Networks (Genetic Search) & 21.26 \\ \hline
+    CART & 21.34 \\ \hline
+    \end{tabular}
+    \label{tbl:overall_results}
+    \end{center}
+\end{table}
 }
@@
 {
+    \frametitle{Conclusions}
+
+    \begin{itemize}
+    \item<1-> ID3 and C4.5 are very fast and build classifiers with a decent error rate.
+    \item<1-> ILP and CN2 build more expressive classifiers, capable of a better interpretation by a human agent, but take longer to finish.
+    \item<1-> The worst error rate is below 25\%.
+    \item<1-> There is a lot of software tools that support the building of classifiers, namely Weka and RapidMiner.
+    \begin{itemize}
+        \item<1-> Open source.
+        \item<1-> Easy to use.
+        \item<1-> APIs for integration with Java applications.
+    \end{itemize}
+    \end{itemize}
+}
+
+\frame
+{
+    \frametitle{Conclusions}
+
+    \begin{itemize}
+    \item<1-> Having a high capital gain clearly contributes to gain more than 50K dollars per year.
+    \item<1-> Middle-aged husbands have a higher tendency to gain more than 50K dollars per year.
+    \item<1-> A high percentage of people who gain less than 50K dollars per year are own childs.
+    \item<1-> People coming from countries such as Taiwan, Thailand or Cambodia usually have higher education and gain more than 50K dollars per year.
+    \end{itemize}
 }

### 2009-12-16 demo/demo.tex
added times to results.

@@
     \begin{center}
-    \begin{tabular}{ | l | l | }
+    \begin{tabular}{ | l | l | l | }
     \hline
-    \textbf{Algorithm} & \textbf{Error rate} \\ \hline
-    CN2 & 15.20 \\ \hline
-    ID3 & 15.68 \\ \hline
-    ADTree & 18.49 \\ \hline
-    C4.5 & 18.84 \\ \hline
-    Bayesian Networks (Simulated Annealing) & 19.15 \\ \hline
-    Bayesian Networks (Hill Climbing) & 20.03 \\ \hline
-    ILP (Aleph) & 20.45 \\ \hline
-    Random Forest & 21.01 \\ \hline
-    Bayesian Networks (Genetic Search) & 21.26 \\ \hline
-    CART & 21.34 \\ \hline
+    \textbf{Algorithm} & \textbf{Error rate} & \textbf{Time} \\ \hline
+    CN2 & 15.20 & 03m29s \\ \hline
+    ID3 & 15.68 & 00m14s \\ \hline
+    ADTree & 18.49 & 01m12s \\ \hline
+    C4.5 & 18.84 & 00m08s \\ \hline
+    Bayesian Networks (Simulated Annealing) & 19.15 & 00m07s \\ \hline
+    Bayesian Networks (Hill Climbing) & 20.03 & 00m07s \\ \hline
+    ILP (Aleph) & 20.45 & $>$ 6hours \\ \hline
+    Random Forest & 21.01 & 00m08s \\ \hline
+    Bayesian Networks (Genetic Search) & 21.26 & 00m04s \\ \hline
+    CART & 21.34 & 07m05s \\ \hline
     \end{tabular}

### 2009-12-16 demo/demo.tex
fixing tilde character.

--- 2009-12-16 demo/demo.tex
@@
     \begin{center}
-    \begin{tabular}{ | l | l | }
+    \begin{tabular}{ | l | l | l | }
     \hline
-    \textbf{Algorithm} & \textbf{Error rate} \\ \hline
-    CN2 & 15.20 \\ \hline
-    ID3 & 15.68 \\ \hline
-    ADTree & 18.49 \\ \hline
-    C4.5 & 18.84 \\ \hline
-    Bayesian Networks (Simulated Annealing) & 19.15 \\ \hline
-    Bayesian Networks (Hill Climbing) & 20.03 \\ \hline
-    ILP (Aleph) & 20.45 \\ \hline
-    Random Forest & 21.01 \\ \hline
-    Bayesian Networks (Genetic Search) & 21.26 \\ \hline
-    CART & 21.34 \\ \hline
+    \textbf{Algorithm} & \textbf{Error rate} & \textbf{Time} \\ \hline
+    CN2 & 15.20 & 03m29s \\ \hline
+    ID3 & 15.68 & 00m14s \\ \hline
+    ADTree & 18.49 & 01m12s \\ \hline
+    C4.5 & 18.84 & 00m08s \\ \hline
+    Bayesian Networks (Simulated Annealing) & 19.15 & 00m07s \\ \hline
+    Bayesian Networks (Hill Climbing) & 20.03 & 00m07s \\ \hline
+    ILP (Aleph) & 20.45 & $\sim$6hours \\ \hline
+    Random Forest & 21.01 & 00m08s \\ \hline
+    Bayesian Networks (Genetic Search) & 21.26 & 00m04s \\ \hline
+    CART & 21.34 & 07m05s \\ \hline
     \end{tabular}

### 2009-12-16 demo/demo.tex
added some cluster description.

@@
     \frametitle{Clusters}
+
+    \begin{itemize}
+    \item<1-> Clustering can be used to detect similar individuals and to build classes in the absence of them.
+    \item<1-> Generating 6 clusters with only three variables (\textbf{age}, \textbf{education\_num} and \textbf{hours\_per\_week}) leads to some interesting observations:
+    \begin{itemize}
+        \item<1-> Younger people are grouped together.
+        \item<1-> People with different levels of education belong to different clusters.
+    \end{itemize}
+    \end{itemize}
 }

### 2009-12-16 demo/demo.tex
added cn2 description to slides.
fixing presentation footer.

@@
 \urldef{\mails}\path|ei05011@fe.up.pt,ei05028@fe.up.pt|
-\title[A comparison of classification algorithms applied to the Adult Database \insertframenumber/\inserttotalframenumber]{A comparison of classification algorithms applied to the Adult Database}
+\title[Classification on the Adult Database \insertframenumber/\inserttotalframenumber]{A comparison of classification algorithms applied to the Adult Database}
 \author{Flávio Cruz and João Azevedo}
@@
 {
+    \frametitle{Rule Induction}
+
+    \begin{itemize}
+    \item<1-> Peter Clark's CN2 software was used to induce rules:
+    \begin{itemize}
+        \item<1-> \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}.
+    \end{itemize}
+    \item<1-> CN2 can generate both ordered and unordered set of rules.
+    \item<1-> Peter Clark's software can use different error estimations:
+    \begin{itemize}
+        \item<1-> Laplacian.
+        \item<1-> Naive.
+        \item<1-> Information Gain.
+        \item<1-> Modified Entropy.
+    \end{itemize}
+    \end{itemize}
 }

### 2009-12-16 demo/demo.tex
association rules slide

@@
 \usepackage{lastpage}
+\usepackage{array}
 \usepackage{url}
@@
     \frametitle{Association Rules}
+
+    \begin{itemize}
+      \item <1-> Association rules can be used to discover interesting relations and patterns between attributes.
+
+    \end{itemize}
+
+    \begin{table}[ht]
+        \begin{center}
+        \scalebox{1.0}{
+
+        \begin{tabular}{ | m{5cm} | c | c | c | }
+        \hline
+        \textbf{Rule / LHS} & \textbf{RHS} & \textbf{Support} & \textbf{Confidence} \\ \hline
+        \footnotesize{age=young, hours\_per\_week=part\_time} & $<$50K & 0.05730782 & 1 \\ \hline
+        \footnotesize{marital\_status=never\_married, relationship=own\_child, hours\_per\_week=part\_time} & $<$50K & 0.04689659 & 1 \\ \hline
+        \footnotesize{age=young, relationship=own\_child, hours\_per\_week=part\_time} & $<$50K & 0.04404042 & 1 \\ \hline
+        \footnotesize{relationship=husband, capital\_gain=high} & $>$50K & 0.02251159 & 0.9918809 \\ \hline
+        \footnotesize{relationship=husband, race=white, sex=male, capital\_gain=high} & $>$50K & 0.02063819 & 0.9926145 \\ \hline
+        \end{tabular}}
+        \label{tbl:arules_income}
+        \end{center}
+    \end{table}
 }

### 2009-12-16 demo/demo.tex
algorithms used slide
move things around a bit

@@
 \section{Algorithms Used}
-
-\subsection{Decision Trees}
 \frame
 {
+  \frametitle{Algorithms Used}
+
+  \begin{itemize}
+    \item<1-> Decision trees
+
+    \begin{itemize}
+      \item<1-> ID3
+      \item<1-> C4.5
+      \item<1-> CART
+      \item<1-> Random Forest
+      \item<1-> Alternating Decision Trees (ADTree)
+    \end{itemize}
+
+    \item<1-> Rule Induction
+
+    \begin{itemize}
+      \item <1-> CN2
+    \end{itemize}
+
+    \item<1-> Inductive Logic Programming: using Aleph
+
+    \item<1-> Bayesian Networks. Using different search algorithms:
+
+    \begin{itemize}
+      \item<1-> Genetic Search
+      \item<1-> Hill Climbing
+      \item<1-> Simulated Annealing
+    \end{itemize}
+
+  \end{itemize}
+}
+
+\section{Conducted Experiments}
+\frame
+{
+  \frametitle{Conducted Experiments}
+
+  \begin{itemize}
+    \item<1-> Classifiers were built using the train set and then tested against the test set
+    \item<1-> C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka
+    \begin{itemize}
+      \item<1-> Various parameters available in Weka algorithms were experimented in hope of getting better error rates
+      \item<1-> Every missing value was replaced with modes and means
+      \item<1-> Two different classifiers were trained for each algorithm, one using only discretized attributes
+                and unchanged attributes to analyze how the algorithms behaves when using numeric attributes
+    \end{itemize}
+
+  \end{itemize}
+}
+
+\frame
+{
+  \frametitle{Conducted Experiments}
+  % colocar ILP CN2 ID3!
 }
@@
     \end{itemize}
-}
-
-\subsection{Inductive Logic Programming}
-\frame
-{
-}
-
-\subsection{Bayesian Classifiers}
-\frame
-{
-}
-
-\section{Conducted Experiments}
-\frame
-{
 }

### 2009-12-16 demo/demo.tex
c4.5, random forest and bayes slides

@@
       \item<1-> Two different classifiers were trained for each algorithm, one using only discretized attributes
-                and unchanged attributes to analyze how the algorithms behaves when using numeric attributes
+                and another with the attributes unmodified
+      \item<1-> Analyze how the algorithm behaves when using numeric attributes
     \end{itemize}
-
   \end{itemize}
@@
   % colocar ILP CN2 ID3!
+}
+
+\subsection{C4.5}
+\frame
+{
+  \frametitle{C4.5}
+
+  \begin{itemize}
+    \item Weka supports a confidence threshold parameter that dictates how much pruning the algorithm will do
+    \item The bigger the threshold the less pruning the algorithm does
+    \item When using only discretized data the algorithm gives error rates of $\approx$25\%
+  \end{itemize}
+
+  \begin{table}[ht]
+    \begin{center}
+    \scalebox{0.8}{
+    \begin{tabular}{ | l | c | c | c | c | c |}
+      \hline
+      \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
+      -U & 9.48 & 21.58 & 0m7.525s & 6422 & 5491 \\ \hline
+      -C 0.05 & 13.78 & 19.49 & 0m7.667s & 149 & 105 \\ \hline
+      -C 0.1 & 13.37 & 19.06 & 0m7.969s & 278 & 212 \\ \hline
+      -C 0.15 & 12.99 & 18.84 & 0m7.708s & 361 & 278 \\ \hline
+      -C 0.2 & 12.95 & 18.89 & 0m7.812s & 373 & 284 \\ \hline
+      -C 0.25 & 12.86 & 18.86 & 0m7.974s & 433 & 331 \\ \hline
+      -C 0.3 & 12.11 & 19.73 & 0m8.064s & 806 & 624 \\ \hline
+      -C 0.35 & 12.02 & 19.77 & 0m8.035s & 878 & 678 \\ \hline
+      -C 0.4 & 11.54 & 20.47 & 0m8.255s & 1264 & 995 \\ \hline
+      -C 0.5 & 11.22 & 20.28 & 0m8.117s & 1603 & 1257 \\ \hline
+    \end{tabular}}
+    \end{center}
+  \end{table}
+}
+
+\subsection{Random Forest}
+\frame
+{
+  \frametitle{Random Forest}
+
+  \begin{itemize}
+    \item<1-> It is an ensemble classifier consisting of many decision trees
+    \item<1-> Outputs the class that is the mode of the class's output by individual trees
+    \item<1-> Better results require more time building the classifier
+    \item<1-> Large quantities of memory needed: 1GB to compute with 400 trees
+    \item<1-> Over-fitting
+  \end{itemize}
+
+  \begin{table}[ht]
+    \begin{center}
+    \scalebox{0.75}{
+    \begin{tabular}{ | l | c | c | c |}
+      \hline
+      \textbf{Trees} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+      2 & 5.72 & 25.70 & 0m4.882s \\ \hline
+      4 & 3.81 & 23.40 & 0m6.086s \\ \hline
+      5 & 3.20 & 23.43 & 0m6.839s \\ \hline
+      7 & 2.86 & 21.01 & 0m8.190s \\ \hline
+      10 & 2.69 & 21.15 & 0m10.162s \\ \hline
+      20 & 2.46 & 20.83 & 0m16.970s \\ \hline
+      40 & 2.44 & 21.26 & 0m30.488s \\ \hline
+      80 & 2.44 & 21.70 & 1m1.347s \\ \hline
+      100 & 2.44 & 21.75 & 1m15.592s \\ \hline
+      400 & 2.44 & 21.62 & 7m51.624s \\ \hline
+    \end{tabular}}
+    \end{center}
+  \end{table}
+}
+
+\subsection{Bayesian Networks}
+\frame
+{
+  \frametitle{Bayesian Networks}
+
+  \begin{itemize}
+    \item<1-> Based on conditional probabilities and the Bayes theorem
+    \item<1-> Represents a set of variables and their conditional independencies via a directed acyclic graph and a set of probability tables
+    \item<1-> The algorithm first learns the network structure / graph and then learns the probability tables
+    \item<1-> We experimented with different algorithms to learn the network structure:
+     \begin{itemize}
+      \item<1-> Genetic Search
+      \item<1-> Hill Climbing
+      \item<1-> Simulated Annealing
+     \end{itemize}
+    \item<1-> Weka provides a very rich set of parameters for each search algorithm
+  \end{itemize}
+
 }
@@
     \item<1-> The worst error rate is below 25\%.
+    \item<1-> Using numeric information results in better error rates
     \item<1-> There is a lot of software tools that support the building of classifiers, namely Weka and RapidMiner.

### 2009-12-16 demo/demo.tex
added more details on CN2 description.

@@
 
-    \item<1-> Inductive Logic Programming: using Aleph
+    \item<1-> Inductive Logic Programming
+    \begin{itemize}
+        \item<1-> Aleph
+    \end{itemize}
 
@@
   \frametitle{Conducted Experiments}
-  % colocar ILP CN2 ID3!
+
+  \begin{itemize}
+    \item<1-> RapidMiner was used to build a classifier with the ID3 algorithm.
+    \item<1-> Peter Clark's CN2 was used to induce rules.
+    \begin{itemize}
+        \item<1-> \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}.
+    \end{itemize}
+    \item<1-> Aleph was used to inductively generate logic clauses.
+    \begin{itemize}
+        \item<1-> \url{http://www.comlab.ox.ac.uk/activities/machinelearning/Aleph/aleph.html}.
+    \end{itemize}
+  \end{itemize}
+  % colocar ILP ID3!
 }
@@
 
-\subsection{Rule Induction}
+\subsection{CN2}
 \frame
 {
-    \frametitle{Rule Induction}
+    \frametitle{CN2}
 
     \begin{itemize}
-    \item<1-> Peter Clark's CN2 software was used to induce rules:
-    \begin{itemize}
-        \item<1-> \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}.
-    \end{itemize}
     \item<1-> CN2 can generate both ordered and unordered set of rules.
@@
     \end{itemize}
+
+\footnotesize
+\begin{table}[ht]
+\begin{center}
+\begin{tabular}{ | l | l | l | l | l |}
+    \hline
+    \textbf{Algorithm} & \textbf{Error estimate} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    Unordered & Laplacian & 13.3 & 15.2 & 3m29s \\ \hline
+    Unordered & Naive & - & - & - \\ \hline
+    Ordered & Laplacian & 0 & 17.6 & 1m45s \\ \hline
+    Ordered & Naive & 0 & 19.9 & 12m21s \\ \hline
+    Ordered & Information Gain & 18.9 & 18.9 & 0m5s \\ \hline
+    Ordered & Modified Entropy & 18.9 & 19.1 & 0m20s \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+\normalsize
 }

### 2009-12-16 demo/demo.tex
added ID3 slide.

@@
 
+\subsection{ID3}
+\frame
+{
+    \frametitle{ID3}
+
+    \begin{itemize}
+        \item ID3 in RapidMiner is implemented in two different flavours: one that supports numerical values, one that doesn't.
+        \item An absolute discretization was conducted on numerical values for the version that doesn't support them.
+        \item RapidMiner supports different splitting criterions: Gain Ratio, Information Gain, Gini Index or Accuracy.
+    \end{itemize}
+
+\footnotesize
+\begin{table}[ht]
+  \begin{center}
+  \begin{tabular}{ | l | c | c | c | c | c |}
+    \hline
+    \textbf{Type} & \textbf{Criterion} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
+    ID3 & Gain Ratio & 18.38 & 18.61 & 0m04s \\ \hline
+    ID3 & Information Gain & 18.48 & 19.05 & 0m03s \\ \hline
+    ID3 & Gini Index & 18.94 & 19.15 & 0m04s \\ \hline
+    ID3 & Accuracy & 20.41 & 20.24 & 0m04s \\ \hline
+    ID3 Numerical & Gain Ratio & 14.93 & 15.68 & 0m14s \\ \hline
+    ID3 Numerical & Information Gain & 15.68 & 16.42 & 0m06s \\ \hline
+    ID3 Numerical & Gini Index & 16.10 & 16.96 & 0m07s \\ \hline
+    ID3 Numerical & Accuracy & 17.99 & 17.92 & 0m12s \\ \hline
+  \end{tabular}
+  \end{center}
+\end{table}
+\normalsize
+}
+
 \subsection{C4.5}

### 2009-12-16 demo/demo.tex
added ilp slides.

@@
 \usepackage{subfigure}
+\usepackage{multirow}
 
@@
 
+\subsection{ILP}
+\frame[containsverbatim]
+{
+    \frametitle{ILP}
+
+    \begin{itemize}
+    \item Aleph was used to genereate logic clauses.
+    \item Uses inverse entailment to generate the bottom clause.
+    \item The search for the best clause is implemented by a restricted branch-and-bound algorithm.
+    \item The generation of hypothesis takes around 6 hours and produces an error rate on the test set of around 20\%.
+    \item Some interesting clauses:
+    \end{itemize}
+
+\scriptsize
+\begin{verbatim}
+more_than_50K(A) :- middle_aged(A), very_high_capital_gain(A), 
+                    is_native_of_united_states(A).
+more_than_50K(A) :- education_prof_school(A), wife(A), workaholic(A).
+more_than_50K(A) :- senior(A), work_private(A), very_high_capital_gain(A).
+more_than_50K(A) :- senior(A), work_federal_gov(A), education_doctorate(A).
+more_than_50K(A) :- education_doctorate(A), medium_capital_loss(A), workaholic(A).
+more_than_50K(A) :- work_self_emp_inc(A), high_capital_gain(A), full_time_worker(A).
+more_than_50K(A) :- middle_aged(A), education_masters(A), is_native_of_iran(A).
+more_than_50K(A) :- senior(A), full_time_worker(A), is_native_of_cambodia(A).
+\end{verbatim}
+\normalsize
+}
+
 \section{Results}

### 2009-12-16 demo/demo.tex
general latex syntax fixes.

@@
     \begin{itemize}
-    \item<1-> Given an adult with some attributes, identify whether he gains more than 50000 dollars per year, or not.
-    \item<1-> Identify what social characteristics allow one person to gain more than someone else.
-    \item<1-> Understanding what affects the income can give a lot of information concerning society's behaviour.
+    \item Given an adult with some attributes, identify whether he gains more than 50000 dollars per year, or not.
+    \item Identify what social characteristics allow one person to gain more than someone else.
+    \item Understanding what affects the income can give a lot of information concerning society's behaviour.
     \end{itemize}
@@
     \begin{itemize}
-    \item<1-> 48842 entries of individual information of people gathered in a 1994 census on the United States of America.
+    \item 48842 entries of individual information of people gathered in a 1994 census on the United States of America.
     \begin{itemize}
-        \item<1-> 32561 records for the training set (67\%).
-        \item<1-> 16281 records for the test set (33\%).
+        \item 32561 records for the training set (67\%).
+        \item 16281 records for the test set (33\%).
     \end{itemize}
-    \item<1-> Each entry is composed by a set of 14 attributes and 2 classes (winning more or less than 50K dollars a year).
-    \item<1-> There are 3620 records with missing values. The only attributes containing missing values are \textbf{Workclass}, \textbf{Occupation} and \textbf{Native-country}.
-    \item<1-> Available in \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}.
+    \item Each entry is composed by a set of 14 attributes and 2 classes (winning more or less than 50K dollars a year).
+    \item There are 3620 records with missing values. The only attributes containing missing values are \textbf{Workclass}, \textbf{Occupation} and \textbf{Native-country}.
+    \item Available in \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}.
     \end{itemize}
@@
     \begin{itemize}
-        \item<1-> \textbf{Age} (continuous).
-        \item<1-> \textbf{Workclass} (categorical).
-        \item<1-> \textbf{FNLWGT} (continuous).
-        \item<1-> \textbf{Education} (ordinal).
-        \item<1-> \textbf{Education-num} (continuous).
-        \item<1-> \textbf{Marital-status} (categorical).
-        \item<1-> \textbf{Occupation} (categorical).
-        \item<1-> \textbf{Relationship} (categorical).
-        \item<1-> \textbf{Race} (categorical).
-        \item<1-> \textbf{Sex} (categorical).
-        \item<1-> \textbf{Capital-gain} (continuous).
-        \item<1-> \textbf{Capital-loss} (continuous).
-        \item<1-> \textbf{Hours-per-week} (continuous).
-        \item<1-> \textbf{Native country} (categorical).
+        \item \textbf{Age} (continuous).
+        \item \textbf{Workclass} (categorical).
+        \item \textbf{FNLWGT} (continuous).
+        \item \textbf{Education} (ordinal).
+        \item \textbf{Education-num} (continuous).
+        \item \textbf{Marital-status} (categorical).
+        \item \textbf{Occupation} (categorical).
+        \item \textbf{Relationship} (categorical).
+        \item \textbf{Race} (categorical).
+        \item \textbf{Sex} (categorical).
+        \item \textbf{Capital-gain} (continuous).
+        \item \textbf{Capital-loss} (continuous).
+        \item \textbf{Hours-per-week} (continuous).
+        \item \textbf{Native country} (categorical).
     \end{itemize}   
@@
     \begin{itemize}
-    \item<1-> There is a higher percentage of older people gaining more than 50K than younger ones.
-    \item<1-> There is a higher percentage of people with more education gaining more than 50K than people with less studies.
-    \item<1-> There is a higher percentage of workaholics gaining more than 50K than people who work in part-time.
+    \item There is a higher percentage of older people gaining more than 50K than younger ones.
+    \item There is a higher percentage of people with more education gaining more than 50K than people with less studies.
+    \item There is a higher percentage of workaholics gaining more than 50K than people who work in part-time.
     \end{itemize}
@@
     \begin{itemize}
-      \item <1-> Association rules can be used to discover interesting relations and patterns between attributes.
+      \item  Association rules can be used to discover interesting relations and patterns between attributes.
 
@@
     \begin{itemize}
-    \item<1-> Clustering can be used to detect similar individuals and to build classes in the absence of them.
-    \item<1-> Generating 6 clusters with only three variables (\textbf{age}, \textbf{education\_num} and \textbf{hours\_per\_week}) leads to some interesting observations:
+    \item Clustering can be used to detect similar individuals and to build classes in the absence of them.
+    \item Generating 6 clusters with only three variables (\textbf{age}, \textbf{education\_num} and \textbf{hours\_per\_week}) leads to some interesting observations:
     \begin{itemize}
-        \item<1-> Younger people are grouped together.
-        \item<1-> People with different levels of education belong to different clusters.
+        \item Younger people are grouped together.
+        \item People with different levels of education belong to different clusters.
     \end{itemize}
@@
     \begin{itemize}
-    \item<1-> The database poses a classification problem.
-    \item<1-> There is no single classifier that works best on all given problems:
+    \item The database poses a classification problem.
+    \item There is no single classifier that works best on all given problems:
     \begin{itemize}
-        \item<1-> Decision trees.
-        \item<1-> Rule induction.
-        \item<1-> Inductive logic programming.
-        \item<1-> Bayesian classifiers.
+        \item Decision trees.
+        \item Rule induction.
+        \item Inductive logic programming.
+        \item Bayesian classifiers.
     \end{itemize}
@@
   \begin{itemize}
-    \item<1-> Decision trees
+    \item Decision trees
 
     \begin{itemize}
-      \item<1-> ID3
-      \item<1-> C4.5
-      \item<1-> CART
-      \item<1-> Random Forest
-      \item<1-> Alternating Decision Trees (ADTree)
+      \item ID3
+      \item C4.5
+      \item CART
+      \item Random Forest
+      \item Alternating Decision Trees (ADTree)
     \end{itemize}
 
-    \item<1-> Rule Induction
+    \item Rule Induction
 
     \begin{itemize}
-      \item <1-> CN2
+      \item  CN2
     \end{itemize}
 
-    \item<1-> Inductive Logic Programming
+    \item Inductive Logic Programming
     \begin{itemize}
-        \item<1-> Aleph
+        \item Aleph
     \end{itemize}
 
-    \item<1-> Bayesian Networks. Using different search algorithms:
+    \item Bayesian Networks. Using different search algorithms:
 
     \begin{itemize}
-      \item<1-> Genetic Search
-      \item<1-> Hill Climbing
-      \item<1-> Simulated Annealing
+      \item Genetic Search
+      \item Hill Climbing
+      \item Simulated Annealing
     \end{itemize}
@@
   \begin{itemize}
-    \item<1-> Classifiers were built using the train set and then tested against the test set
-    \item<1-> C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka
+    \item Classifiers were built using the train set and then tested against the test set
+    \item C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka
     \begin{itemize}
-      \item<1-> Various parameters available in Weka algorithms were experimented in hope of getting better error rates
-      \item<1-> Every missing value was replaced with modes and means
-      \item<1-> Two different classifiers were trained for each algorithm, one using only discretized attributes
+      \item Various parameters available in Weka algorithms were experimented in hope of getting better error rates
+      \item Every missing value was replaced with modes and means
+      \item Two different classifiers were trained for each algorithm, one using only discretized attributes
                 and another with the attributes unmodified
-      \item<1-> Analyze how the algorithm behaves when using numeric attributes
+      \item Analyze how the algorithm behaves when using numeric attributes
     \end{itemize}
@@
   \begin{itemize}
-    \item<1-> RapidMiner was used to build a classifier with the ID3 algorithm.
-    \item<1-> Peter Clark's CN2 was used to induce rules.
+    \item RapidMiner was used to build a classifier with the ID3 algorithm.
+    \item Peter Clark's CN2 was used to induce rules.
     \begin{itemize}
-        \item<1-> \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}.
+        \item \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}.
     \end{itemize}
-    \item<1-> Aleph was used to inductively generate logic clauses.
+    \item Aleph was used to inductively generate logic clauses.
     \begin{itemize}
-        \item<1-> \url{http://www.comlab.ox.ac.uk/activities/machinelearning/Aleph/aleph.html}.
+        \item \url{http://www.comlab.ox.ac.uk/activities/machinelearning/Aleph/aleph.html}.
     \end{itemize}
   \end{itemize}
-  % colocar ILP ID3!
 }
@@
   \begin{itemize}
-    \item<1-> It is an ensemble classifier consisting of many decision trees
-    \item<1-> Outputs the class that is the mode of the class's output by individual trees
-    \item<1-> Better results require more time building the classifier
-    \item<1-> Large quantities of memory needed: 1GB to compute with 400 trees
-    \item<1-> Over-fitting
+    \item It is an ensemble classifier consisting of many decision trees
+    \item Outputs the class that is the mode of the class's output by individual trees
+    \item Better results require more time building the classifier
+    \item Large quantities of memory needed: 1GB to compute with 400 trees
+    \item Over-fitting
   \end{itemize}
@@
   \begin{itemize}
-    \item<1-> Based on conditional probabilities and the Bayes theorem
-    \item<1-> Represents a set of variables and their conditional independencies via a directed acyclic graph and a set of probability tables
-    \item<1-> The algorithm first learns the network structure / graph and then learns the probability tables
-    \item<1-> We experimented with different algorithms to learn the network structure:
+    \item Based on conditional probabilities and the Bayes theorem
+    \item Represents a set of variables and their conditional independencies via a directed acyclic graph and a set of probability tables
+    \item The algorithm first learns the network structure / graph and then learns the probability tables
+    \item We experimented with different algorithms to learn the network structure:
      \begin{itemize}
-      \item<1-> Genetic Search
-      \item<1-> Hill Climbing
-      \item<1-> Simulated Annealing
+      \item Genetic Search
+      \item Hill Climbing
+      \item Simulated Annealing
      \end{itemize}
-    \item<1-> Weka provides a very rich set of parameters for each search algorithm
+    \item Weka provides a very rich set of parameters for each search algorithm
   \end{itemize}
@@
     \begin{itemize}
-    \item<1-> CN2 can generate both ordered and unordered set of rules.
-    \item<1-> Peter Clark's software can use different error estimations:
+    \item CN2 can generate both ordered and unordered set of rules.
+    \item Peter Clark's software can use different error estimations:
     \begin{itemize}
-        \item<1-> Laplacian.
-        \item<1-> Naive.
-        \item<1-> Information Gain.
-        \item<1-> Modified Entropy.
+        \item Laplacian.
+        \item Naive.
+        \item Information Gain.
+        \item Modified Entropy.
     \end{itemize}
@@
     \begin{itemize}
-    \item<1-> ID3 and C4.5 are very fast and build classifiers with a decent error rate.
-    \item<1-> ILP and CN2 build more expressive classifiers, capable of a better interpretation by a human agent, but take longer to finish.
-    \item<1-> The worst error rate is below 25\%.
-    \item<1-> Using numeric information results in better error rates
-    \item<1-> There is a lot of software tools that support the building of classifiers, namely Weka and RapidMiner.
+    \item ID3 and C4.5 are very fast and build classifiers with a decent error rate.
+    \item ILP and CN2 build more expressive classifiers, capable of a better interpretation by a human agent, but take longer to finish.
+    \item The worst error rate is below 25\%.
+    \item Using numeric information results in better error rates
+    \item There is a lot of software tools that support the building of classifiers, namely Weka and RapidMiner.
     \begin{itemize}
-        \item<1-> Open source.
-        \item<1-> Easy to use.
-        \item<1-> APIs for integration with Java applications.
+        \item Open source.
+        \item Easy to use.
+        \item APIs for integration with Java applications.
     \end{itemize}
@@
     \begin{itemize}
-    \item<1-> Having a high capital gain clearly contributes to gain more than 50K dollars per year.
-    \item<1-> Middle-aged husbands have a higher tendency to gain more than 50K dollars per year.
-    \item<1-> A high percentage of people who gain less than 50K dollars per year are own childs.
-    \item<1-> People coming from countries such as Taiwan, Thailand or Cambodia usually have higher education and gain more than 50K dollars per year.
+    \item Having a high capital gain clearly contributes to gain more than 50K dollars per year.
+    \item Middle-aged husbands have a higher tendency to gain more than 50K dollars per year.
+    \item A high percentage of people who gain less than 50K dollars per year are own childs.
+    \item People coming from countries such as Taiwan, Thailand or Cambodia usually have higher education and gain more than 50K dollars per year.
     \end{itemize}

### 2009-12-17 demo/demo.tex
fixing typos in the presentation.

@@
     \end{itemize}
-    \item Each entry is composed by a set of 14 attributes and 2 classes (winning more or less than 50K dollars a year).
+    \item Each entry is composed by a set of 14 attributes and a class (winning more or less than 50K dollars a year).
     \item There are 3620 records with missing values. The only attributes containing missing values are \textbf{Workclass}, \textbf{Occupation} and \textbf{Native-country}.
-    \item Available in \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}.
+    \item Available in \url{ftp://ftp.ics.uci.edu/pub/machine-learning-databases/adult}
     \end{itemize}
@@
   \begin{itemize}
-    \item Classifiers were built using the train set and then tested against the test set
-    \item C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka
+    \item Classifiers were built using the train set and then tested against the test set.
+    \item C4.5, Random Forest, CART, ADTree and Bayesian classifiers built using Weka:
     \begin{itemize}
-      \item Various parameters available in Weka algorithms were experimented in hope of getting better error rates
-      \item Every missing value was replaced with modes and means
+      \item Various parameters available in Weka algorithms were experimented in hope of getting better error rates.
+      \item Every missing value was replaced with modes and means.
       \item Two different classifiers were trained for each algorithm, one using only discretized attributes
-                and another with the attributes unmodified
-      \item Analyze how the algorithm behaves when using numeric attributes
+                and another with the attributes unmodified.
+      \item Analyze how the algorithm behaves when using numeric attributes.
     \end{itemize}
@@
     \begin{itemize}
-        \item \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}.
+        \item \url{http://www.cs.utexas.edu/users/pclark/software/\#cn2}
     \end{itemize}
@@
     \begin{itemize}
-        \item \url{http://www.comlab.ox.ac.uk/activities/machinelearning/Aleph/aleph.html}.
+        \item \url{http://www.comlab.ox.ac.uk/activities/machinelearning/Aleph/aleph.html}
     \end{itemize}
@@
   \begin{itemize}
-    \item Weka supports a confidence threshold parameter that dictates how much pruning the algorithm will do
-    \item The bigger the threshold the less pruning the algorithm does
-    \item When using only discretized data the algorithm gives error rates of $\approx$25\%
+    \item Weka supports a confidence threshold parameter that dictates how much pruning the algorithm will do.
+    \item The bigger the threshold the less pruning the algorithm does.
+    \item When using only discretized data the algorithm gives error rates of $\approx$25\%.
   \end{itemize}
@@
   \begin{itemize}
-    \item It is an ensemble classifier consisting of many decision trees
-    \item Outputs the class that is the mode of the class's output by individual trees
-    \item Better results require more time building the classifier
-    \item Large quantities of memory needed: 1GB to compute with 400 trees
-    \item Over-fitting
+    \item It is an ensemble classifier consisting of many decision trees.
+    \item Outputs the class that is the mode of the class's output by individual trees.
+    \item Better results require more time building the classifier.
+    \item Large quantities of memory needed: 1GB to compute with 400 trees.
+    \item Over-fitting.
   \end{itemize}
@@
   \begin{itemize}
-    \item Based on conditional probabilities and the Bayes theorem
-    \item Represents a set of variables and their conditional independencies via a directed acyclic graph and a set of probability tables
-    \item The algorithm first learns the network structure / graph and then learns the probability tables
+    \item Based on conditional probabilities and the Bayes theorem.
+    \item Represents a set of variables and their conditional independencies via a directed acyclic graph and a set of probability tables.
+    \item The algorithm first learns the network structure / graph and then learns the probability tables.
     \item We experimented with different algorithms to learn the network structure:
@@
      \end{itemize}
-    \item Weka provides a very rich set of parameters for each search algorithm
+    \item Weka provides a very rich set of parameters for each search algorithm.
   \end{itemize}
